### 2022-02-08 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers & Casey Sakamoto"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lmerTest)
library(nlme)
library(broom.mixed)
library(segmented)
library(performance)
library(AGD)
library(rspiro)
library(lubridate)
library(plotly)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r include=FALSE}
# Import cleaned data from Casey
df = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
df$X = NULL
# Fill in MRN and CFF ID for everyone
df = df %>% 
  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
# Assume that blanks for pancreatic and G tube indicate "no"
df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
# Convert dates
date_vars = c("DOB","Start","Date")
df[,date_vars] = lapply(df[,date_vars], function(d){
  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
})
# Time from start and filtering
df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
# Add UM (they did not provide dates, so need Time column)
um_a1cs = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
um_bmi = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
um_demo = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
um_ogtt = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
um_pfts = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
um = full_join(um_a1cs,um_bmi)
um = full_join(um,um_ogtt)
um = full_join(um,um_pfts)
um = left_join(um,um_demo)
um = um %>% filter(!is.na(CFRD.Status))
df = full_join(df,um)
# Filter and fill constants
df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
  ungroup() %>% filter(!is.na(Time),abs(Time) <= 365,!is.na(CFF.ID),
                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
df$Sex = as.factor(df$Sex)
levels(df$Sex) = c("F","M")
# BMI category
df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days")),df$Age,na.rm = T)/365.25
df$BMIz = y2z(df$BMI,x = df$Age,sex = df$Sex,ref=nl4.bmi)
df$BMI_group = apply(df,1,function(r){
  age = as.numeric(r["Age"])
  if(!is.na(age) & age > 19){
    bmi = as.numeric(r["BMI"])
    group = cut(bmi,c(-Inf,18.5,25,30,35,40,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)",
                   "Obese (Class II)","Obese (Class III+)"),right = F)
  } else if (!is.na(age) & age <= 19){
    z = as.numeric(r["BMIz"])
    perc = pnorm(z)
    group = cut(perc,c(-Inf,5,85,95,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)"),right = F)
  } else {group = NA}
  return(as.character(group))
})
df$BMI_group = as.factor(df$BMI_group)
df$BMI_group = ordered(df$BMI_group,c("Under Weight","Normal Weight","Overweight",
                                      "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
# PFTs
old <- c("Caucasian","Black","Other")
new <- c(1,2,5)
df$ppFEV = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FEV1)
df$ppFVC = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FVC)
```

```{r}
data = df
outcome = "HbA1c"
# Piecewise model function
piecewise = function(outcome,data = df,ordinal = F){
  # Add spline column
  data$Change = ifelse(data$Time < 0,0,1)
  data$Change = data$Change * data$Time
  # Model
  f = as.formula(paste0(outcome,"~Time + Change + (1|CFF.ID)"))
  mod = lmer(f,data = data)
  res = tidy(mod,"fixed")
  res$effect = NULL
  # Fit ordinal GEE (for example BMI group)
  if(ordinal){
    f = as.formula(paste0(outcome,"~Time + Change"))
    mod = ordgee(f,data = data,id = data$CFF.ID)
    res = summary(mod)
  }
  # Segmented fit
  fit = lme(as.formula(paste0(outcome,"~Time + Change")),random = ~1|CFF.ID,
            data = data,na.action = na.omit)
  seg = suppressWarnings(segmented(fit, seg.Z = ~Time, npsi=1))
  # Model check if necessary
  check = check_model(mod)
  # Coordinates for model plotting
  x1_start = -365
  y1_start = res$estimate[1] + res$estimate[2]*x1_start
  x1_end = 0
  y1_end = res$estimate[1] 
  x2_start = 0
  y2_start = res$estimate[1] 
  x2_end = 365
  y2_end = res$estimate[1] + (res$estimate[2]+res$estimate[3])*x2_end
  # Plot
  p = ggplot(data,aes_string(x = "Time",y = outcome,group = "CFF.ID",color = "Site")) +
    geom_point(alpha = 0.5) + geom_line(alpha = 0.2) +
    annotate("segment",x=x1_start,y=y1_start,xend=x1_end,yend=y1_end,color = "blue",size = 1) +
    annotate("segment",x=x2_start,y=y2_start,xend=x2_end,yend=y2_end,color = "blue",size = 1) +
    xlab("Days from Trikafta Start") + theme_bw()
  # Interactive plot and model results
  return(list(plot = ggplotly(p),table = res,check = check,seg = seg))
}
```

# Methods

For each outcome, we fit a piecewise linear mixed effects model with random intercept for each person and a "knot" at 0. This approach can be thought of as fitting two separate linear models, while ensuring that the two regression lines meet at Trikafta start. In the model results, "Slope Prior" indicates the slope prior to start of treatment (i.e. the average change in outcome for each "1 unit increase" in time). The "Change" term indicates the *change in slope* after medication start. To find the slope post-Trikafta, you add the "Slope Prior" and "Change" estimates.

The random intercept model was selected because it performed better (by AIC) than a model with random intercept and continuous AR1 correlation structure. Random slope models did not converge for some outcomes, so were not used for the sake of consistency between results.

# HbA1c

```{r results='asis'}
a1c = piecewise(outcome = "HbA1c",df)
ggplotly(a1c$plot)
kable(a1c$table,digits = 3)
a1c$check
```

`r if(class(a1c$seg)=="lme"){"Automated R package was not able to find a better breakpoint."}`

# BMI

Models did not converge with BMI converted to under weight, normal weight, etc. This is most likely due to lack of within-patient variability in terms of BMI group (i.e. people's BMIs aren't changing enough that they move from one group to another).

# Insulin

Most participants only have one observation of insulin dose, so these results would be the same as Maxie's previous reports.

# ppFEV

```{r results='asis'}
ppFEV = piecewise(outcome = "ppFEV",df)
ggplotly(ppFEV$plot)
kable(ppFEV$table,digits = 3)
ppFEV$check
```

# ppFVC

```{r results='asis'}
ppFVC = piecewise(outcome = "ppFVC",df)
ggplotly(ppFVC$plot)
kable(ppFVC$table,digits = 3)
ppFVC$check
```