### 2022-02-08 Christine Chan/CF Metabolomics/chan_cfrd_metabolomics.Rmd
Update chan_cfrd_metabolomics.Rmd
PLSDA permutation tests

---
title: ""
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(Hmisc)
library(mixOmics)
library(combinat)
library(tidyverse)
library(parallel)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD//PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
clean_metab = function(data,prefix = "M"){
  # Replace 1 with NA
  data[data == 1] = NA
  # Compound names
  data$compound_id = paste0(prefix,1:nrow(data))
  rownames(data) = data$compound_id
  compounds = data[,c("compound_id","Compound")]
  # Get samples only
  data = data[,grep("area2height",colnames(data))]
  data = data[,!grepl("blank",tolower(colnames(data)))]
  colnames(data) = gsub("Lipid_Pos_|Lipid_Neg_|Hilic_Pos_|_area2height.raw.|_dup","",colnames(data))
  # Transpose
  data = data.frame(t(data))
  # Fix subject names
  data$subject_id = str_pad(rownames(data),3,"left","0")
  # Return data and compounds
  return(list(data,compounds))
}
```

```{r}
# Clinical data
source("/Users/timvigers/Dropbox/Mac/Documents/GitHub/CF/Christine Chan/CF Metabolomics/import_data.r")
df = data %>% group_by(medical_record_number) %>% filter(row_number() == 1) %>% ungroup() %>%
  select(subject_id,group_category.factor,age_at_visit_1,gender.factor,race_ethnicity.factor,
         average_bmi_z_score,fvc,fev1_percent,poc_a1c,ag,fructosamine,
         percent_glycated_albumin)
df$subject_id = as.character(sub("--\\d","",df$subject_id))

# Import metabolomics
metab_pos = read.csv("./Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - C18 Pos.csv",na.strings = "")
metab_neg = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - C18 Neg.csv",na.strings = "")
hilic = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/CF Metabolomics/Data_Raw/Christine Chan - CF-Diabetes - HILIC Pos.csv",na.strings = "")
# Clean
## Positive
metab_pos = clean_metab(metab_pos,prefix = "P")
compounds = metab_pos[[2]]
## Negative
metab_neg = clean_metab(metab_neg,prefix = "N")
compounds = rbind(compounds,metab_neg[[2]])
## HILIC
hilic = clean_metab(hilic,prefix = "H")
compounds = rbind(compounds,hilic[[2]])
# Merge
df = df %>% filter(subject_id %in% metab_pos[[1]]$subject_id) %>%
  left_join(.,metab_pos[[1]],by = "subject_id") %>% 
  left_join(.,metab_neg[[1]],by = "subject_id") %>%
  left_join(.,hilic[[1]],by = "subject_id")
# Larger groups
df$combined = df$group_category.factor
levels(df$combined) = c("HC & CFNGT","HC & CFNGT","CFAGT & CFRD", "CFAGT & CFRD")
# Remove low variance
df = df[,-nearZeroVar(df)$Position]
# Matrices for mixOmics
X = df %>% select(P1:H651)
Y_all = df$group_category.factor
Y_two = df$combined
```

# PCA

## Four groups

```{r}
pca_res = pca(X,center = T,scale = T)
plotIndiv(pca_res,group = df$group_category.factor,legend = TRUE,pch = 16,
          title = "PCA Components 1 - 2")
```

## Control and CFNGT vs. Abnormal Glycemia

```{r}
plotIndiv(pca_res,group = df$combined,legend = TRUE,pch = 16,
          title = "PCA Components 1 - 2")
```

# PLS-DA

## Four groups

```{r}
plsda_res = splsda(X, Y_all)
plotIndiv(plsda_res,group = df$group_category.factor,legend = TRUE,pch = 16)
```

## Control and CFNGT vs. Abnormal Glycemia

```{r}
plsda_res = splsda(X, Y_two)
plotIndiv(plsda_res,group = df$combined,legend = TRUE,pch = 16)
```

### Permutation testing

```{r cache=TRUE}
real_auc_1 = auroc(plsda_res,print = F,plot = F)$Comp1[1]
real_auc_2 = auroc(plsda_res,print = F,plot = F)$Comp2[1]
perms = 1000
cl = makeCluster(8,type = "FORK")
aucs = parLapply(cl,1:perms, function(s){
  set.seed(s)
  Y_perm = sample(df$combined,length(df$combined))
  plsda_res = splsda(X, Y_perm)
  aucs = auroc(plsda_res,print = F,plot = F)
  comp1_auc = aucs$Comp1[1]
  comp2_auc = aucs$Comp2[1]
  return(c(comp1 = comp1_auc,comp2 = comp2_auc))
})
stopCluster(cl)
aucs = as.data.frame(do.call(rbind,aucs))
```

Two group outcome was "shuffled" `r perms` times and sPLSDA re-fit on the permuted outcome.

#### Component 1

```{r}
ggplot(aucs,aes(x = comp1)) + geom_histogram(binwidth = 0.02) + theme_bw() + 
  geom_vline(xintercept = real_auc_1,color = "red") +
  xlab("AUC") + ylab("Count")
```

Permutation p value for component 1 is `r round((1 + sum(aucs$comp1 > real_auc_1))/perms,3)`

#### Component 2

```{r}
ggplot(aucs,aes(x = comp2)) + geom_histogram(binwidth = 0.02) + theme_bw() + 
  geom_vline(xintercept = real_auc_2,color = "red") +
  xlab("AUC") + ylab("Count")
```

Permutation p value for component 1 is `r round((1 + sum(aucs$comp2 > real_auc_2))/perms,3)`