### 2022-04-20 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
correlations

---
title: "Pre-/Post-Trikafta (Prospective)"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(cgmanalysis)
library(AGD)
library(pracma)
library(corrplot)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
```

```{r}
# Import
df = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/trikafta_2022-02-8.csv",
              na.strings = c(""))
# Clean up
df$Sex = factor(df$Sex..Male.1.)
levels(df$Sex) = c("F","M")
df = df %>% group_by(Study.ID) %>% 
  mutate(trikafta_duration = sum(Time.on.trikafta.in.months)) %>% ungroup()
df[,c("Kalydeco","Orkambi","Symdeko")] = 
  lapply(df[,c("Kalydeco.1","orkambi.1","tezacaftor.ivacaftor..symdeko..1")],function(c){
    c = as.factor(c)
    levels(c) = c("No","Yes")
    return(c)
  })
# Genotype
df$Genotype = apply(df,1,function(r){
  if(tolower(r["CFTR.Mutation.1"]) == "f508del" & tolower(r["CFTR.Mutation.2"]) == "f508del"){
    g = "Homozygous F508del"
  } else if (tolower(r["CFTR.Mutation.1"]) == "f508del"){
    g = "Heterozygous F508del"
  } else {
    g = "Other"
  }
  return(g)
})
# BMI z score (count everyone over 20 as 20)
df$Age = ifelse(df$Age.at.visit. >= 20, 20, df$Age.at.visit.)
df$BMIz = y2z(df$BMI,x = df$Age,sex = df$Sex,ref=nl4.bmi)
# Weight z score
df$weightz = y2z(df$Weight..kg.,x = df$Age,sex = df$Sex,ref=nl4.wgt)
```

```{r cgm data,include=FALSE}
# Merge
cgm = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/cgm_vars.csv")
cgm$Study.ID = str_extract(cgm$subject_id,"^[[:upper:]]{2}\\d{4}")
cgm$Event.Name = 
  str_extract(tolower(cgm$subject_id),"visit\\d{1}|visit \\d{1}|yr\\d{1}|year\\d{1}|year \\d{1}")
cgm$Event.Name = paste("Visit",gsub("[[:alpha:]]|\\s","",cgm$Event.Name))
df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
keep_a1c = df %>% group_by(Study.ID) %>% 
  summarise(n = sum(!is.na(A1C.result))) %>%
  filter(n == 2) %>% pull(Study.ID)
# Update DXA variables
dxa_vars = c("Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
            "BMD.Total.z.score","BMD.Subtotal.z.score",
            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score")
dxa = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/dxa_data.csv")
dxa = dxa %>% select(Study.ID,all_of(dxa_vars))
df = left_join(df,dxa)
keep_dxa = dxa$Study.ID
```

# Table 1: Demographics at Baseline

For calculating BMI z scores for adults > 20, we used CDC references for 20 year olds.

```{r}
# Labels for all tables - not the prettiest code
newlabels = list(Age.at.visit. = "Age (years)",Sex..Male.1. = "Sex",
                 BMIz = "BMI z score",FEV1.. = "ppFEV1",FVC.... = "ppFVC",
                 trikafta_duration = "Duration on Trikafta (months) at Visit 2",
                 Pancreatic.Status = "Pancreatic Status")
newlabels["A1C.result"] = "HbA1c (%)"
newlabels["Weight..kg."] = "Weight (kg)"
newlabels["weightz"] = "Weight Z"
newlabels["Glucose.0.minute"] = "Fasting glucose (mg/dL)"
newlabels["Glucose.60.min"] = "1hr glucose (mg/dL)"
newlabels["Glucose.120.min"] = "2hr glucose (mg/dL)"
newlabels["c.peptide.0.minute"] = "Fasting C-peptide (pmol/L)"
newlabels["iAUC_ratio"] = "C pep iAUC:Glucose iAUC"
newlabels["homa2B"] = "HOMA2 %B"
newlabels["homa2S"] = "HOMA2 %S"
newlabels["homa2IR"] = "HOMA2 IR"
newlabels["mean_sensor"] = "Mean SG (mg/dL)"
newlabels["max_sensor"] = "Max. SG (mg/dL)"
newlabels["min_sensor"] = "Min SG (mg/dL)"
newlabels["percent_time_70_180"] = "% TIR 70-180 mg/dL"
newlabels["percent_time_over_140"] = "% Time >140 mg/dL"
newlabels["percent_time_over_180"] = "% Time >180 mg/dL"
newlabels["percent_time_over_250"] = "% Time >250 mg/dL"
newlabels["percent_time_under_70"] = "% Time <70 mg/dL"
newlabels["percent_time_under_54"] = "% Time <54 mg/dL"
newlabels["standard_deviation"] = "SD"
newlabels["cv"] = "CV"
newlabels["r_mage"] = "MAGE"
newlabels["trikafta_duration"] = "Modulator Duration"
newlabels[dxa_vars] = 
  c("Tissue (g)","Total Fat Mass (%)","Fat Body Mass (g)","Fat Free Mass (g)",
    "Lean Body Mass (g)","Total Mass (kg)","Fat Free Mass %","Fat Mass Index (FMI)",
    "Lean body mass/fat mass (LBM/FM)","Fat Free Mass Index (FFMI)","Lean Body Mass Index (LBMI)",
    "BMD Total z-score","BMD Subtotal z-score","DXA FMI calculated z-score",
    "DXA LBMI calculated z-score")
```

## Full cohort

```{r results='asis'}
t1 = tableby(~Age.at.visit.+Weight..kg.+weightz+trikafta_duration+
               Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep_a1c,])
summary(t1,digits = 2,labelTranslations = newlabels)
```

## Reduced Cohort

```{r results='asis'}
t1 = tableby(~Age.at.visit.+Weight..kg.+weightz+trikafta_duration+
               Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep_dxa,])
summary(t1,digits = 2,labelTranslations = newlabels)
```

# Table 2: Pre-/Post-Trikafta Comparisons

All metrics calculated based on 2 hour OGTT (0, 10, 30, 60, 90, and 120 minutes). Group was treated as an ordered variable with NGT < AGT < CFRD.

```{r}
# Make dataset for HOMA2 calculations - glucose in mmol/L and cpep in nmol/L
homa = df %>% select(Study.ID,Event.Name,Glucose.0.minute,c.peptide.0.minute)
homa$Glucose.0.minute = homa$Glucose.0.minute/18
homa$c.peptide.0.minute = homa$c.peptide.0.minute/1000*0.33
write.csv(homa,"./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/homa_data.csv",
          na = "",row.names = F)
# Import HOMA2 data (calculated with https://www.dtu.ox.ac.uk/homacalculator/download.php)
homa = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/homa_calculated.csv")
df = left_join(df,homa,by = c("Study.ID", "Event.Name"))
# Glucose tolerance group
glu_cols = c("Glucose.0.minute","Glucose.60.min","Glucose.120.min")
df$Group = apply(df[,glu_cols],1,function(r){
  if(r[1]<126 & r[2] < 200 & r[3] < 140){
    return("NGT")
  } else if (r[1] >= 126 | r[3] >= 200){
    return("CFRD")
  } else {
    return("AGT")
  }
})
df$Group = factor(df$Group,levels = c("NGT","AGT","CFRD"),ordered = T)
# convert Insulin to mcU/mL
ins_cols = colnames(df)[head(grep("insulin",tolower(colnames(df))),-1)]
df[,ins_cols] = lapply(df[,ins_cols], function(c){
  return(as.numeric(c) * 0.02869)
})
# cpeptide to pmol/L
cpep_cols = colnames(df)[head(grep("c.peptide",tolower(colnames(df))),-2)]
df[,cpep_cols] = lapply(df[,cpep_cols], function(c){
  return(c * 0.33)
})
# Calculate AUCs
df$`Glucose AUC` = apply(df %>% select(Glucose.0.minute:Glucose.120.min),1,function(r){
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
df$`C Peptide AUC` = apply(df %>% select(c.peptide.0.minute:c.peptide.120.min),1,function(r){
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
df$`Insulin AUC` = apply(df %>% select(Insulin.0.minute:Insulin.120.min),1,function(r){
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
# Incremental AUCs
df$`Glucose iAUC` = apply(df %>% select(Glucose.0.minute:Glucose.120.min),1,function(r){
  r = r - r[1]
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
df$`C Peptide iAUC` = apply(df %>% select(c.peptide.0.minute:c.peptide.120.min),1,function(r){
  r = r - r[1]
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
df$`Insulin iAUC` = apply(df %>% select(Insulin.0.minute:Insulin.120.min),1,function(r){
  r = r - r[1]
  trapz(c(0,10,30,60,90,120),as.numeric(r))
})
# Other indices
# Cpeptide index = delta cpeptide 0-30/delta glucose 0-30
df$`C Peptide Index` = (df$c.peptide.30.min - df$c.peptide.0.minute)/
  (df$Glucose.30.min - df$Glucose.0.minute)
# Matsuda c peptide (500,000/√[fasting glucose*fasting cpep*mean glucose*mean cpep])   [Glucose mg/dL, cpeptide pmol/L]
glu_cols = colnames(df)[head(grep("glucose\\.\\d{1,}",tolower(colnames(df))),-2)]
glu_cols = glu_cols[-2]
ins_cols = colnames(df)[head(grep("insulin\\.\\d{1,}",tolower(colnames(df))),-2)]
ins_cols = ins_cols[-1]
df$mean_glucose = rowMeans(df[,glu_cols])
df$mean_cpep = rowMeans(df[,cpep_cols])
df$`Adapted Matsuda` = 500000/(sqrt(df$Glucose.0.minute*df$c.peptide.0.minute*
                                      df$mean_glucose*df$mean_cpep))
df$Matsuda = apply(df,1,function(r){
  glu = round(as.numeric(r[glu_cols]))
  ins = round(as.numeric(r[ins_cols]),1)
  matsuda = 10000/sqrt(glu[1]*ins[1]*(glu[1]+glu[2]*2+glu[3]*2+glu[4]*2+glu[5])/
                         8*(ins[1]+ins[2]*2+ins[3]*2+ins[4]*2+ins[5])/8)
})
# C Pep ODI
df$`C Peptide ODI` = df$`C Peptide Index`*df$`Adapted Matsuda`
# Matsuda ODI
df$`ODI (Cpeptide index x Matsuda)` = df$`C Peptide Index`*df$Matsuda
# iAUC ratio
df$iAUC_ratio = df$`C Peptide iAUC`/df$`Glucose iAUC`
```

## Full cohort

```{r results='asis'}
t2_vars = c("A1C.result","Weight..kg.","weightz","BMIz","FEV1..","FVC....",
            "trikafta_duration","Group","Glucose.0.minute",
            "Glucose.60.min","Glucose.120.min","c.peptide.0.minute",
            "`Glucose iAUC`","`C Peptide iAUC`","`Insulin iAUC`",
            "`C Peptide Index`","Matsuda",
            "`Adapted Matsuda`","`C Peptide ODI`","`ODI (Cpeptide index x Matsuda)`",
            "iAUC_ratio","homa2B","homa2S","homa2IR",
            "Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
            "BMD.Total.z.score","BMD.Subtotal.z.score",
            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score",
            "mean_sensor","max_sensor","min_sensor","percent_time_70_180",
            "percent_time_over_140","percent_time_over_180","percent_time_over_250",
            "percent_time_under_70","percent_time_under_54","standard_deviation","cv","r_mage")
f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

## Reduced Cohort

```{r results='asis'}
f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
p = paired(f,data = df[df$Study.ID %in% keep_dxa,], 
           control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
           id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

# Table 3: Adults vs. Children

```{r}
df$`Age Group` = cut(df$Age.at.visit.,c(-Inf,18,Inf),labels = c("Child","Adult"))
```

The following table includes `r as.numeric(table(df[df$Event.Name == "Visit 1","Age Group"])["Child"])` children and `r as.numeric(table(df[df$Event.Name == "Visit 1","Age Group"])["Adult"])` adults.

```{r results='asis'}
f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
            strata = `Age Group`)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

# Table 4: Prior Modulator Use

Unable to perform statistical tests due to too many missing values.

# Correlations

White text indicates the correlation coefficient $R$, and results are only shown for correlations significant at the p < 0.05 level.

```{r fig.width=10,fig.height=10}
change_df = df %>% group_by(Study.ID) %>% 
  summarise_at(vars(Weight..kg.,c.peptide.0.minute,`C Peptide Index`,
                    `Adapted Matsuda`,`C Peptide ODI`,iAUC_ratio,homa2B:homa2IR),
               .funs = diff) %>%
  rename("Change in Weight (kg)" = Weight..kg.,
         "Change in Fasting C-Peptide (pmol/L)" = c.peptide.0.minute,
         "Change in C Peptide Index" = `C Peptide Index`,
         "Change in Adapted Matsuda" = `Adapted Matsuda`,
         "Change in C Peptide ODI" = `C Peptide ODI`,
         "Change in C pep iAUC:Glucose iAUC" = iAUC_ratio,
         "Change in HOMA2 %B" = homa2B, "Change in HOMA2 %S" = homa2S,
         "Change in HOMA2 IR" = homa2IR)

M = cor(change_df[,2:ncol(change_df)])
testRes = cor.mtest(change_df[,2:ncol(change_df)])
corrplot(M, p.mat = testRes$p, method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='white', number.cex = 0.8, order = 'AOE', diag=FALSE)
```

# Other Figures

## HbA1c

```{r}
ggplot(df,aes(x = Event.Name, y = A1C.result,fill = Event.Name)) + 
  geom_boxplot() + geom_jitter(width = 0.25) + 
  xlab("Visit") + ylab("HbA1c (%)") + theme_bw() + theme(legend.position = "none")
```

## Glucose

```{r}
plot_df = df %>% pivot_longer(all_of(glu_cols)) %>% select(Study.ID,Event.Name,name,value)
plot_df$name = as.numeric(gsub("Glucose|\\.|min|minute","",plot_df$name))
plot_df = plot_df %>% group_by(name,Event.Name) %>% 
  summarise(m = mean(value,na.rm = T),
            se = sd(value)/sqrt(n()),.groups = "drop")
ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
  geom_line() + geom_point() + 
  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
  xlab("Time (minutes)") + ylab("Glucose (mg/dL)") + ylim(0,200) + 
  theme_bw()
```

## Glucose

```{r}
plot_df = df %>% pivot_longer(all_of(cpep_cols)) %>% select(Study.ID,Event.Name,name,value)
plot_df$name = as.numeric(gsub("c.peptide|\\.|min|minute","",plot_df$name))
plot_df = plot_df %>% group_by(name,Event.Name) %>% 
  summarise(m = mean(value,na.rm = T),
            se = sd(value)/sqrt(n()),.groups = "drop")
ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
  geom_line() + geom_point() + 
  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
  xlab("Time (minutes)") + ylab("C Peptide (pmol/L)") + 
  theme_bw()
```

## Insulin

```{r}
plot_df = df %>% pivot_longer(all_of(ins_cols)) %>% select(Study.ID,Event.Name,name,value)
plot_df$name = as.numeric(gsub("Insulin|insulin|\\.|min|minute","",plot_df$name))
plot_df = plot_df %>% group_by(name,Event.Name) %>% 
  summarise(m = mean(value,na.rm = T),
            se = sd(value)/sqrt(n()),.groups = "drop")
ggplot(plot_df,aes(x = name,y = m,color = Event.Name)) + 
  geom_line() + geom_point() + 
  geom_errorbar(aes(ymin=m-(1.96*se), ymax=m+(1.96*se)), width=2) +
  xlab("Time (minutes)") + ylab("Insulin (mcU/mL)") + 
  theme_bw()
```

# Notes

1. The A1C numbers look correct to me. The p value in Table 2 is calculated for only those people with a measure at V1 and V2, so in this case it was calculated based on n = 15. Similarly, for "HbA1c (%) for those with CGM," there are 12 values missing out of 20, so the p value is for 8 participants.

2. The citation for the trapezoidal method R package is:

    Borchers, Hans W. Pracma: Practical Numerical Math Functions, 2022. https://CRAN.R-project.org/package=pracma.

3. Citation for R (version `r paste0(R.version$major,".","1.3")`):

    R Core Team. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing, 2022. https://www.R-project.org/.