### 2022-05-15 StatNet/knot_models.qmd
Create knot_models.qmd

---
title: "Piecewise Regression for Nutritional Outcomes Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
format: revealjs
editor: visual
bibliography: /Users/timvigers/Dropbox/Miscellaneous/references.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-alphabetical.csl
---

```{r}
#| include: false
# Load libraries, data, etc.
library(DiagrammeR)
library(MultiKink)
library(ggplot2)
data("triceps")
```

## Study Overview

-   Do vitamin levels change after CFTR modulator start?

-   Registry data through 2021:

    -   Most patients have multiple nutrition measures before and after CFTR modulator start.
    -   A subset of patients were on ivacaftor followed by elexacaftor/tezacaftor/ivacaftor (ETI).

## Study Overview

-   The first approach was to limit the analysis dataset to those with two measures prior to modulator start, and at least one measure after.

    -   Then we did a t test on the differences between measures closest to modulator start.

-   However, this results in a lot of excluded data and it was unclear how to treat those on two different modulators.

-   Piecewise models allow us to use all the data and include all the participants in the same model.

## Piecewise Linear Regression Basics

-   See [module and presentation](https://bookdown.org/tpinto_home/Beyond-Linearity) by Jaroslaw Harezlak & Armando Teixeira-Pinto for the Biostatistics Collaboration of Australia for additional detail.

-   Will use the "triceps" data included in the `MultiKink` R package [@wanMultiKinkEstimationInference2020] because it's a little easier to see the trends.

-   892 measurements of the triceps skin fold.

    -   Interested in modeling changes with age.

---

Triceps data:

```{r}
tri.age.plot = ggplot(triceps,aes(x = age,y = triceps)) + geom_point(alpha = 0.5) +
  xlab("Age") + ylab("Triceps") + theme_bw()
tri.age.plot
```

---

Add knots at ages 5, 10, 20, 30, and 40:

```{r}
pred1 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age<5,]))
pred2 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age >=5 & triceps$age<10,]))
pred3 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=10 & triceps$age<20,]))
pred4 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=20 & triceps$age<30,]))
pred5 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=30 & triceps$age<40,]))
pred6 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=40,]))
tri.age.plot + 
  geom_line(data=triceps[triceps$age<5,], 
            aes(y = pred1, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age >=5 & triceps$age<10,], 
            aes(y = pred2, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=10 & triceps$age<20,], 
            aes(y = pred3, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=20 & triceps$age<30,], 
            aes(y = pred4, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=30 & triceps$age<40,], 
            aes(y = pred5, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=40,], 
            aes(y = pred6, x=age), size = 1, col="blue") 
```

## Continuity restriction

-   We can adapt the previous model to make sure that the regressions "join up" at each knot:

$$
y = \beta_0 + \beta_1 x+\beta_2(x-k_1)_+\beta_3(x-k_2)_+ +...+ \epsilon
$$

with

$$
(x-k)_+=
\left\{ 
  \begin{array}{ c l }
    0, & \quad \textrm{if } x < k \\
    x-k,                 & \quad \textrm{if } x \geq k 
  \end{array}
\right.
$$

```{r}
#| eval: false
grViz("
digraph graph2 {
graph [layout = dot]
# node definitions with substituted label text
node [shape = rectangle]
a [label = '@@1']
b [label = '@@2']
c [label = '@@3']
d [label = '@@4']
e [label = '@@5']
f [label = '@@6']
g [label = '@@7']
h [label = '@@8']
i [label = '@@9']
j [label = '@@10']
k [label = '@@11']
l [label = '@@12']
m [label = '@@13']

a -> b
b -> c
c -> d
c -> e
c -> f
d -> g
e -> h
f -> i
g -> j
h -> k
i -> l
j -> m
k -> m
l -> m
}

[1]: paste0('Clinical Database (n = ', ncdb, ')')
[2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
[3]: paste0('In Our Dataset (n = ',length(intersect(mods$`Patient ID`,sweat$`Patient ID`)), ')')
[4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
[5]: paste0('Treated With Trikafta Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Trikafta']), ')')
[6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
[7]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
[8]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
[9]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
[10]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
[11]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
[12]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
[13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
")
```

# References