### 2022-05-18 StatNet/knot_models.qmd
Update knot_models.qmd

---
title: "Piecewise Regression for Nutritional Outcomes Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
format: revealjs
editor: visual
bibliography: /Users/timvigers/Dropbox/Miscellaneous/references.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-alphabetical.csl
cache: true
---

```{r}
#| include: false
# Load libraries, data, etc.
library(DiagrammeR)
library(MultiKink)
library(ggplot2)
library(broom)
library(lmerTest)
library(mcp)
library(knitr)
data("triceps")
cf = read.csv("/Users/timvigers/GitHub/CF/StatNet/data.csv",
              stringsAsFactors = F,na.strings = "")
```

## Study Overview

-   Do vitamin levels change after CFTR modulator start?

-   Registry data through 2021:

    -   Most patients have multiple nutrition measures before and after CFTR modulator start.
    -   A subset of patients were on ivacaftor followed by elexacaftor/tezacaftor/ivacaftor (ETI).

## Study Overview

-   The first approach was to limit the analysis dataset to those with two measures prior to modulator start, and at least one measure after.

    -   Then we did a t test on the differences between measures closest to modulator start.

-   However, this results in a lot of excluded data and it was unclear how to treat those on two different modulators.

-   Piecewise models allow us to use all the data and include all the participants in the same model.

## Piecewise Linear Regression Basics

-   See [module and presentation](https://bookdown.org/tpinto_home/Beyond-Linearity) by Jaroslaw Harezlak & Armando Teixeira-Pinto for the Biostatistics Collaboration of Australia for additional detail.

-   Will use the "triceps" data included in the `MultiKink` R package [@wanMultiKinkEstimationInference2020] because it's a little easier to see the trends.

-   892 measurements of the triceps skin fold.

    -   Interested in modeling changes with age.

------------------------------------------------------------------------

Triceps data:

```{r}
tri.age.plot = ggplot(triceps,aes(x = age,y = triceps)) + geom_point(alpha = 0.6) +
  xlab("Age") + ylab("Triceps") + theme_bw()
tri.age.plot
```

------------------------------------------------------------------------

Add knots at ages 5, 10, and 20:

```{r}
pred1 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age<5,]))
pred2 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age >=5 & triceps$age<10,]))
pred3 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=10 & triceps$age<20,]))
pred4 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=20,]))
tri.age.plot + 
  geom_line(data=triceps[triceps$age<5,], 
            aes(y = pred1, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age >=5 & triceps$age<10,], 
            aes(y = pred2, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=10 & triceps$age<20,], 
            aes(y = pred3, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=20,], 
            aes(y = pred4, x=age), size = 1, col="blue") 
```

## Continuity restriction

-   We can adapt the previous model to make sure that the regressions "join up" at each knot:

$$
y = \beta_0 + \beta_1 x+\beta_2(x-k_1)_+\beta_3(x-k_2)_+ +...+ \epsilon
$$

with

$$
(x-k)_+=
\left\{ 
  \begin{array}{ c l }
    0, & \quad \textrm{if } x < k \\
    x-k,                 & \quad \textrm{if } x \geq k 
  \end{array}
\right.
$$

## Manual coding

-   Can add these values by hand easily (using the difference \* an indicator):

```{r}
#| echo: true
# Make new variables
triceps$c5 = (triceps$age-5)*(triceps$age>=5)
triceps$c10 = (triceps$age-10)*(triceps$age>=10)
triceps$c20 = (triceps$age-20)*(triceps$age>=20)
```

-   Examples of what the new variables should look like:

```{r}
kable(triceps[c(which(triceps$age<5)[1],which(triceps$age<10&triceps$age>5)[1],
          which(triceps$age<20&triceps$age>10)[1],
          which(triceps$age>=20)[1]),c("age","c5","c10","c20")],row.names = F)
```

------------------------------------------------------------------------

Plot predicted values:

```{r}
# Linear model
mod = lm(triceps ~ age + c5 + c10 + c20,data = triceps)
# Plot
pred = predict(mod)
tri.age.plot +
  geom_line(data=triceps, 
            aes(y = pred, x=age), size = 1, col="blue") 
```

## Model interpretation

```{r}
kable(tidy(mod),digits = 3)
```

\
The estimates of c5, c10, and c20 indicate change in slope at each respective knot.

## Pre-/Post-CFTR Modulator Data

-   Vitamin D measures at annual visits from CF registry and our site.

-   Some patients started on ivacaftor who then transitioned to ETI.

## Pre-/Post-CFTR Modulator Data

```{r}
vita_plot = ggplot(cf,aes(x = Time,y = Vitamin.D,color = factor(Patient.ID))) + 
  geom_point() + geom_line(alpha = 0.2) +
  ylab("Vitamin D") + theme_bw() + theme(legend.position = "none")
vita_plot
```

## Pre-/Post-CFTR Modulator Model

-   The model is similar to triceps data example, but the second knot varies by patient and we include a random intercept to account for repeated measures:

$$
\text{Vitamin D}_{ij} = \beta_0 + \beta_1 \text{Time}_{ij} +\\ \beta_2(\text{Time}_{ij}-k_1)_+ + \beta_3(\text{Time}_{ij}-k_{2i})_+ + b_i + \epsilon_i
$$

## Pre-/Post-CFTR Modulator Model

with

$$
(x-k_1)_+=
\left\{ 
  \begin{array}{ c l }
    0, & \quad \textrm{if } \text{Time} < 0 \\
    x-k_1,                 & \quad \textrm{if } \text{Time} \geq 0 
  \end{array}
\right.
$$

and

$$
(x-k_2)_+=
\left\{ 
  \begin{array}{ c l }
    0, & \quad \textrm{if } \text{time before }2^{nd}\text{ mod}  \\
    \text{time after }2^{nd}\text{ mod},                 & \quad \textrm{if } \text{time after }2^{nd}\text{ mod}
  \end{array}
\right.
$$

and

$$
\epsilon \sim N(0,\sigma^2_{\epsilon})\text{ and } b\sim N(0,\sigma^2_b)
$$

## Pre-/Post-CFTR Modulator Results

```{r}
mod = lmer(Vitamin.D ~ Time + c1 + c2 + (1|Patient.ID),data = cf)
plot_df = mod@frame
plot_df$pred = predict(mod)
ggplot(plot_df,aes(x = Time,y = Vitamin.D,group = factor(Patient.ID))) +
    geom_line(aes(y = pred),alpha = 0.5) + geom_point(aes(y = pred),alpha = 0.5) +
    theme_bw() + theme(legend.position = "none") + xlab("Time (years)")
```

## Pre-/Post-CFTR Modulator Results

```{r}
kable(broom.mixed::tidy(mod,"fixed")[,-1],digits = 3)
```

## A Bayesian Approach

-   The `mcp` package provides an easy interface for fitting Bayesian regression models with multiple change points.
    -   Essentially it translates `lme4` syntax into JAGS.
-   Uninformative "t-tail" priors are used unless the user specifies something else.
    -   "The default priors should be suitable for estimation in the absence of strong prior knowledge."[@lindelovMcpPackageRegression2020a]

## On the Triceps Data

```{r}
#| echo: true
# Define the model
model = list(
  triceps ~ 1 + age,          # intercept + slope
  ~ 0 + age,  # joined slope 
  ~ 0 + age,  # joined slope 
  ~ 0 + age   # joined slope 
)
```

## On the Triceps Data

```{r}
#| include: false
bayes_fit = mcp(model, data = triceps, sample = "both")
s = summary(bayes_fit)
p = plot(bayes_fit)
```

```{r}
print(p)
```

-   Visually, this approach seems to generally confirm the knots we used earlier.

## On the Triceps Data

-   The model results bear this out as well:

```{r}
kable(s,digits = 3)
```

------------------------------------------------------------------------

## On the Triceps Data

-   `mcp` also allows us to test hypotheses about the location of these knots:

```{r}
#| echo: true
hypothesis(bayes_fit, "cp_1 = 5")
hypothesis(bayes_fit, "cp_2 = 10")
hypothesis(bayes_fit, "cp_3 = 20")
```

-   None of the hypothesis tests are significant, indicating that the knots found with `mcp` are reasonably close to the ones we pre-specified (the CI on cp_1 is fairly wide though).

## A Bayesian Model for Pre-/Post-CFTR Modulator

-   `mcp` can model joined slopes with varying change points:

```{r}
#| echo: true
# Define the model
model = list(
  Vitamin.D ~ 1 + Time, # intercept + slope
  ~ 0 + Time,  # joined slope
  1 + (1|Patient.ID) ~ 0 + Time
)
```

## A Bayesian Model for Pre-/Post-CFTR Modulator

```{r}
#| include: false
fit = mcp(model, data = cf)
s = summary(fit)
p = plot(fit)
```

```{r}
kable(s,digits = 3)
```

## A Bayesian Model for Pre-/Post-CFTR Modulator

```{r}
print(p)
```

-   This looks terrible!

## A Bayesian Model for Pre-/Post-First CFTR Modulator

-   Try fitting on just those with one modulator, with a N(0,1) prior on a change at 0:

```{r}
#| echo: true
# Define the model
model = list(
  Vitamin.D ~ 1 + Time, # intercept + slope
  ~ 0 + Time  # joined slope
)
priors = list(
  cp_1 = "dnorm(0,1)"
)
```

```{r}
#| include: false
fit = mcp(model,prior = priors,data = cf)
s = summary(fit)
p = plot(fit)
```

## A Bayesian Model for Pre-/Post-First CFTR Modulator

```{r}
kable(s,digits = 3)
```

## A Bayesian Model for Pre-/Post-First CFTR Modulator

```{r}
print(p)
```

## Limitations

-   Only 20 two-modulator participants with vitamin D values after starting ETI (21 measures among them).

    -   Large change in slope after second modulator start is likely driven by a few influential observations.

-   Somewhat arbitrarily limited pre-modulator values to \< 5 years from modulator start.

-   General data quality issues that we are in the process of resolving.

## Questions

-   How to handle truncated data (e.g. a value of "\< 0.1")?

-   Should I model the two-modulator and one-modulator groups separately, but still use piecewise models?

-   Any general advice (particularly for Bayesian methods)?

# References