### 2022-09-23 Edith Zemanick/CF Nutrition/cf_nutrition.qmd
trying to figure out liver labs in python

---
title: "CF Nutrition Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    self-contained: true
editor: visual
---

```{r include=FALSE}
library(readxl)
library(tidyverse)
library(parsedate)
library(arsenal)
```

```{r}
# Import all data
vit = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report Vitamins_throughDecember2021_revised20220712.xlsx") %>%
  select(`Patient ID`,Date,VitaminA_Retinol:`Vitamin E Gamma`)
fev = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_PFTs_ThroughDecember2021.xlsx") %>% 
  rename(`Patient ID` = PatientID) %>%
  select(`Patient ID`,Date,BMI:`FEF25-75 % pred`)
lft = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report_LFTs_ThroughDecember2021.xlsx") %>%
  select(`Patient ID`,Date,ALT:GGTP)
# Format dates
vit$Date = as.Date(parse_date(vit$Date,approx = F))
fev$Date = as.Date(parse_date(fev$Date,approx = F))
lft$Date = as.Date(parse_date(lft$Date,approx = F))
# Merge and sort
df = full_join(vit,fev,by = c("Patient ID", "Date"))
df = full_join(df,lft,by = c("Patient ID", "Date"))
df = df %>% arrange(`Patient ID`,Date)
# Conver to numeric
df[,3:ncol(df)] = lapply(df[,3:ncol(df)],as.numeric)
# Format modulator info
mods = read_excel("/Users/timvigers/Documents/Work/Vigers/CF/Edith Zemanick/CF Nutrition/Data_Raw/Report CFTR Modulators_throughDecember2021.xlsx")
mods$`Age at Start (years)` = NULL
mods = mods %>% filter(Modulator %in% c("Kalydeco","Trikafta"))
mods$`Earliest Date in Registry` = as.Date(parse_date(mods$`Earliest Date in Registry`,
                                                      approx = T))
mods$`Last Date in Registry` = as.Date(parse_date(mods$`Last Date in Registry`,
                                                      approx = T))
mods = mods %>% pivot_wider(id_cols = `Patient ID`,names_from = Modulator,
                     values_from = `Earliest Date in Registry`)
mods$`Effective Modulator Start` = pmin(mods$Trikafta,mods$Kalydeco,na.rm = T)
mods$fmt = mods$`Effective Modulator Start` == mods$Trikafta
mods$fmk = mods$`Effective Modulator Start` == mods$Kalydeco
mods[,c("fmt","fmk")][is.na(mods[,c("fmt","fmk")])]=0
mods$`First Modulator` = c("Trikafta","Kalydeco")[max.col(mods[,c("fmt","fmk")])]
mods[,c("fmt","fmk")] = NULL
# Put together
df = df %>% filter(`Patient ID` %in% mods$`Patient ID`)
df = left_join(df,mods,by = "Patient ID")
df$`Pre or Post` = factor(df$Date <= df$`Effective Modulator Start`,
                          levels = c("TRUE","FALSE"),labels = c("Pre","Post"))
# Filter and sort again, just in case
df$`Days from Start` = as.numeric(difftime(df$Date,df$`Effective Modulator Start`,units = "days"))
df = df %>% filter(`Days from Start` >= -365*3,
                   `Days from Start` < 0 | `Days from Start` >=90) %>% 
  arrange(`Patient ID`,Date)
# Average pre-modulator values
pre = df %>% group_by(`Patient ID`,`Pre or Post`) %>% summarise_at(vars(BMI:GGTP),mean, na.rm = T) %>% filter(`Pre or Post` == "Pre")
# Find most recent vitamin measure, get any other measures within 30 days
post = df[df$`Pre or Post`=="Post",]
```

# Data Cleaning

-   All vitamin, PFT, and BMI percentile values were converted to numeric, so values such as "LESS THAN 0.4" were marked as missing.

-   We excluded visits more than 3 years prior to effective modulator start and within 90 after effective modulator start.

-   All values prior to effective modulator start were averaged to calculate the "pre" value. We selected the most recent vitamin values as the "post" value, and included any BMI percentile, FEV1 % predicted, ALT, AST, and GGTP measures within a month of that visit.

-   Sex was only included in the PFT datasheet, so those without PFT data were missing this variable.

# Baseline value grouped by first modulator

```{r}
t1 = tableby(First.Modulator~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df[df$Timepoint == "Pre",])
summary(t1,pfootnote = T)
```

# Pre/post comparison

```{r results='asis'}
p = paired(Timepoint~Vitamin.A+Vitamin.D+Vitamin.E.Alpha+
           ppFEV1+BMI.Percentile+ALT+AST+GGTP,data = df,id=ID,control = list(numeric.test = "signed.rank"))
summary(p,pfootnote = T)
```