### 2022-10-20 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
added tables by specific genotype

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(readxl)
library(arsenal)
library(tidyverse)
library(pROC)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test"
  source("C:/Users/timvigers/Documents/GitHub/CF/get_genotype_class_severity.R")
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
  source("~/GitHub/CF/get_genotype_class_severity.R")
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
  source("~/GitHub/CF/get_genotype_class_severity.R")
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
```

```{r}
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Numeric variables
num_vars = c("Amount","Na","K","Cl")
num_vars = c(paste0(num_vars,"_Left"),paste0(num_vars,"_Right"))
df[,num_vars] = suppressWarnings(lapply(df[,num_vars],as.numeric))
# Filter per Maxie's code
df <- df %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
  filter(
    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
    !(Diagnosis == "Non-CF" & PatientID %in%
      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)),
    Modulator == "None"
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE")
# Genotype info
## By severity
df$geno1 <- get_cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- get_cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "_")
df$geno_risk <- factor(df$geno_risk,
  levels = c(
    "Severe_Severe","Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild",
    "Not CF_Not CF","Severe_Unknown", "Unknown_Severe","Unknown_Unknown"
  ),
  labels = c(
    "High","Low", "Low", "Low", "Low", "Not CF",  "Unknown",
    "Unknown", "Unknown"
  )
)
## By function
df$geno1_class <- get_cf_genotype_class_severity(df$Genotypes1,out = "class")
df$geno1_class[grep("508",df$Genotypes1)] = "F508del"
df$geno1_class = replace(df$geno1_class,df$geno1_class %in% c("I","II","III"),"MF")
df$geno1_class = replace(df$geno1_class,df$geno1_class %in% c("IV","V"),"RF")

df$geno2_class <- get_cf_genotype_class_severity(df$Genotypes2,out = "class")
df$geno2_class[grep("508",df$Genotypes2)] = "F508del"
df$geno2_class = replace(df$geno2_class,df$geno2_class %in% c("I","II","III"),"MF")
df$geno2_class = replace(df$geno2_class,df$geno2_class %in% c("IV","V"),"RF")

df$geno_class = factor(paste0(pmin(df$geno1_class,df$geno2_class),"/",
                       pmax(df$geno1_class,df$geno2_class)))
# Which arm?
df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
  k <- sum(!is.na(r))
  if (k == 6) {
    return("Both")
  } else if (k == 3) {
    return("Arm Not Noted")
  } else {
    return(NA)
  }
})

# Difference between arms
df$`Cl Diff. Between Left and Right Arm` = as.numeric(df$Cl_Left) - as.numeric(df$Cl_Right)
df$`Na Diff. Between Left and Right Arm` = as.numeric(df$Na_Left) - as.numeric(df$Na_Right)
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sodium to chloride ratio
df$`Na/Cl Ratio` = df$Na /df$Cl
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
# Table options (most variables are not normally distributed)
mycontrols  <- tableby.control(numeric.test="kwt",numeric.stats=c("Nmiss", "median", "q1q3"))
labels = list(Age.at.Test.in.Years = "Age (Years)", 
              geno_risk = "Genotype Severity",geno_class = "Specific Genotype")
```

# Patient Demographics and Sample Characteristics

For those with two sweat electrolyte measures (left and right), the average of the two was used. Each participants' first sweat test value was used for the following tables.

## By diagnosis

```{r results='asis'}
t1_df <- df %>%
  group_by(PatientID) %>%
  filter(row_number() == 1) # First measure for everyone
t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + fe(geno_risk) + fe(geno_class) + 
                Amount + Cl + Na + K + `Na/Cl Ratio`,
  data = t1_df,control = mycontrols,simulate.p.value = T
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

## By method

```{r results='asis'}
t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + fe(geno_risk) + fe(geno_class) +
                Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + 
                `Na Diff. Between Left and Right Arm`,
  data = t1_df,control = mycontrols,simulate.p.value = T
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

# Electrolyte comparisons

## By diagnosis

```{r results='asis'}
t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
                `Cl Diff. Between Left and Right Arm` + 
                `Na Diff. Between Left and Right Arm`,
  data = t1_df,control = mycontrols,
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### Gibson-Cooke only

```{r results='asis'}
t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
                `Cl Diff. Between Left and Right Arm` + 
                `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Method == "Gibson-Cooke",],control = mycontrols,
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### Macroduct only

```{r results='asis'}
t1 <- tableby(Diagnosis ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
                `Cl Diff. Between Left and Right Arm` + 
                `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Method == "Macroduct",],control = mycontrols,
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

## By genotype risk

```{r results='asis'}
t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
                `Cl Diff. Between Left and Right Arm` + 
                `Na Diff. Between Left and Right Arm`,
  data = t1_df,control = mycontrols,
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

## By genotype risk and diagnosis

### Cystic fibrosis

```{r results='asis'}
t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### CFTR-related disorder

```{r results='asis'}
t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### CRMS

```{r results='asis'}
t1 <- tableby(geno_risk ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "CRMS",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### Non-CF

```{r results='asis'}
t1 <- tableby( ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "Non-CF",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### Plots

```{r warning=FALSE}
ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Sweat Sodium")

ggplot(df, aes(x = Diagnosis, y = K)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Sweat Potassium")

ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Sweat Chloride")

ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Na/Cl Ratio")
```

## By specific genotype

```{r results='asis'}
t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + 
                `Cl Diff. Between Left and Right Arm` + 
                `Na Diff. Between Left and Right Arm`,
  data = t1_df,control = mycontrols,
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

## By specific genotype and diagnosis

### Cystic fibrosis

```{r results='asis'}
t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "Cystic Fibrosis",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### CFTR-related disorder

```{r results='asis'}
t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "CFTR-related disorder",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### CRMS

```{r results='asis'}
t1 <- tableby(geno_class ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "CRMS",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### Non-CF

```{r results='asis'}
t1 <- tableby( ~ Amount + Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,
  data = t1_df[t1_df$Diagnosis == "Non-CF",],control = mycontrols
)
summary(t1,pfootnote = T,
  labelTranslations = labels
)
```

### Plots

```{r warning=FALSE}
ggplot(df, aes(x = geno_class, y = Na)) +
  geom_boxplot() +
  theme_bw() + xlab("Specific Genotype") +
  ggtitle("Sweat Sodium")

ggplot(df, aes(x = geno_class, y = K)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Sweat Potassium")

ggplot(df, aes(x = geno_class, y = Cl)) +
  geom_boxplot() +
  theme_bw() + xlab("Specific Genotype") +
  ggtitle("Sweat Chloride")

ggplot(df, aes(x = geno_class, y = `Na/Cl Ratio`)) +
  geom_boxplot() +
  theme_bw() + xlab("Specific Genotype") +
  ggtitle("Na/Cl Ratio")
```

# Between occasion

For all participants with more than one sweat test, we compared their first and last values.

```{r results='asis'}
# Get first and last values for everyone with more than 1
multiples = df %>% group_by(PatientID) %>% 
  filter(n() > 1) %>% # Multiple values only
  filter(row_number()==1 | row_number()==n()) %>% # First and last
  mutate(time = row_number())
multiples$time = factor(multiples$time,levels = 1:2,labels = c("First","Last"))
# Paired testing options
mycontrols  <- tableby.control(numeric.test="signed.rank",cat.test = "mcnemar",
                               numeric.stats=c("Nmiss", "median", "q1q3"))
```

## Overall

```{r results='asis'}
p <- paired(time ~ Diagnosis + Cl + Na + K + `Na/Cl Ratio` + 
              `Cl Diff. Between Left and Right Arm` + 
              `Na Diff. Between Left and Right Arm`,
            signed.rank.exact = F,data = multiples, 
            id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

## By diagnosis

Not enough observations in CRMS or CFTR-related disorder groups to calculate p values. Also not enough observations to further break down the table by sweat collection method.

```{r results='asis'}
p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,strata = Diagnosis,
            data = multiples %>% filter(Diagnosis %in% c("Cystic Fibrosis","Non-CF")), 
            id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

## By genotype risk

```{r results='asis'}
p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + `Cl Diff. Between Left and Right Arm` + `Na Diff. Between Left and Right Arm`,signed.rank.exact = F,strata = geno_risk,
            data = multiples, 
            id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

## By genotype risk and diagnosis

Not enough paired observations in the CRMS or CFTR-related disorder groups.

### Cystic fibrosis

```{r results='asis'}
p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_risk,
            data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"), 
            id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

### Non-CF

```{r results='asis'}
p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_risk,
            data = multiples %>% filter(Diagnosis == "Non-CF"), 
            id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

## By specific genotype

```{r results='asis'}
p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio` + 
              `Cl Diff. Between Left and Right Arm` + 
              `Na Diff. Between Left and Right Arm`,
            signed.rank.exact = F,strata = geno_class,numeric.test = "notest",
            data = multiples, id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

## By genotype class and diagnosis

Not enough paired observations in the CRMS or CFTR-related disorder groups.

### Cystic fibrosis

```{r results='asis'}
p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,
            signed.rank.exact = F,strata = geno_class,
            data = multiples %>% filter(Diagnosis == "Cystic Fibrosis"), 
            id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

### Non-CF

```{r results='asis'}
p <- paired(time ~ Cl + Na + K + `Na/Cl Ratio`,signed.rank.exact = F,strata = geno_class,
            data = multiples %>% filter(Diagnosis == "Non-CF"), 
            id = PatientID,control = mycontrols)
summary(p,pfootnote = T,labelTranslations = labels)
```

# Chloride groups

We performed receiver operating characteristic (ROC) analysis to find a sweat sodium cutoff that distinguishes between chloride groups. Only the first sweat test from each participant was used. Because the groups are based on chloride value, Na/Cl ratio is correlated with the outcome so Na alone was used instead. Also, ROC analyses generally assume a binary outcome (i.e. looking for a cutoff to distinguish between two groups), so

In the following results, "Low" indicates a sweat chloride under 30, "Medium" indicates a sweat chloride of 30 - 59, and "High" indicates a sweat chloride greater than or equal to 60.

```{r}
# Three groups
t1_df$Group = cut(t1_df$Cl,breaks = c(0,30,60,Inf),right = F,labels = c("Low","Medium","High"))
# Low vs. medium or high
t1_df$Low = t1_df$Group
levels(t1_df$Low) = c("Low","Medium or High","Medium or High")
t1_df$Low = relevel(t1_df$Low,ref = "Medium or High")
# High vs. medium or low
t1_df$High = t1_df$Group
levels(t1_df$High) = c("Low or Medium","Low or Medium","High")
t1_df$High = relevel(t1_df$High,ref = "Low or Medium")
```

## ROC analysis

```{r}
roc_low <- roc(response = t1_df$Low,predictor = t1_df$Na,quiet = T)
coords_low = round(coords(roc_low,"best",best.method="youden"),3)
roc_high <- roc(response = t1_df$High,predictor = t1_df$Na,quiet = T)
coords_high = round(coords(roc_high,"best",best.method="youden"),3)

plot(roc_low,print.auc = TRUE, col = "blue") 
plot(roc_high, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between low vs. medium or high, and the green is the ROC curve for distinguishing between low or medium vs. high. The optimal threshold for distinguishing between low vs. medium or high was `r coords_low$threshold` (sensitivity = `r coords_low$sensitivity`, specificity = `r coords_low$specificity`) based on the Youden index. The optimal threshold for distinguishing between high vs. low or medium was `r coords_high$threshold` (sensitivity = `r coords_high$sensitivity`, specificity = `r coords_high$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_low$auc,roc_high$auc)),3)`.

## Histogram by group

Vertical red lines indicate the thresholds identified above.

```{r warning=FALSE}
ggplot(t1_df[!is.na(t1_df$Group),],aes(Na,fill = Group)) + 
  geom_histogram(alpha=0.3, position="identity",binwidth = 1) + 
  geom_vline(xintercept = coords_low$threshold,color = "red") +
  geom_vline(xintercept = coords_high$threshold,color = "red") +
  ylab("Count") + theme_bw()
```