### 2022-12-11 Christine Chan/CF Registry/cf_registry_analysis.Rmd
registry mostly done

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(knitr)
library(parallel)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/home/tim/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data")
```

```{r eval=FALSE}
# Read in CSVs
df <- read.csv("CFF19_encountersMerged_Del1.csv")
annual <- read.csv("CFF19_Annualized_Del1.csv")
colnames(annual)[colnames(annual) == "ReviewYear"] <- "reviewyear"
demo <- read.csv("CFF19_DemogCFDiag_Del1.csv")
# Combine and format demographics columns
demo$Race <- apply(demo, 1, function(r) {
  race <- r[paste0("Race", 1:6)]
  race <- as.numeric(which(race == 1))
  race <- paste(race, collapse = ", ")
})
demo$Race[grep(", ", demo$Race)] <- "More Than One"
demo$Race <- factor(demo$Race,
                    levels = c(1:6, "More Than One"),
                    labels = c(
                      "White", "Black or African American",
                      "American Indian or Alaska Native", "Asian",
                      "Native Hawaiian or Other Pacific Islander",
                      "Some other race", "More Than One"
                    )
)
demo$Hispanicrace <- factor(demo$Hispanicrace,
                            levels = 1:2,
                            labels = c("Yes", "No")
)
# Merge
df <- left_join(df, annual, by = c("eDWID", "reviewyear"))
df <- left_join(df, demo, by = "eDWID")
# Group years by 2003-2009, 2010-2013, and 2014-2018
df$Period <- cut(df$reviewyear,
                 breaks = c(2002, 2009, 2013, 2018),
                 labels = c("2003 - 2009", "2010 - 2013", "2014 - 2018")
)
df$smoking <- factor(df$smoking, levels = 1:4, labels = c(
  "No",
  "Occasionally",
  "Yes, Regularly, less than 1 ppd",
  "Yes, Regularly, 1 ppd or more"
), ordered = T)
df$second_smoke <- factor(df$second_smoke, levels = c(1:4), labels = c(
  "No",
  "Occasionally",
  "Yes, Regularly, less than 1 ppd",
  "Yes, Regularly, 1 ppd or more"
), ordered = T)
# Factors
ed_levels <- c(
  "Less than High School",
  "High School diploma or equivalent",
  "Some College",
  "College Graduate",
  "Masters/Doctoral level degree"
)
df$patient_education <- factor(df$patient_education,
                               levels = c(1:5),
                               labels = ed_levels
)
df$father_education <- factor(df$father_education,
                              levels = c(1:5),
                              labels = ed_levels
)
df$mother_education <- factor(df$mother_education,
                              levels = c(1:5),
                              labels = ed_levels
)
df$spouse_education <- factor(df$spouse_education,
                              levels = c(1:5),
                              labels = ed_levels
)
df$pregnant <- factor(df$pregnant, levels = 0:1, labels = c("No", "Yes"), ordered = T)
# Combine education checkboxes
df$employment <- apply(df, 1, function(r) {
  emp <- r[paste0("employment", 1:7)]
  emp <- as.numeric(emp)
  if (all(is.na(emp) | emp == 0)) {
    return(NA)
  } else {
    emp <- as.numeric(emp)
    emp <- which(emp == 1)
    emp <- paste(emp, collapse = ", ")
    return(emp)
  }
})
df$employment <- sub("1", "Part Time", df$employment)
df$employment <- sub("2", "Full time homemaker", df$employment)
df$employment <- sub("3", "Full time employment", df$employment)
df$employment <- sub("4", "Unemployed", df$employment)
df$employment <- sub("5", "Student", df$employment)
df$employment <- sub("6", "Disabled", df$employment)
df$employment <- sub("7", "Retired", df$employment)
df$employment <- as.factor(df$employment)
# Bronchodilator use
df$bronchodilators <- apply(df, 1, function(r) {
  bronch <- r[grep("bronchodilators",names(r))]
  names(bronch) = c("Beta agonist",
                    "Theophylline product",
                    "Short acting beta agonist",
                    "Long acting beta agonist",
                    "Short acting anticholinergic",
                    "Long acting anticholinergic",
                    "Combination beta agonist and anticholinergic"
  )
  if (all(is.na(bronch))) {
    return(NA)
  } else {
    bronch <- names(bronch)[which(bronch == 1)]
    bronch <- paste(bronch, collapse = ", ")
    return(bronch)
  }
})
# Exacerbation medication
df$act_primary <- apply(df, 1, function(r) {
  act <- r[grep("A_act_primary",names(r))]
  names(act) = c("Positive Expiratory Pressure (PEP)",
                 "Postural drainage with clapping (CPT)",
                 "Forced expiratory techniques",
                 "Oscillating PEP",
                 "High frequency chest wall oscillation",
                 "Exercise",
                 "None",
                 "Other"
  )
  if (all(is.na(act) | act == 0)) {
    return(NA)
  } else {
    act <- names(act)[which(act == 1)]
    act <- paste(act, collapse = ", ")
    return(act)
  }
})
# Oxygen therapy
df$oxygen_therapy <- factor(df$oxygen_therapy,levels = 1:5,
                            labels = c("Yes, Continuously",
                                       "Yes, Nocturnal and/or with exertion",
                                       "Yes, During exacerbation",
                                       "Yes, prn",
                                       "No"))
# Save formatted data for re-import later
save(df,file = "Data_Clean/formatted_data.Rdata")
```

```{r}
load("Data_Clean/formatted_data.Rdata")
df = df %>% arrange(eDWID,reviewyear,encounternum)
```

# Table 1: Participant characteristics at last visit in time period

```{r warning=FALSE,results='asis'}
# Summarize by ID and period
t1_df <- df %>%
  group_by(eDWID, Period) %>%
  summarise(
    Age = max(Age_YrEnd), Race = last(na.omit(Race)),
    Hispanic = last(na.omit(Hispanicrace)), Gender = last(na.omit(Gender)),
    `Patient Education` = last(na.omit(patient_education)),
    `Father Education` = last(na.omit(father_education)),
    `Mother Education` = last(na.omit(mother_education)),
    `Spouse Education` = last(na.omit(spouse_education)),
    `Marital Status` = last(na.omit(marital_status)),
    Employment = last(na.omit(employment)),
    Pregnancy = max(pregnant, na.rm = T)
  )
# Replace infinite values with missing
t1_df <- data.frame(lapply(t1_df, function(c) {
  c[is.infinite(c)] <- NA
  return(c)
}))
# Print table 1
cols <- colnames(t1_df)[3:ncol(t1_df)]
t1 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
  data = t1_df
)
summary(t1)
```

# Table 2: Pulmonary function

Continuous and count variables are from the last year of each period.

```{r warning=FALSE,results='asis'}
# Summarize by ID and period
t2_df <- df %>%
  group_by(eDWID, Period) %>%
  summarise(
    Smoking = max(smoking,na.rm = T),
    `Secondhand smoke`=max(second_smoke,na.rm = T),
    `FEV1 % Predicted` = last(na.omit(GLI_FEV1_pct_predicted)),
    `FVC % Predicted` = last(na.omit(GLI_FVC_pct_predicted)),
    `Pulm. Exacerbations` = last(na.omit(NumPulmExacerbation)),
    .groups = "keep"
  )
# Replace infinite values with missing
# t2_df <- data.frame(lapply(t2_df, function(c) {
#   c[is.infinite(c)] <- NA
#   return(c)
# }))
# Print table 1
cols <- colnames(t2_df)[3:ncol(t2_df)]
t2 <- tableby(as.formula(paste0("Period~", paste0("`", cols, "`", collapse = "+"))),
  data = t2_df
)
summary(t2)
```

# Questions

1. Which variables do we include for microbiome stuff? Include "othermicroorganismstypes"?