### 2022-12-15 Christine Chan/CF Registry/cf_registry_analysis.Rmd
starting over CF registry

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: no
    page-layout: full
---

```{r setup, include=FALSE}
library(table1)
knitr::opts_chunk$set(echo = FALSE,cache = TRUE,cache.lazy = FALSE)
knitr::opts_knit$set(root.dir = "~/Work/CHCO/Christine Chan/CF registry data")
```

```{r}

```