### 2023-01-10 Emily Muther/mental_health_screening.qmd
Update mental_health_screening.qmd

---
title: "Mental Health Screening in CF"
author: "Tim Vigers"
date: last-modified
format:
  html:
    code-fold: true
    self-contained: true
    page-layout: full
    title-block-banner: true
editor: source
---

# Data Cleaning

```{python}
import pandas as pd
# Data import
gad7 = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/GAD-7_General_Anxiety_Disorder_Full_Data_data_11-15-22.csv")
gad7 = gad7[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','GAD-7 Total Score','GAD-7 Totals']]
gad7.dropna(axis=0,inplace=True)
gad7.drop_duplicates(inplace=True)
phq9 = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PHQ-9_Full_Data_data_11-11-22.csv")
phq9 = phq9[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','PHQ-9 Total Score (calculated)','PHQ-9 Totals']]
phq9.dropna(axis=0,inplace=True)
phq9.drop_duplicates(inplace=True)
psc = pd.read_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv")
psc = psc[['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT ','Score','PSC Totals']]
psc.dropna(axis=0,inplace=True)
psc.drop_duplicates(inplace=True)
# Merge
df = gad7.merge(phq9,on=['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT '],how="outer")
df = df.merge(psc,on=['PAT_MRN_ID','CONTACT_DATE','AGE AT CONTACT '],how="outer")
# Rename
df.rename({'PAT_MRN_ID':'ID','CONTACT_DATE':'Date','AGE AT CONTACT ':'Age','GAD-7 Total Score':'GAD-7 Total','GAD-7 Totals':'GAD-7 Group','PHQ-9 Total Score (calculated)':'PHQ-9 Total','PHQ-9 Totals':'PHQ-9 Group','Score':'PSC-17 Total','PSC Totals':'PSC-17 Group'},inplace=True,axis=1)
# Write to CSV
df.to_csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/cf_mental_health_screening.csv")
```

- On first import, there are 2,646 observations each for GAD-7, PHQ-9, and PSC-17. 
- After deleting rows missing any any of patient ID, contact date, age at contact, total score, or score group, there were: 
  - 559 observations for GAD-7
  - 548 for PHQ-9
  - 308 for PSC-17
- After dropping duplicate rows, there are: 
  - 390 GAD-7 observations
  - 383 for PHQ-9
  - 216 for PSC-17
- After merging these together based on ID, date, and age, there are 613 unique observations with at least one of the three screening measures.

# Profiling

```{python}
from pandas_profiling import ProfileReport
profile = ProfileReport(df, title="Data Profiling Report")
profile.to_notebook_iframe()
```