### 2023-02-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
separate cleaning and analysis files Colleen Wood

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(arsenal)
load("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_dataset.rdata")
```

# Participant characteristics prior to CFTR modulator start

```{r results='asis'}
# Table
t1 <- tableby(~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
  `G-tube in past 12 months` + `Pancreatic Status` + Site +
  Genotype + `Age Group` + Race,
data = df %>% filter(`Pre or Post` == "Pre")
)
summary(t1)
```

# Pre-post

## HbA1c and others

### All groups

```{r results='asis'}
p <- paired(`Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
  ppFEV1 + ppFVC + TDD + `Long acting insulin dose`,
data = df, id = `CFF ID`,
signed.rank.exact = FALSE, 
numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

### By site

```{r results='asis'}
p <- paired(`Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
  ppFEV1 + ppFVC + TDD,
data = df, id = `CFF ID`,strata = Site,
signed.rank.exact = FALSE, numeric.test = "notest",
numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

## CGM

Only 13 of those with CGM also had HbA1c values pre- and post-CFTR modulator.

### Demographics

```{r results='asis'}
cgm = df[!is.na(df$average_sensor),]
# Table
t1 <- tableby(~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFVC +
  `G-tube in past 12 months` + `Pancreatic Status` + Site +
  Genotype + `Age Group` + Race,
data = cgm %>% filter(`Pre or Post` == "Pre")
)
summary(t1)
```

### Overall CGM variables

```{r results='asis'}
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
cgm_formula = update(cgm_formula,.~.+TDD + `Long acting insulin dose`)
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

### Daytime CGM variables

```{r results='asis'}
cgm$daytime_cv = cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
day_cgm = c(day_cgm,"daytime_cv")
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

### Nighttime CGM variables

```{r results='asis'}
cgm$nighttime_cv = cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
night_cgm = c(night_cgm,"nighttime_cv")
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```