### 2023-03-29 Scott Sagel/PEx in CF/nacfc_2023.qmd
Update nacfc_2023.qmd
tidyverse styling

---
title: "The Changing Face of Pediatric Pulmonary Exacerbations in CF (NACFC 2023)"
author: "Tim Vigers"
date: "today"
format:
  html:
    toc: false
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
execute: 
  echo: false
---

```{r}
#| include: false
library(readxl)
library(tidyverse)
library(gt)
library(gtsummary)
source("~/GitHub/CF/genotype_class_severity.R")
# Import
chco <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx", na = c("", "NULL"))
cnh <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
# Format CHCO columns
chco$`P. aeruginosa` <- coalesce(chco$`P. aeruginosa mucoid`, chco$`P. aeruginosa non-mucoid`)
chco <- chco %>%
  rename(
    Race = `Race (White/Black/Other)`, Ethnicity = `Hispanic/Latinx`
  ) %>%
  mutate(
    Genotypes1 = case_when(
      Genotypes1 == "F508del" | Genotypes1 == "delF508" ~ "F508del",
      .default = "Other"
    ),
    Genotypes2 = case_when(
      Genotypes2 == "F508del" | Genotypes2 == "delF508" ~ "F508del",
      .default = "Other"
    ),
    `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))
  ) %>%
  unite(Genotype, Genotypes1, Genotypes2, sep = "/")
# Format CNH to match CHCO
cnh <- cnh %>%
  rename(
    `# Days in Admit` = `Total hospitalization days`,
    `Age at Admit` = `Age at admission`,
    `FEV1 % Predicted at Discharge` = `End of treamtent ppFEV1`,
    `Unique number` = `Study ID`
  ) %>%
  mutate(
    Site = "CNH",
    `Modulator at Time of Admit` = case_when(
      None == "Unchecked" ~ "Yes",
      None == "Checked" ~ "No"
    ),
    MRSA = case_when(
      MRSA == "Unchecked" ~ NA,
      MRSA == "Checked" ~ TRUE
    ),
    `Staph aureus` = case_when(
      MSSA == "Unchecked" ~ NA,
      MSSA == "Checked" ~ TRUE
    ),
    `P. aeruginosa` = case_when(
      `P. aeruginosa` == "Unchecked" ~ NA,
      `P. aeruginosa` == "Checked" ~ TRUE
    ),
    Genotype = case_when(
      `CFTR genotype, groups` == "F508del heterozygous" ~ "F508del/Other",
      `CFTR genotype, groups` == "F508del homozygous" ~ "F508del/F508del",
      `CFTR genotype, groups` == "Other" ~ "Other/Other"
    ),
    `Specific Modulator at Admit` = case_when(
      `Ivacaftor (Kalydeco)` == "Checked" ~ "Other",
      `Ivacaftor/lumacaftor (Orkambi)` == "Checked" ~ "Other",
      `Tezacaftor/ivacaftor (Symdeko)` == "Checked" ~ "Other",
      `Elexacaftor/tezacaftor/ivacaftor (Trikafta)` == "Checked" ~ "Trikafta",
      .default = "None"
    ),
    `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`)
  )
# Merge
df <- full_join(chco, cnh)
# Format column types
df$`Modulator at Time of Admit` <- factor(df$`Modulator at Time of Admit`,
  levels = c("No", "Yes"), ordered = T
)
df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit` == "No"] <- "None"
df$`Specific Modulator at Admit` <- factor(tolower(df$`Specific Modulator at Admit`),
  levels = c("kalydeco", "none", "orkambi", "symdeko", "trikafta"),
  labels = c("Other", "None", "Other", "Other", "Trikafta")
)
df$Race <- factor(df$Race,
  levels = c(
    "Amer. Indian or Alaskan Naitive", "Asian",
    "Black", "Black or African American", "Black/ AA",
    "Decline to Answer", "Hispanic", "Other",
    "Unknown / Not Reported", "White"
  ),
  labels = c(
    "Other", "Other", "Black", "Black", "Black",
    "Unknown / Not Reported", "Other", "Other",
    "Unknown / Not Reported", "White"
  )
)
df$Ethnicity <- factor(df$Ethnicity,
  levels = c(
    "FALSE", "Hispanic or Latino", "HLS",
    "NOT Hispanic or Latino", "NOT HLS",
    "TRUE", "UNKNOWN"
  ),
  labels = c(
    "NOT Hispanic or Latino", "Hispanic or Latino",
    "Hispanic or Latino", "NOT Hispanic or Latino",
    "NOT Hispanic or Latino", "Hispanic or Latino",
    "Unknown"
  )
)
bugs <- c("P. aeruginosa", "P. aeruginosa mucoid", "P. aeruginosa non-mucoid", "Staph aureus", "MRSA", "S. maltophilia")
df[, bugs] <- lapply(df[, bugs], function(c) {
  c[is.na(c)] <- "No"
  c <- factor(c, levels = c("No", TRUE), labels = c("No", "Yes"), ordered = T)
})
df$`S. maltophilia`[df$Site == "CNH"] <- NA
# Make summary table
summ <- df %>%
  group_by(`Unique number`, Site, Year) %>%
  mutate(`Change in FEV1` = `FEV1 % Predicted at Discharge` - `FEV1 % Predicted at Admit`) %>%
  summarise(
    `# of Hospitalizations per Person` = n(),
    across(c(Sex:Genotype, `Specific Modulator at Admit`), ~ first(na.omit(.x))),
    across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
    across(
      all_of(c(
        "Age at Admit", "# Days in Admit", "Change in FEV1",
        "FEV1 % Predicted at Admit", "FEV1 % Predicted at Discharge"
      )),
      ~ mean(.x, na.rm = T)
    ),
    .groups = "drop"
  ) %>%
  mutate(Genotype = factor(Genotype,
    levels = c(
      "F508del/F508del", "F508del/Other",
      "Other/F508del", "Other/Other"
    ),
    labels = c(
      "F508del/F508del", "F508del/Other",
      "F508del/Other", "Other/Other"
    )
  )) %>%
  select(
    `Unique number`, `Site`, Year, `# of Hospitalizations per Person`, `# Days in Admit`, `Age at Admit`,
    `Change in FEV1`, `FEV1 % Predicted at Admit`, `FEV1 % Predicted at Discharge`,
    `Modulator at Time of Admit`, `Specific Modulator at Admit`,
    all_of(bugs), Sex:Genotype
  )
```

# Data cleaning notes

1. For each person with multiple admissions, continuous variables were averaged prior to generating summary tables.
2. If a participant was on a modulators during any admission, they were counted as on modulators for that year. For example, participant 4 from CHOC was hospitalized 3 times in 2018 and was on a modulator during one of those admissions, so they were counted as on modulators in 2018.
3. CNH only included combined P. aeruginosa information.
4. Participants were uniquely identified by ID, site, and year.

The number of hospitalizations for PEx decreased from `r sum(df$Year==2018)` in 2018 to `r sum(df$Year==2022)` in 2022.

# Summary Statistics

```{r}
#| label: tbl-all
#| tbl-cap: Summary by Year
#| message: false
summ %>%
  select(-`Unique number`, -Site) %>%
  tbl_summary(
    by = Year,
    type = list(`# of Hospitalizations per Person` ~ "continuous"),
    statistic = list(
      c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
      c(
        "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
        "FEV1 % Predicted at Discharge"
      ) ~ "{mean} ({sd})"
    ),
    missing_text = "(Missing)"
  ) %>%
  add_ci(pattern = "{stat} [{ci}]") %>%
  add_p(test = list(
    `Age at Admit` ~ "t.test",
    `Change in FEV1` ~ "t.test"
  )) %>%
  separate_p_footnotes() %>%
  as_gt() %>%
  as_raw_html()
```

```{r}
#| label: tbl-site
#| tbl-cap: Stratified by Site
#| message: false
summ %>%
  select(-`Unique number`, -`S. maltophilia`) %>%
  tbl_strata(
    strata = Site,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = Year,
          type = list(`# of Hospitalizations per Person` ~ "continuous"),
          statistic = list(
            c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
            c(
              "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
              "FEV1 % Predicted at Discharge"
            ) ~ "{mean} ({sd})"
          ),
          missing_text = "(Missing)"
        ) %>%
        add_ci(pattern = "{stat} [{ci}]") %>%
        add_p(test = list(
          `Age at Admit` ~ "t.test",
          `Change in FEV1` ~ "t.test"
        )) %>%
        separate_p_footnotes()
  ) %>%
  as_gt() %>%
  as_raw_html()
```