### 2023-04-03 Jordana Hoppe/nacfc_2023.qmd
Fleiss' kappa for Jordana

---
title: "Remote Collection of Respiratory Samples (NACFC 2023)"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(irr)
library(gtsummary)
df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
# Age
df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
# Combine race columns
df <- df %>%
  rowwise() %>%
  mutate(
    Race = paste(
      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
      Race..choice.Unknown.or.not.reported.) == "Checked"),
      collapse = "+"
    ),
    Race = factor(Race, levels = c("", "1", "5"), labels = c("Unknown", "American Indian or Alaska Native", "White"))
  )
# Shorten organism names
bugs = colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.",colnames(df))]
bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
bugs = bugs[1:16]
colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
```

# Table 1: Participant Characteristics

```{r}
#| label: tbl-1
df %>%
  select(
    Age, FEV1...predicted, FVC...predicted,
    Sex, Race, Hispanic.or.Latinx, Rural.location.
  ) %>%
  tbl_summary(
    label = list(
      Age ~ "Age at clinic culture collection",
      FEV1...predicted ~ "ppFEV", FVC...predicted ~ "ppFVC",
      Hispanic.or.Latinx ~ "Hispanic or Latinx",
      Rural.location. ~ "Rural location?"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    )
  )
```

# Pairwise Comparisons

Cohen's kappa is an agreement statistic ranging from -1 (perfect disagreement) to 1 (perfect agreement). A rough guideline for interpreting Cohen's kappa is:

| Cohen's Kappa | Interpretation         |
|---------------|------------------------|
| 0             | No agreement           |
| 0.10 - 0.20   | Slight agreement       |
| 0.21 - 0.40   | Fair agreement         |
| 0.41 - 0.60   | Moderate agreement     |
| 0.61 - 0.80   | Substantial agreement  |
| 0.81 - 0.99   | Near perfect agreement |
| 1             | Perfect agreement      |

```{r}
sens_spec <- function(gold_standard, other_test, bug, type = "Home Collection") {
  t <- table(other_test, gold_standard)
  if (all(dim(t) == c(1, 1))) {
    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("All samples marked as", unique(gold_standard))
    cat("\n")
  } else {
    # Sensitivity and specificity
    sen <- sensitivity(data = factor(other_test), reference = factor(gold_standard))
    spec <- specificity(data = factor(other_test), reference = factor(gold_standard))
    # Kappa
    k = kappa2(cbind(gold_standard,other_test))
    # Formatting
    dimnames(t) <- lapply(dimnames(t), function(n) {
      n <- sub("Checked", "Present", n)
      n <- sub("Unchecked", "Absent", n)
    })
    dimnames(t)$other_test <- paste(type, dimnames(t)$other_test)
    dimnames(t)$gold_standard <- paste("Clinic", dimnames(t)$gold_standard)

    cat("\n")
    cat("###", bug)
    cat("\n")
    print(knitr::kable(as.data.frame.matrix(t)))
    cat("\n")
    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", paste0(round(spec, 3),"."))
    cat("\n")
    cat("Cohen's kappa for the two tests was", round(k$value, 3), paste0("(p = ", format.pval(k$p.value,digits = 3, eps = 0.001),")"))
    cat("\n")
  }
}
```

## Delayed Processing

```{r results='asis'}
sens_spec(df$Methicillin.sensitive.Staph.aureus.,
  df$Methicillin.sensitive.Staph.aureus..1,
  type = "Delayed Processing", bug = "MSSA"
)
```

```{r results='asis'}
sens_spec(df$Methicillin.resistant.Staph.aureus.,
  df$Methicillin.resistant.Staph.aureus..1,
  type = "Delayed Processing", bug = "MRSA"
)
```

```{r results='asis'}
sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..1,
  type = "Delayed Processing", bug = "A. xylosoxidans"
)
```

```{r results='asis'}
sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..1,
  type = "Delayed Processing", bug = "Achromobacter (not xylosoxidans)"
)
```

```{r results='asis'}
sens_spec(df$Acinetobacter., df$Acinetobacter..1,
  type = "Delayed Processing", bug = "Acinetobacter"
)
```

```{r results='asis'}
sens_spec(df$H..influenzae., df$H..influenzae..1,
  type = "Delayed Processing", bug = "H. influenzae"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa., df$P..aeruginosa..1,
  type = "Delayed Processing", bug = "P. aeruginosa"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...1,
  type = "Delayed Processing", bug = "P. aeruginosa (mucoid)"
)
```

```{r results='asis'}
sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..1,
  type = "Delayed Processing", bug = "Pseudomonas (not aeruginosa)"
)
```

```{r results='asis'}
sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..1,
  type = "Delayed Processing", bug = "B. cepacia complex"
)
```

```{r results='asis'}
sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..1,
  type = "Delayed Processing", bug = "Stenotrophomonas"
)
```

```{r results='asis'}
sens_spec(df$Aspergillus., df$Aspergillus..1,
  type = "Delayed Processing", bug = "Aspergillus"
)
```

```{r results='asis'}
sens_spec(df$Penicillium.species., df$Penicillium.species..1,
  type = "Delayed Processing", bug = "Penicillium species"
)
```

```{r results='asis'}
sens_spec(df$Sceosporium.species., df$Sceosporium.species..1,
  type = "Delayed Processing", bug = "Sceosporium species"
)
```

```{r results='asis'}
sens_spec(df$M..avium.complex., df$M..avium.complex..1,
  type = "Delayed Processing", bug = "M. avium complex"
)
```

```{r results='asis'}
sens_spec(df$M..abscessus., df$M..abscessus..1,
  type = "Delayed Processing", bug = "M. abscessus"
)
```

## Home Collection

```{r results='asis'}
sens_spec(df$Methicillin.sensitive.Staph.aureus.,
  df$Methicillin.sensitive.Staph.aureus..2,
  type = "Home Collection", bug = "MSSA"
)
```

```{r results='asis'}
sens_spec(df$Methicillin.resistant.Staph.aureus.,
  df$Methicillin.resistant.Staph.aureus..2,
  type = "Home Collection", bug = "MRSA"
)
```

```{r results='asis'}
sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..2,
  type = "Home Collection", bug = "A. xylosoxidans"
)
```

```{r results='asis'}
sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..2,
  type = "Home Collection", bug = "Achromobacter (not xylosoxidans)"
)
```

```{r results='asis'}
sens_spec(df$Acinetobacter., df$Acinetobacter..2,
  type = "Home Collection", bug = "Acinetobacter"
)
```

```{r results='asis'}
sens_spec(df$H..influenzae., df$H..influenzae..2,
  type = "Home Collection", bug = "H. influenzae"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa., df$P..aeruginosa..2,
  type = "Home Collection", bug = "P. aeruginosa"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...2,
  type = "Home Collection", bug = "P. aeruginosa (mucoid)"
)
```

```{r results='asis'}
sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..2,
  type = "Home Collection", bug = "Pseudomonas (not aeruginosa)"
)
```

```{r results='asis'}
sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..2,
  type = "Home Collection", bug = "B. cepacia complex"
)
```

```{r results='asis'}
sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..2,
  type = "Home Collection", bug = "Stenotrophomonas"
)
```

```{r results='asis'}
sens_spec(df$Aspergillus., df$Aspergillus..2,
  type = "Home Collection", bug = "Aspergillus"
)
```

```{r results='asis'}
sens_spec(df$Penicillium.species., df$Penicillium.species..2,
  type = "Home Collection", bug = "Penicillium species"
)
```

```{r results='asis'}
sens_spec(df$Sceosporium.species., df$Sceosporium.species..2,
  type = "Home Collection", bug = "Sceosporium species"
)
```

```{r results='asis'}
sens_spec(df$M..avium.complex., df$M..avium.complex..2,
  type = "Home Collection", bug = "M. avium complex"
)
```

```{r results='asis'}
sens_spec(df$M..abscessus., df$M..abscessus..2,
  type = "Home Collection", bug = "M. abscessus"
)
```

# Agreement between all 3 methods

Fleiss' kappa is interpreted analagously to Cohen's kappa, but applies to more than two tests. Tables with NaN and NAs indicate that all tests had the same values (i.e. all absent or all present).

```{r}
fleiss = function(bug,bug_name){
  fk = kappam.fleiss(df[,c(bug,paste0(bug,".1"),paste0(bug,".2"))],detail = T)
  t = as.data.frame.matrix(fk$detail)
  t$p.value = format.pval(t$p.value,digits = 3,eps = 0.001)
  cat("\n")
    cat("##", bug_name)
    cat("\n")
    print(knitr::kable(t))
    cat("\n")
    cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value,digits = 3, eps = 0.001),")"))
    cat("\n")
}
```

```{r results='asis'}
fleiss(bug = "Methicillin.sensitive.Staph.aureus.",bug_name = "MSSA")
```

```{r results='asis'}
fleiss(bug = "Methicillin.resistant.Staph.aureus.",bug_name = "MRSA")
```

```{r results='asis'}
fleiss("A..xylosoxidans.", bug_name = "A. xylosoxidans")
```

```{r results='asis'}
fleiss("Achromobacter.not.xylosoxidans.",bug_name = "Achromobacter (not xylosoxidans)")
```

```{r results='asis'}
fleiss("Acinetobacter.", bug_name = "Acinetobacter")
```

```{r results='asis'}
fleiss("H..influenzae.", bug_name = "H. influenzae")
```

```{r results='asis'}
fleiss("P..aeruginosa.", bug_name = "P. aeruginosa")
```

```{r results='asis'}
fleiss("P..aeruginosa..mucoid..",bug_name = "P. aeruginosa (mucoid)")
```

```{r results='asis'}
fleiss("Pseudomonas.species..not.aeruginosa.",bug_name = "Pseudomonas (not aeruginosa)")
```

```{r results='asis'}
fleiss("B..cepacia.complex.", bug_name = "B. cepacia complex")
```

```{r results='asis'}
fleiss("Stenotrophomonas.", bug_name = "Stenotrophomonas")
```

```{r results='asis'}
fleiss("Aspergillus.",bug_name = "Aspergillus")
```

```{r results='asis'}
fleiss("Penicillium.species.",bug_name = "Penicillium species")
```

```{r results='asis'}
fleiss("Sceosporium.species.",bug_name = "Sceosporium species")
```

```{r results='asis'}
fleiss("M..avium.complex.",bug_name = "M. avium complex")
```

```{r results='asis'}
fleiss("M..abscessus.", bug_name = "M. abscessus")
```