### 2023-04-06 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd

---
title: "Remote Collection of Respiratory Samples (NACFC 2023)"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(irr)
library(gtsummary)
df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
# Age
df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
# Combine race columns
df <- df %>%
  rowwise() %>%
  mutate(
    Race = paste(
      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
      Race..choice.Unknown.or.not.reported.) == "Checked"),
      collapse = "+"
    ),
    Race = factor(Race, levels = c("1", "5"), labels = c("American Indian or Alaska Native", "White"))
  )
# Shorten organism names
bugs = colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.",colnames(df))]
bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
bugs = bugs[1:16]
colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
```

```{r}
#| label: tbl-1
#| tbl-cap: Participant Characteristics
df %>%
  select(
    Age, FEV1...predicted, FVC...predicted,
    Sex, Race, Hispanic.or.Latinx, Rural.location.
  ) %>%
  tbl_summary(
    label = list(
      Age ~ "Age at clinic culture collection",
      FEV1...predicted ~ "ppFEV", FVC...predicted ~ "ppFVC",
      Hispanic.or.Latinx ~ "Hispanic or Latinx",
      Rural.location. ~ "Rural location?"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    )
  )
```

### Agreement between all 3 methods

Fleiss' kappa is interpreted analagously to Cohen's kappa, but applies to more than two tests. Tables with NaN and NAs indicate that all tests had the same values (i.e. all absent or all present).

| Cohen's Kappa | Interpretation         |
|---------------|------------------------|
| 0             | No agreement           |
| 0.10 - 0.20   | Slight agreement       |
| 0.21 - 0.40   | Fair agreement         |
| 0.41 - 0.60   | Moderate agreement     |
| 0.61 - 0.80   | Substantial agreement  |
| 0.81 - 0.99   | Near perfect agreement |
| 1             | Perfect agreement      |

```{r}
fleiss = function(bug,bug_name){
  d = df[,c(bug,paste0(bug,".1"),paste0(bug,".2"))]
  colnames(d) = c("Clinic","Delayed","Home")
  d[d=="Checked"] = "Present"
  d[d=="Unchecked"] = "Absent"
  fk = kappam.fleiss(d)
  cat("\n")
    cat("##", bug_name)
    cat("\n")
    print(ftable(d))
    cat("\n")
    cat("Overall Fleiss' kappa for the three tests was", round(fk$value, 3), paste0("(p = ", format.pval(fk$p.value,digits = 3, eps = 0.001),")"))
    cat("\n")
}
```

```{r}
fleiss(bug = "Methicillin.sensitive.Staph.aureus.",bug_name = "MSSA")
```

```{r}
fleiss(bug = "Methicillin.resistant.Staph.aureus.",
       bug_name = "MRSA")
```

```{r}
fleiss("A..xylosoxidans.", bug_name = "A. xylosoxidans")
```

```{r}
fleiss("Achromobacter.not.xylosoxidans.",
       bug_name = "Achromobacter (not xylosoxidans)")
```

```{r}
fleiss("Acinetobacter.", bug_name = "Acinetobacter")
```

```{r}
fleiss("H..influenzae.", bug_name = "H. influenzae")
```

```{r}
fleiss("P..aeruginosa.", bug_name = "P. aeruginosa")
```

```{r}
fleiss("P..aeruginosa..mucoid..",bug_name = "P. aeruginosa (mucoid)")
```

```{r}
fleiss("Pseudomonas.species..not.aeruginosa.",
       bug_name = "Pseudomonas (not aeruginosa)")
```

```{r}
fleiss("B..cepacia.complex.", bug_name = "B. cepacia complex")
```

```{r}
fleiss("Stenotrophomonas.", bug_name = "Stenotrophomonas")
```

```{r}
fleiss("Aspergillus.",bug_name = "Aspergillus")
```

```{r}
fleiss("Penicillium.species.",bug_name = "Penicillium species")
```

```{r}
fleiss("Sceosporium.species.",bug_name = "Sceosporium species")
```

```{r}
fleiss("M..avium.complex.",bug_name = "M. avium complex")
```

```{r}
fleiss("M..abscessus.", bug_name = "M. abscessus")
```

### Figure

```{r message=FALSE}
f = df %>% 
  select(Subject.ID,contains("Methicillin.sensitive"),contains("H..influenzae."))
colnames(f) = c("ID",paste("MSSA",c("Clinic","Delayed","Home")),
                paste("H. influenzae",c("Clinic","Delayed","Home")))

counts=f %>% 
  pivot_longer(`MSSA Clinic`:`H. influenzae Home`,
                       names_to = c("Bug","Test"),
                       names_pattern = "(.*) (.*)") %>%
  filter(value=="Checked") %>%
  group_by(Bug,ID) %>%
  summarise(Test=paste0(Test,collapse = "+")) %>%
  count(Test) %>%
  mutate(prop = prop.table(n))

ggplot(counts,aes(x=Bug,y=n,fill=Test,label=n))+
  geom_col(colour = "black", position = "fill") +
  geom_text(aes(label = n), position = position_fill(vjust = 0.5)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Pastel1")+
  xlab("Pathogen")+ylab("% of Samples")

```