### 2023-04-10 Christine Chan/CFRD Screening Survey/cfrd_screening_analysis.Rmd
center type comparisons CFRD screening

---
title: ""
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(broom)
library(gtsummary)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,message = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
source("~/GitHub/CF/genotype_class_severity.R")
```

```{r}
endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/endo.csv",na.strings = "")
non_endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/non_endo.csv",na.strings = "")
```

# OGTT Completion Rate

## Centers with at least 50% of patients with state-funded health insurance have less than 50% OGTT completion rates (CF center directors survey, questions #4 and #6).

```{r}
# Combine levels
non_endo$state_funded = factor(non_endo$What.percent.of.your.CF.patients.have.state.funded.health.insurance..i.e...Medicaid..etc...)
levels(non_endo$state_funded) = c('less than 50%','more than 50%','less than 50%','more than 50%',NA)
non_endo$ogtt_completion = factor(non_endo$What.percent.of.eligible.CF.patients....10.years.old.at.your.center.were.screened.for.CFRD.with.an.OGTT.in.the.past.year.)
levels(non_endo$ogtt_completion) = c('less than 50%','more than 50%','less than 50%','more than 50%',NA)
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=state_funded, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

```{r}
# Logistic model
m = glm(ogtt_completion ~ state_funded,data = non_endo,family = binomial)
kable(tidy(m,exponentiate = T),digits = 3)
```

There was no association between state-funded health insurance and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`).

## Centers with an identified Endocrinologist in coordinated clinic have >50% OGTT completion rates (CF center directors survey, questions #6 and #8)

For this question, all "Other" responses were counted as "No."

```{r}
# Combine levels
non_endo$endocrinologist = factor(non_endo$Does.your.center.have.an.identified.Endocrinologist.s..to.whom.you.refer.your.CF.patients.)
levels(non_endo$endocrinologist) = c('No','No','Yes')
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=endocrinologist, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

```{r}
# Logistic model
m = glm(ogtt_completion ~ endocrinologist,data = non_endo,family = binomial)
kable(tidy(m,exponentiate = T),digits = 3)
```

There was no association between having an endocrinologist and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`).

## Centers with a protocol or standardized method for ordering OGTTs are more likely to have >50% OGTT completion rates (CF center directors survey, questions #6 and #9)

```{r}
# Combine levels
non_endo$protocol = factor(non_endo$Do.you.have.a.protocol.or.standardized.method.for.how.OGTTs.are.ordered.at.your.center.)
levels(non_endo$protocol) = c('No','No','Yes','Yes')
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=protocol, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

Unable to perform a statistical tests because there were 0 centers without a protocol and over 50% OGTT completion.

## Centers with an identified role for placing orders for OGTTs are more likely to have >50% OGTT completion rates (test for 1 role response vs 2 responses vs 3 vs 4 or more); CF center directors survey, questions #6 and #10)

```{r}
# Combine levels
non_endo$role = apply(non_endo,1,function(r){
  s = strsplit(as.character(r['Who.typically.places.the.orders..even.if.a.co.signer.is.needed..for.OGTTs.at.your.center...please.select.all.that.apply..']),",")
  return(length(s[[1]]))
})
non_endo$role = factor(non_endo$role)
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=role, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

```{r}
# Logistic model
m = glm(ogtt_completion ~ role,data = non_endo,family = binomial)
kable(tidy(m,exponentiate = T),digits = 3)
```

There was no association between OGTT ordering roles and OGTT completion.

## Centers with an identified role for contacting patients to schedule the OGTT are more likely to have >50% completion rates (test for 1 role response vs 2 responses vs 3 vs 4 or more); CF center directors survey, questions #6 and #11

```{r}
# Combine levels
non_endo$role = apply(non_endo,1,function(r){
  s = strsplit(as.character(r['Who.typically.contacts.patients.to.schedule.the.OGTT...please.select.all.that.apply..']),",")
  return(length(s[[1]]))
})
non_endo$role = factor(non_endo$role)
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=role, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

Unable to perform a statistical test due to 0 centers with 3 roles and < 50% completion.

## Centers with an identified role for providing directions for patients re: how to prepare for the OGTT are more likely to have >50% completion rates (test for 1 role response vs 2 responses vs 3 vs 4 or more); CF center directors survey, questions #6 and #12

```{r}
# Combine levels
non_endo$role = apply(non_endo,1,function(r){
  s = strsplit(as.character(r['Who.typically.provides.directions.to.patients.regarding.how.to.prepare.for.the.OGTT...please.select.all.that.apply...']),",")
  return(length(s[[1]]))
})
non_endo$role = factor(non_endo$role)
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=role, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

```{r}
# Logistic model
m = glm(ogtt_completion ~ role,data = non_endo,family = binomial)
kable(tidy(m,exponentiate = T),digits = 3)
```

There was no association between OGTT direction roles and OGTT completion.

## Centers that perform OGTTs in clinic are more likely to have >50% OGTT completion rates (CF center directors survey, questions #6 and #13)

```{r}
# Combine levels
non_endo$location = apply(non_endo,1,function(r){
  s = as.character(r['In.what.setting.s..do.your.patients.have.OGTTs.done..please.select.all.that.apply...'])
  if(grepl("In our CF clinic",s)){
    return("In clinic")
  } else {
    return("Other")
  }
})
non_endo$location = as.factor(non_endo$location)
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=location, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

```{r}
# Logistic model
m = glm(ogtt_completion ~ state_funded,data = non_endo,family = binomial)
kable(tidy(m,exponentiate = T),digits = 3)
```

There was no association between in-clinic OGTTs and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`).

## Centers that always/almost always provide reminders to patients for OGTTs are more likely to have >50% OGTT completion rates (CF center directors survey, questions #6 and #15)

```{r}
# Combine levels
non_endo$reminders = as.factor(non_endo$How.often.does.your.center.provide.routine.reminders.to.patients.when.an.OGTT.is.upcoming..due..and.or.overdue..)
levels(non_endo$reminders) = c("Other","Yes, always or almost always","Other")
```

### Cross-tab

```{r}
# Table
tbl_cross(non_endo, row=reminders, col=ogtt_completion, percent="row",missing = "no")
```

### Model results

```{r}
# Logistic model
m = glm(ogtt_completion ~ reminders,data = non_endo,family = binomial)
kable(tidy(m,exponentiate = T),digits = 3)
```

There was a significant association between reminders and OGTT completion (p = `r round(summary(m)$coefficients[2,4],3)`). Centers with regular reminders were approximately 5 times more likely to have an over 50% OGTT completion rate.

# Center Age Group

```{r}
non_endo$age_group = 
  factor(non_endo$What.patient.group.does.your.CF.center.take.care.of.)
endo$age_group = 
  factor(endo$What.age.group.of.CF.patients.do.you.take.care.of...please.select.all.that.apply.)
levels(endo$age_group) = c("Pediatric","Both","Both","Both","Adult")
```

## CF center directors

```{r results='asis'}
t1 = tableby(age_group ~ state_funded+ogtt_completion + endocrinologist + protocol + role + location +reminders,
             data = non_endo,control = list(cat.test = "fe"))
summary(t1,pfootnote = T)
```

# Combined data

```{r}
endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/endo_checkboxes.csv",na.strings = "")
non_endo = read.csv("./Christine Chan/CFRD Screening/Data_Raw/non_endo_checkboxes.csv",na.strings = "")
```

## Alternative OGTT sources

```{r}
# Combine and format
endo_alt = factor(endo$How.often.have.you.recommended.using.an.alternative.glucose.source.for.OGTT.testing..i.e...juice..pop.soda..jelly.beans..licorice..etc...)
non_endo_alt = factor(non_endo$How.often.does.your.center.use.alternative.glucose.sources.for.OGTT.testing..i.e...juice..pop.soda..jelly.beans..licorice..etc...)
length(endo_alt) = length(non_endo_alt)
levels(non_endo_alt) = c(NA,"Never","Rarely","Sometimes")
plot_df = data.frame(list("Endocrinologist" = endo_alt,"Directors" = non_endo_alt))
plot_df = plot_df %>% pivot_longer(cols = c("Endocrinologist","Directors")) %>%
  group_by(name) %>% count(value) %>% drop_na(.)
# Plot
ggplot(plot_df,aes(x = value,y = n,group = name,fill = name)) + 
  geom_bar(stat="identity",position=position_dodge()) +
  theme_bw() + theme(legend.title = element_blank()) + 
  xlab("Alternative Glucose Sources for OGTT")
```

Center directors and endocrinologists were not significantly different in terms of how often they recommend alternative glucose sources (p = 0.823).

## OGTT Time Points Collected

```{r}
# Combine and format
endo_alt = factor(endo$What.OGTT.glucose.time.points.are.collected.)
non_endo_alt = non_endo %>% 
  select(What.OGTT.time.points.for.glucose.levels.does.your.center.collect...choice.0..30..60..90..and.120.min.,
         What.OGTT.time.points.for.glucose.levels.does.your.center.collect...choice.0..60..and.120.min.,
         What.OGTT.time.points.for.glucose.levels.does.your.center.collect...choice.0.and.120.min.)
non_endo_alt = factor(apply(non_endo_alt,1,function(r){
  w = which(as.character(r) == "Checked")
  if(length(w)>0){
    names(r)[w[1]]
  } else {NA}
}))
levels(non_endo_alt) = c("0, 30, 60, 90, and 120 min","0, 60, and 120 min","0 and 120 min")
length(endo_alt) = length(non_endo_alt)
plot_df = data.frame(list("Endocrinologist" = endo_alt,"Directors" = non_endo_alt))
plot_df = plot_df %>% pivot_longer(cols = c("Endocrinologist","Directors")) %>%
  group_by(name) %>% count(value) %>% drop_na(.)
# Plot
ggplot(plot_df,aes(x = value,y = n,group = name,fill = name)) + 
  geom_bar(stat="identity",position=position_dodge()) +
  theme_bw() + theme(legend.title = element_blank()) + 
  xlab("OGTT Time Points Collected")
```

Center directors and endocrinologists were significantly different in terms of which timepoints were collected (p = 0.004).

## Other Diabetes Screening Methods

```{r}
# Combine and format
endo_alt = endo %>% select(Have.you.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.HbA1c.:Have.you.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.Other..specify.in.comment..) %>% 
  pivot_longer(everything()) %>% group_by(name) %>% summarise(n = sum(value == "Checked")) 
endo_alt$name = factor(endo_alt$name)
levels(endo_alt$name) = c("CGM","Fasting glucose","2 hr. glucose","HbA1c","Home fingerstick",
                          "Other","Random glucose")
endo_alt$type = "Endocrinologist"

non_endo_alt = non_endo %>% 
  select(Has.your.center.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.HbA1c.:
           Has.your.center.used.any.of.the.following.tests.to.screen.diagnose.CFRD.in.the.outpatient.setting...please.select.all.that.apply...choice.None.of.the.above.) %>%
  pivot_longer(everything()) %>% group_by(name) %>% summarise(n = sum(value == "Checked")) 
non_endo_alt$name = factor(non_endo_alt$name)
levels(non_endo_alt$name) = c("CGM","Fasting glucose","2 hr. glucose","HbA1c","Home fingerstick",
                              "Other","Other","Random glucose")
non_endo_alt = non_endo_alt %>% group_by(name) %>% summarise(n = sum(n))
non_endo_alt$type = "Directors"

plot_df = data.frame(rbind(endo_alt,non_endo_alt))
t = plot_df %>% pivot_wider(id_cols = name,values_from = n,names_from = type) %>%
  column_to_rownames(.,var = "name")

```

### Percentages by responder type

```{r}
# Table
kable(proportions(data.matrix(t),margin = 2)*100,digits = 2)
```

### Figure

```{r}
# Plot
ggplot(plot_df,aes(x = name,y = n,group = type,fill = type)) + 
  geom_bar(stat="identity",position=position_dodge()) +
  theme_bw() + theme(legend.title = element_blank()) + 
  xlab("Other Diabetes Screening Methods")
```

Center directors and endocrinologists were not significantly different in terms of how often they use alternative screening measures (p = 0.116).