### 2023-08-23 Stacey Martiniano/ntm_agreement.qmd
finished initial agreement, started LMMs

---
title: "NTM CT Scoring - Initial Agreement"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/Library/CloudStorage/Dropbox/Miscellaneous/american-medical-association.csl
editor: source
---

```{r data cleaning}
#| message: false
library(tidyverse)
library(redcapAPI)
library(Hmisc)
library(irr)
library(knitr)
# Import data
api <- read.table("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/api.txt", header = F)
rcon <- redcapConnection(
  url = "https://redcap.ucdenver.edu/api/",
  token = api[1, 1]
)
data <- exportRecordsTyped(rcon, labels = F)
data <- data.frame(lapply(data, as.character))
data$scan_id <- as.numeric(data$scan_id)
# Make scorer IDs by initial
data$scorer <- paste0(
  toupper(substr(data$reader_first_name, 1, 1)),
  toupper(substr(data$reader_last_name, 1, 1))
)
data$scorer[data$scorer == "GG"] <- "GS"
# Deidentify readers
old <- c("AF", "DL", "JW")
new <- c("A", "B", "C")
data$scorer[data$scorer %in% old] <- new[match(data$scorer, old, nomatch = 0)]
# Revert duplicate IDs back to original to match with gold standard
training <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/training.key.csv")
dups <- read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Stacey Martiniano/NTM CT Scoring/Data_Clean/duplicated.key.csv")
training$Training_id <- as.numeric(sub("T00", "", training$Training_id))
# Change duplicates to OG IDs
old <- dups$duplicated_ssid..dates.shifted..17.days.
new <- dups$original_ssid
data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
# Change training IDs to OG IDs
old <- training$Training_id
new <- training$Cffid.Scoring.Set
data$scan_id[data$scan_id %in% old] <- new[na.omit(match(data$scan_id, old))]
# Convert to long
data <- data %>%
  select(
    scan_id, scorer, scoresheet_timestamp,
    bronchiectasis_rul:atelectasis___6
  ) %>%
  arrange(scan_id, scorer, scoresheet_timestamp) %>%
  group_by(scan_id, scorer) %>%
  mutate(scorer_repeat = row_number()) %>%
  select(-scoresheet_timestamp) %>%
  pivot_longer(bronchiectasis_rul:atelectasis___6,
    names_to = c("feature", "location"),
    names_pattern = "(.*)_(.*)$"
  ) %>%
  mutate(
    feature = gsub("__", "", feature),
    feature = str_to_title(gsub("_", " ", feature))
  )
data$location <- factor(data$location,
  levels = c(
    "1", "2", "3", "4", "5", "6", "ling", "lll",
    "lul", "rll", "rml", "rul"
  ),
  labels = c(
    "RUL", "RML", "RLL", "LUL", "Ling", "LLL",
    "Ling", "LLL", "LUL", "RLL", "RML", "RUL"
  )
)
# Format values
data$value <- factor(data$value,
  levels = c(
    "Absent", "Checked", "Cylindric", "Cystic", "Unchecked", "Varicose"
  ),
  labels = c("Absent", "Present", "Present", "Present", "Absent", "Present"),
  ordered = T
)
# Separate out gold standards and duplicates
data_dups <- data %>% filter(scan_id %in% dups$original_ssid)
data_gs <- data %>% filter(scan_id %in% training$Cffid.Scoring.Set)
data <- data %>%
  filter(!scan_id %in% c(dups$original_ssid, training$Cffid.Scoring.Set))
# Second readings only for comparison to GS
data_gs <- data_gs %>%
  group_by(scan_id, scorer, feature, location) %>%
  slice_max(scorer_repeat)
```

# Proprortion of agreement

## Singletons

### By feature

```{r}
#| message: false
data %>%
  group_by(feature, scan_id, scorer) %>%
  summarise(v = max(value, na.rm = T)) %>%
  summarise(a = sum(v == "Present")) %>%
  summarise(
    agreed = sum(a == 0 | a == 3),
    disagreed = sum(a == 1 | a == 2),
    proportion = round(agreed / (agreed + disagreed), 3)
  ) %>%
  kable()
```

### Feature by lobe

```{r}
#| message: false
data %>%
  group_by(feature, location, scan_id) %>%
  summarise(a = factor(sum(value == "Present", na.rm = T))) %>%
  summarise(
    agreed = sum(a == 0 | a == 3),
    disagreed = sum(a == 1 | a == 2),
    proportion = round(agreed / (agreed + disagreed), 3)
  ) %>%
  kable()
```

## Duplicates

### By feature

```{r}
#| message: false
data_dups %>%
  group_by(feature, scorer, scan_id, scorer_repeat) %>%
  summarise(v = max(value, na.rm = T)) %>%
  summarise(a = v[scorer_repeat == 1] == v[scorer_repeat == 2]) %>%
  summarise(p = mean(a)) %>%
  pivot_wider(
    names_from = scorer, values_from = p,
    names_prefix = "prop. agreed "
  ) %>%
  kable()
```

### Feature by lobe

```{r}
#| message: false
#| warning: false
data_dups %>%
  group_by(feature, location, scorer, scan_id, scorer_repeat) %>%
  summarise(v = max(value, na.rm = T)) %>%
  summarise(a = v[scorer_repeat == 1] == v[scorer_repeat == 2]) %>%
  summarise(p = mean(a, na.rm = T)) %>%
  pivot_wider(
    names_from = scorer, values_from = p,
    names_prefix = "prop. agreed "
  ) %>%
  kable()
```

## Agreement with gold standard

### By feature

```{r}
#| message: false
data_gs %>%
  group_by(feature, scan_id, scorer) %>%
  summarise(v = max(value, na.rm = T)) %>%
  summarise(
    A = v[scorer == "A"] == v[scorer == "GS"],
    B = v[scorer == "B"] == v[scorer == "GS"],
    C = v[scorer == "C"] == v[scorer == "GS"]
  ) %>%
  summarise(
    `prop. agreed A` = sum(A) / n(),
    `prop. agreed B` = sum(B) / n(),
    `prop. agreed C` = sum(C) / n()
  ) %>%
  kable()
```

### Feature by lobe

Missing values were ignored in creating the table below. So if a scorer marked a feature and location blank on one scan and agreed with the gold standard on 3, their score will be 0.75 rather than 0.6.

```{r}
#| message: false
data_gs %>%
  group_by(feature, location, scan_id) %>%
  summarise(
    A = value[scorer == "A"] == value[scorer == "GS"],
    B = value[scorer == "B"] == value[scorer == "GS"],
    C = value[scorer == "C"] == value[scorer == "GS"]
  ) %>%
  summarise(
    `prop. agreed A` = mean(A, na.rm = T),
    `prop. agreed B` = mean(B, na.rm = T),
    `prop. agreed C` = mean(C, na.rm = T)
  ) %>%
  kable()
```

# Using Cohen's kappa

| Value of Kappa | Level of Agreement | % of Data that are Reliable |
| -------------- | ------------------ | --------------------------- |
| 0–.20          | None               | 0–4%                        |
| .21–.39        | Minimal            | 4–15%                       |
| .40–.59        | Weak               | 15–35%                      |
| .60–.79        | Moderate           | 35–63%                      |
| .80–.90        | Strong             | 64–81%                      |
| Above.90       | Almost Perfect     | 82–100%                     |

: Kappa interpretation table @mchughInterraterReliabilityKappa2012a {#tbl-kappa}

## Singletons

### By feature

```{r}
#| message: false
kappa <- data %>%
  group_by(feature, scan_id, scorer) %>%
  summarise(v = max(value, na.rm = T), .groups = "drop") %>%
  pivot_wider(names_from = scorer, values_from = v)
kappa <- split.data.frame(kappa, ~feature)
kappas <- lapply(names(kappa), function(n) {
  d <- kappa[[n]]
  k <- kappam.fleiss(d[, c("A", "B", "C")])
  return(c(n, k$value, k$p.value))
})
kappas <- data.frame(do.call(rbind, kappas))
colnames(kappas) <- c("feature", "kappa", "p")
kappas$kappa <- as.numeric(kappas$kappa)
kappas$p <- format.pval(as.numeric(kappas$p), digits = 3, eps = 0.001)
kable(kappas, digits = 3)
```

### Feature by lobe

Because you technically cannot calculate Cohen's kappa for matrices where every cell is the same, values of `-Inf` can be interpreted as perfect agreement for a feature that had the same value across all scans. For example, none of the 15 scans showed consolidation in the	RML.

```{r}
#| message: false
kappa <- data %>%
  pivot_wider(names_from = scorer, values_from = value)
kappa <- split.data.frame(kappa, ~ feature + location)
kappas <- lapply(names(kappa), function(n) {
  d <- kappa[[n]]
  k <- kappam.fleiss(d[, c("A", "B", "C")])
  return(c(strsplit(n, "\\.")[[1]], k$value, k$p.value))
})
kappas <- data.frame(do.call(rbind, kappas))
colnames(kappas) <- c("feature", "location", "kappa", "p")
kappas$kappa <- as.numeric(kappas$kappa)
kappas$p <- format.pval(as.numeric(kappas$p), digits = 3, eps = 0.001)
kable(kappas, digits = 3)
```

## Duplicates

Unfortunately, because there weren't many duplicate scans it was difficult to calculate kappa for many features.

### By feature

```{r}
#| message: false
kappa <- data_dups %>%
  group_by(feature, scan_id, scorer, scorer_repeat) %>%
  summarise(v = max(value, na.rm = T), .groups = "drop") %>%
  pivot_wider(names_from = c(scorer, scorer_repeat), values_from = v)
kappa <- split.data.frame(kappa, ~feature)
kappas <- lapply(names(kappa), function(n) {
  d <- kappa[[n]]
  ka <- kappa2(d[, c("A_1", "A_2")])
  kb <- kappa2(d[, c("B_1", "B_2")])
  kc <- kappa2(d[, c("C_1", "C_2")])
  return(c(
    strsplit(n, "\\.")[[1]],
    ka$value, ka$p.value,
    kb$value, kb$p.value,
    kc$value, kb$p.value
  ))
})
kappas <- data.frame(do.call(rbind, kappas))
colnames(kappas) <- c(
  "feature", "kappa A", "p A", "kappa B", "p B",
  "kappa C", "p C"
)
kappas[, !grepl("feature", colnames(kappas))] <- lapply(kappas[, !grepl("feature", colnames(kappas))], as.numeric)
kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
  format.pval,
  digits = 3, eps = 0.001
)
kable(kappas, digits = 3)
```

### Feature by lobe

```{r}
#| message: false
kappa <- data_dups %>%
  group_by(feature,location, scan_id, scorer, scorer_repeat) %>%
  pivot_wider(names_from = c(scorer, scorer_repeat), values_from = value)
kappa <- split.data.frame(kappa, ~feature+location)
kappas <- lapply(names(kappa), function(n) {
  d <- kappa[[n]]
  ka <- kappa2(d[, c("A_1", "A_2")])
  kb <- kappa2(d[, c("B_1", "B_2")])
  kc <- kappa2(d[, c("C_1", "C_2")])
  return(c(
    strsplit(n, "\\.")[[1]], 
    ka$value, ka$p.value,
    kb$value, kb$p.value,
    kc$value, kb$p.value
  ))
})
kappas <- data.frame(do.call(rbind, kappas))
colnames(kappas) <- c(
  "feature", "location","kappa A", "p A", "kappa B", "p B",
  "kappa C", "p C"
)
kappas[, !grepl("feature|location", colnames(kappas))] <- 
  lapply(kappas[, !grepl("feature|location", colnames(kappas))], as.numeric)
kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
  format.pval,
  digits = 3, eps = 0.001
)
kable(kappas, digits = 3)
```

## Agreement with gold standard

Unfortunately, because there weren't many gold standard scans it was difficult to calculate kappa for many features.

### By feature

```{r}
#| message: false
kappa <- data_gs %>%
  group_by(feature, scan_id, scorer) %>%
  summarise(v = max(value, na.rm = T), .groups = "drop") %>%
  pivot_wider(names_from = scorer, values_from = v)
kappa <- split.data.frame(kappa, ~feature)
kappas <- lapply(names(kappa), function(n) {
  d <- kappa[[n]]
  ka <- kappa2(d[, c("A", "GS")])
  kb <- kappa2(d[, c("B", "GS")])
  kc <- kappa2(d[, c("C", "GS")])
  return(c(
    strsplit(n, "\\.")[[1]],
    ka$value, ka$p.value,
    kb$value, kb$p.value,
    kc$value, kb$p.value
  ))
})
kappas <- data.frame(do.call(rbind, kappas))
colnames(kappas) <- c(
  "feature", "kappa A", "p A", "kappa B", "p B",
  "kappa C", "p C"
)
kappas[, !grepl("feature", colnames(kappas))] <- lapply(kappas[, !grepl("feature", colnames(kappas))], as.numeric)
kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
  format.pval,
  digits = 3, eps = 0.001
)
kable(kappas, digits = 3)
```

### Feature by lobe

```{r}
#| message: false
kappa <- data_gs %>%
  group_by(feature,location, scan_id, scorer) %>%
  pivot_wider(names_from = scorer, values_from = value)
kappa <- split.data.frame(kappa, ~feature+location)
kappas <- lapply(names(kappa), function(n) {
  d <- kappa[[n]]
  ka <- kappa2(d[, c("A", "GS")])
  kb <- kappa2(d[, c("B", "GS")])
  kc <- kappa2(d[, c("C", "GS")])
  return(c(
    strsplit(n, "\\.")[[1]], 
    ka$value, ka$p.value,
    kb$value, kb$p.value,
    kc$value, kb$p.value
  ))
})
kappas <- data.frame(do.call(rbind, kappas))
colnames(kappas) <- c(
  "feature", "location","kappa A", "p A", "kappa B", "p B",
  "kappa C", "p C"
)
kappas[, !grepl("feature|location", colnames(kappas))] <- 
  lapply(kappas[, !grepl("feature|location", colnames(kappas))], as.numeric)
kappas[, grep("p ", colnames(kappas))] <- lapply(kappas[, grep("p ", colnames(kappas))],
  format.pval,
  digits = 3, eps = 0.001
)
kable(kappas, digits = 3)
```