### 2023-10-10 Scott Sagel/PEx in CF/nacfc_2023.qmd
pex separate data cleaning file

---
title: "The Changing Face of Pediatric Pulmonary Exacerbations in CF (NACFC 2023)"
author: "Tim Vigers"
date: "today"
format:
  html:
    toc: false
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
execute: 
  echo: false
---

```{r}
#| include: false
library(tidyverse)
library(gtsummary)
load("/Users/timvigers/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
```

# Data cleaning notes

1. FEV1 % predicted at discharge was calculated using the `rspiro` package and GLI-2012 equations.
  - We did not distinguish between NE and SE Asian when capturing race, so all Asian participants were assumed to be NE Asian for these calculations (this was picked randomly).
  - Some participants are under age 3 and/or under 1 meter in height, so the ppFEV1 calculations may not be exactly accurate for these people.

```{r}
# Make summary table
summ <- df %>%
  group_by(sid,redcap_data_access_group,admit_year) %>%
  mutate(`Change in FEV1` = ppfev1_discharge - ppfev1_admit) %>%
  summarise(
    `# of Hospitalizations` = n(),
    across(c(sex:genotype), ~ first(na.omit(.x))),
    across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
    across(
      all_of(c(
        "Age at Admit", "# Days in Admit", "Change in FEV1",
        "FEV1 % Predicted at Admit", "FEV1 % Predicted at Discharge"
      )),
      ~ mean(.x, na.rm = T)
    ),
    .groups = "drop"
  ) %>%
  select(
    `Unique number`, `Site`, Year, `# of Hospitalizations per Person`, `# Days in Admit`, `Age at Admit`,
    `Change in FEV1`, `FEV1 % Predicted at Admit`, `FEV1 % Predicted at Discharge`,
    `Modulator at Time of Admit`, `Specific Modulator at Admit`,
    all_of(bugs), Sex:Genotype
  )
```

# Data cleaning notes

1. For each person with multiple admissions, continuous variables were averaged prior to generating summary tables.
2. If a participant was on a modulators during any admission, they were counted as on modulators for that year. For example, participant 4 from CHOC was hospitalized 3 times in 2018 and was on a modulator during one of those admissions, so they were counted as on modulators in 2018.
3. CNH only included combined P. aeruginosa information.
4. Participants were uniquely identified by ID, site, and year.

The number of hospitalizations for PEx decreased from `r sum(df$Year==2018)` in 2018 to `r sum(df$Year==2022)` in 2022.

# Summary Statistics

```{r}
#| label: tbl-all
#| tbl-cap: Summary by Year
#| message: false
summ %>%
  select(-`Unique number`, -Site) %>%
  tbl_summary(
    by = Year,
    type = list(`# of Hospitalizations per Person` ~ "continuous"),
    statistic = list(
      c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
      c(
        "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
        "FEV1 % Predicted at Discharge"
      ) ~ "{mean} ({sd})"
    ),
    missing_text = "(Missing)"
  ) %>%
  add_ci(pattern = "{stat} [{ci}]") %>%
  add_p(test = list(
    `Age at Admit` ~ "t.test",
    `Change in FEV1` ~ "t.test"
  )) %>%
  separate_p_footnotes() %>%
  as_gt() %>%
  as_raw_html()
```

```{r}
#| label: tbl-site
#| tbl-cap: Stratified by Site
#| message: false
summ %>%
  select(-`Unique number`, -`S. maltophilia`) %>%
  tbl_strata(
    strata = Site,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = Year,
          type = list(`# of Hospitalizations per Person` ~ "continuous"),
          statistic = list(
            c("# of Hospitalizations per Person", "# Days in Admit") ~ "{median} ({p25}, {p75})",
            c(
              "Age at Admit", "Change in FEV1", "FEV1 % Predicted at Admit",
              "FEV1 % Predicted at Discharge"
            ) ~ "{mean} ({sd})"
          ),
          missing_text = "(Missing)"
        ) %>%
        add_ci(pattern = "{stat} [{ci}]") %>%
        add_p(test = list(
          `Age at Admit` ~ "t.test",
          `Change in FEV1` ~ "t.test"
        )) %>%
        separate_p_footnotes()
  ) %>%
  as_gt() %>%
  as_raw_html()
```