### 2023-11-03 Christine Chan/One-hour OGTT/tim_1_hour_ogtt_analysis.qmd
CF update

---
title: "1 hour OGTT glucose in CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

```{r libraries}
#| include: false
library(tidyverse)
```

```{r data cleaning}

```

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets

```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets

```