### 2024-01-02 StatNet/LMM Power/statnet_lmm_power.qmd
StatNet power presentation

---
title: "Power and Sample Size in General Linear Mixed Models"
format: revealjs
bibliography: zotero.bib
csl: american-medical-association.csl
---

```{r include=FALSE}
library(tidyverse)
```

## A quick refresher on terminology

- Type 1 error rate ($\alpha$): the probability that we incorrectly reject the null hypothesis ($H_0$).
- Type 2 error rate ($\beta$): the probability that we incorrectly fail to reject $H_0$.
- Power ($P$): the probability that we correctly reject $H_0$.
  - Can also be written as $1-\beta$

## A quick refresher on terminology

<br>
<br>

|                     | $H_0$ True  |$H_0$ False|
| ------------------- | ----------- |---------- |
| Reject $H_0$        | $\alpha$    | $1-\beta$ |
| Do Not Reject $H_0$ | $1-\alpha$  | $\beta$   |

## The difference between two independent means

<br>

![](t_test_indep_means.jpg) @faulStatisticalPowerAnalyses2009

## Next slide

- As the deviation from the null hypothesis ($d$) approaches 0, $P$ approaches $\alpha$. @diggleAnalysisLongitudinalData2013
- We typically choose $d$ to be the smallest clinically meaningful difference that we want to detect.
- Measurement variation ($\sigma^2$): the variance of the response.
  - Ideally approximated through previous studies, but sometimes estimated based on clinical knowledge.

## References