### 2024-01-05 StatNet/LMM Power/statnet_lmm_power.qmd
statnet LMM power presentation almost done

---
title: "Power and Sample Size in General Linear Mixed Models"
format: 
  revealjs:
    self-contained: true
bibliography: zotero.bib
csl: american-medical-association.csl
---

```{r include=FALSE}
library(tidyverse)
library(simr)
```

## A quick refresher on terminology

- Type 1 error rate ($\alpha$): 
  - The probability that we incorrectly reject the null hypothesis ($H_0$).
- Type 2 error rate ($\beta$): 
  - The probability that we incorrectly fail to reject $H_0$.
- Power ($P$): 
  - The probability that we correctly reject $H_0$.
  - Can also be written as $1-\beta$

## A quick refresher on terminology

<br>
<br>

|                     | $H_0$ True  |$H_0$ False|
| ------------------- | ----------- |---------- |
| Reject $H_0$        | $\alpha$    | $1-\beta$ |
| Do Not Reject $H_0$ | $1-\alpha$  | $\beta$   |

## The difference between two independent means

<br>

![](t_test_indep_means.jpg)

## Power basics

- As the deviation from the null hypothesis ($d$) approaches 0, $P$ approaches $\alpha$.
- We typically choose $d$ to be the smallest clinically meaningful difference that we want to detect.
- Measurement variation ($\sigma^2$): 
  - The variance of the response.
  - Ideally approximated through previous studies, but sometimes estimated based on clinical knowledge.

## Power in longitudinal studies

- Number of repeated observations per person ($n$): 
  - The number of observations per person.
  - May be constrained by practical matters like budget or participant burden.
- Correlation among the repeated observations ($\rho$): 
  - The pattern of correlation among repeated observations.
  - Can sometimes be estimated based on previous studies.

## Cough monitor study

- Participants provided with AI-based cough monitor for 3 months.
  - Can distinguish between coughs and other noises.
- Cough frequency assessed by cough episodes per hour and cough seconds per hour.
- Compare the frequency of nighttime cough in study participants with CF versus healthy controls.
  - Controls will most likely be siblings of participants with CF.
  - This is the aim that we are powering on.
- Also interested in feasibility and the relationship between nighttime cough frequency and respiratory symptom scores.

## Power for a continuous outcome

- Given a fairly simple statistical model:

$$
Y_{ij} = \beta_0+\beta_1 x_i+\epsilon_{ij} \\
j=1,...,n \\ 
i=1,...,2m
$$

  - $x_i$ is the treatment indicator for the $ith$ subject.
  - Assume that $Var(\epsilon_{ij})=\sigma^2$ and $Corr(Y_{ij},Y_{ik})=\rho$ for all $j\neq k$. 
  - Also assume that each person has the same set of explanatory variables ($x_{ij}=x_{j}$).

## Power for a continuous outcome

- The required number of subjects per group is:

$$
m=\frac{2(z_\alpha+z_{1-P})^2\sigma^2\{1+(n-1)\rho\}}{nd^2}=\\
$$

$$
\frac{2(z_\alpha+z_{1-P})^2\{1+(n-1)\rho\}}{n(\frac{d}{\sigma})^2}
$$

- Need to fill in $\rho$, $d$, and $\sigma^2$

## Previous studies

- Unfortunately there are very few previous studies on this.
  - Only found one study that reported variability of night time coughs in CF.
- According to Van Der Giessen et al.@vandergiessenNocturnalCoughChildren2009 the distributions of outcomes are skewed, so they report medians and IQR.
  - For the purposes of this power calculation we are assuming normality since we don't have access to the data.
- They report: 
  - 0.9 csec/hr (IQR 0.3-6.1) the first night and 0.96 csec/hr (IQR 0.1-3.4) the second night
  - Correlation between night 1 and night 2 was 0.29 for cough seconds and 0.35 for csec/hr.
- Assume that controls will not cough and pwCF will cough approximately 0.9-1 times per hour (same for csec/hr). Also assume that $\sigma = \frac{IQR}{1.35}$ and $\rho=0.3$

# Amalia's awesome workbook (demonstration)

# GLIMMPSE demo

## SIMR

- Unfortunately not as easy to apply to this scenario since we don't have access to the prior data, but possible.
- First, set up a basic data matrix with 10 participants (one pwCF and one control from 5 families). Each person starts with 10 nights of data.
- Use the model:

$$
y∼group+time+group*time+(1|family/id)+\epsilon
$$

  - $(1|family/id)$ indicates random effects for participant nested within family.

```{r}
# Set up X matrix with 10 nights of data for 5 families
nights <- 1:10
groups <- c("Control", "CF")
families <- as.character(1:5)
X <- expand.grid(time = nights, group = groups, family = families)
X$id <- paste0(X$group, X$family)
# Set simulation parameters
b <- c(0, 8, 0, 0) # Intercept and effect of night and group
V1 <- list(0.5, 2) # Random effect variance
# Create glmer object
glmer_obj <- makeGlmer(y ~ group * time + (1 | family / id),
  family = "poisson", fixef = b, VarCorr = V1, data = X
)
glmer_obj
```

```{r}
#| cache: true
powerSim(glmer_obj, nsim = 1000, test = "group", progress = F)
```

```{r}
#| cache: true
# Stable variables
nights <- 1:10
groups <- c("Control", "CF")
b <- c(0, 8, 0, 0) # Intercept and effect of night and group
V1 <- list(0.5, 2) # Random effect variance
# Alter sample size (number of families) and plot results
n <- lapply(2:100, function(f) {
  # Set up data
  families <- as.character(1:f)
  X <- expand.grid(time = nights, group = groups, family = families)
  X$id <- paste0(X$group, X$family)
  # Create glmer object
  glmer_obj <- makeGlmer(y ~ group * time + (1 | family / id),
    family = "poisson", fixef = b, VarCorr = V1, data = X
  )
  ps = powerSim(glmer_obj, nsim = 1000, test = "group", progress = F)
})

```

## Questions for the group

- Do these assumptions seem reasonable given the limited prior knowledge?
- Does anyone have code they'd be willing to share for these simulations?
- Suggestions or software recommendations?

## References

---
nocite: |
  @faulStatisticalPowerAnalyses2009, @diggleAnalysisLongitudinalData2013, @greenSIMRPackagePower2016
---