### 2024-01-29 Emily Muther/mental_health_screening_psc.qmd
PSC screening rough draft done

---
title: "Mental Health Screening in CF - PSC Screening"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gtsummary)
library(lme4)
library(GLMMadaptive)
library(plotly)
library(knitr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| warning: false
# Import
df <- read.csv("./Emily Muther/General - Muther Research Group/Mental Health Screening/Data_Raw/PSC_Full_Data_data_11-11-22.csv",
  check.names = F, na.strings = ""
)
# Remove duplicate rows
df <- df %>% distinct()
# Sort by ID and contact date
df$CONTACT_DATE <- lubridate::mdy(df$CONTACT_DATE)
df <- df %>% arrange(PAT_MRN_ID, CONTACT_DATE)
# Exclude those with no PSC data
df <- df %>% filter(!is.na(Score))
# Make visit number
df <- df %>%
  group_by(PAT_MRN_ID) %>%
  mutate(visit_num = row_number()) %>%
  ungroup()
```

# Data cleaning

1. Duplicate rows were removed (n=21).
2. All rows with missing data for `Score` were excluded (n=2319).
3. In Table 1, demographics and first contact date are reported.

# Demographics

```{r table 1}
cat_demo_vars <- c(
  "DEPARTMENT_NAME", "ENC_T", "PAYOR_NAME"
)
cont_demo_vars <- c("AGE AT CONTACT")
df %>%
  group_by(PAT_MRN_ID) %>%
  summarise(
    across(all_of(cont_demo_vars), ~ first(na.omit(.x))),
    across(all_of(cat_demo_vars), ~ first(na.omit(.x)))
  ) %>%
  select(-PAT_MRN_ID) %>%
  tbl_summary(
    missing_text = "Missing",
    label = list("Age at First Contact" = "AGE AT CONTACT")
  )
```

# Prevalence of behavioral health difficulties

## All PSC screeners

```{r table 2}
df %>%
  select(
    `Internalizing subscale`:`Externalizing subscale`, Score,
    `PSC Attention`:`PSC Totals`
  ) %>%
  tbl_summary(
    missing_text = "Missing",
    type = list(`Internalizing subscale` ~ "continuous")
  )
```

## By person

The maximum score on each subscale was calculated for all participants.

```{r}
#| warning: false
df %>%
  mutate(across(`PSC Attention`:`PSC Totals`, ~ factor(.x, ordered = T))) %>%
  group_by(PAT_MRN_ID) %>%
  summarise(across(c(
    `Internalizing subscale`:`Externalizing subscale`, Score,
    `PSC Attention`:`PSC Totals`
  ), ~ max(.x, na.rm = T))) %>%
  mutate(across(
    c(`Internalizing subscale`:`Externalizing subscale`, Score),
    ~ case_when(.x == -Inf ~ NA, .default = .x)
  )) %>%
  select(-PAT_MRN_ID) %>%
  tbl_summary(
    missing_text = "Missing",
    type = list(`Internalizing subscale` ~ "continuous")
  )
```

# Mixed models

Because some participants had multiple PSC scores, we used mixed effects models to account for correlation within each participant. Also, due to the large number of 0 scores, we used a zero-inflated negative binomial model to examine the association between PSC score and age.

## Continuous scores

### Total score

```{r age vs. total psc}
# Plot
p <- ggplot(df, aes(
  x = AGE_AT_CONTACT_IN_YEARS, y = Score,
  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(p)
# Model
gm <- mixed_model(Score ~ AGE_AT_CONTACT_IN_YEARS,
  random = ~ 1 | PAT_MRN_ID, data = df,
  family = zi.poisson(), zi_fixed = ~1
)
kable(summary(gm)$coef_table, digits = 3)
```

### Internalizing subscale

```{r age vs. internalizing psc}
# Plot
p <- ggplot(df, aes(
  x = AGE_AT_CONTACT_IN_YEARS, y = `Internalizing subscale`,
  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(p)
# Model
gm <- mixed_model(`Internalizing subscale` ~ AGE_AT_CONTACT_IN_YEARS,
  random = ~ 1 | PAT_MRN_ID, data = df,
  family = zi.poisson(), zi_fixed = ~1, control = list(iter_EM = 0)
)
kable(summary(gm)$coef_table, digits = 3)
```

### Attention subscale

```{r age vs. attention psc}
# Plot
p <- ggplot(df, aes(
  x = AGE_AT_CONTACT_IN_YEARS, y = `Attention subscale`,
  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(p)
# Model
gm <- mixed_model(`Attention subscale` ~ AGE_AT_CONTACT_IN_YEARS,
  random = ~ 1 | PAT_MRN_ID, data = df,
  family = zi.poisson(), zi_fixed = ~1, control = list(iter_EM = 0)
)
kable(summary(gm)$coef_table, digits = 3)
```

### Externalizing subscale

```{r age vs. externalizing psc}
# Plot
p <- ggplot(df, aes(
  x = AGE_AT_CONTACT_IN_YEARS, y = `Externalizing subscale`,
  group = factor(PAT_MRN_ID), color = factor(PAT_MRN_ID)
)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none")
ggplotly(p)
# Model
gm <- mixed_model(`Externalizing subscale` ~ AGE_AT_CONTACT_IN_YEARS,
  random = ~ 1 | PAT_MRN_ID, data = df,
  family = zi.poisson(), zi_fixed = ~1, control = list(iter_EM = 0)
)
kable(summary(gm)$coef_table, digits = 3)
```

## Low vs. high risk

### Total score

```{r age vs. total psc categorical}
# Model
gm <- glmer(factor(`PSC Totals`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
  data = df, family = "binomial"
)
kable(summary(gm)$coefficients, digits = 3)
```

### Internalizing subscale

```{r age vs. internalizing psc categorical}
# Model
gm <- glmer(
  factor(`PSC Internalizing`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
  data = df, family = "binomial"
)
kable(table(factor(df$`PSC Internalizing`), df$AGE_AT_CONTACT_IN_YEARS))
```

The above model is not converging due to low number of high risk scores on this subscale.

### Attention subscale

```{r age vs. attention psc categorical}
# Model
gm <- glmer(
  factor(`PSC Attention`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
  data = df, family = "binomial"
)
kable(summary(gm)$coefficients, digits = 3)
```

### Externalizing subscale

```{r age vs. externalizing psc categorical}
# Model
gm <- glmer(
  factor(`PSC Externalizing`) ~ AGE_AT_CONTACT_IN_YEARS + (1 | PAT_MRN_ID),
  data = df, family = "binomial"
)
kable(summary(gm)$coefficients, digits = 3)
```

# Questions for Emily

1. How do we handle people with multiple screenings at the same age?