### 2024-01-31 Christine Chan/CF Registry/cf_registry_analysis.qmd
cf registry time varying covariates fixed

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    fig-width: 10
    fig-height: 10
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
execute:
  echo: false
  warning: false
editor: source
---

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(forestmodel)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(icenReg)
library(knitr)
library(eulerr)
library(broom)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

<!-- # CFRD incidence and prevalence -->

<!-- ```{r} -->
<!-- # All data -->
<!-- encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv", -->
<!--   na.strings = c("", "N/A") -->
<!-- ) -->
<!-- annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv", -->
<!--   na.strings = c("", "N/A") -->
<!-- ) -->
<!-- encounter <- encounter %>% -->
<!--   arrange(eDWID, reviewyear, encounternum) -->
<!-- annual <- annual %>% arrange(eDWID, ReviewYear) -->
<!-- # Add age -->
<!-- ages <- annual %>% -->
<!--   mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>% -->
<!--   rename(reviewyear = ReviewYear) %>% -->
<!--   group_by(eDWID, reviewyear) %>% -->
<!--   summarise( -->
<!--     `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), -->
<!--       decreasing = TRUE -->
<!--     )[1])), -->
<!--     Age = Age_YrEnd[1], .groups = "drop" -->
<!--   ) -->
<!-- encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear")) -->
<!-- encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`, -->
<!--   c(-Inf, 10, 19, 31, Inf), -->
<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
<!--   right = F -->
<!-- ) -->
<!-- encounter$`Age Group` <- cut(encounter$Age, -->
<!--   c(-Inf, 10, 19, 31, Inf), -->
<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
<!--   right = F -->
<!-- ) -->
<!-- # For those with CFRD by registry criteria, find first visit with diagnosis -->
<!-- first_cfrd_cffpr <- encounter %>% -->
<!--   filter(cfrd_status == 3) %>% -->
<!--   group_by(eDWID) %>% -->
<!--   filter(row_number() == 1) %>% -->
<!--   select(eDWID, reviewyear, `Age in 2018`) %>% -->
<!--   rename(first_cfrd_cffpr = reviewyear) -->
<!-- # Add to encounter level data -->
<!-- encounter <- left_join(encounter, first_cfrd_cffpr, -->
<!--   by = c("eDWID", "Age in 2018") -->
<!-- ) -->
<!-- # Fill down CFRD Yes/No by year -->
<!-- encounter <- encounter %>% -->
<!--   group_by(eDWID) %>% -->
<!--   mutate( -->
<!--     cfrd_cffpr = case_when( -->
<!--       reviewyear < first_cfrd_cffpr ~ "No", -->
<!--       reviewyear >= first_cfrd_cffpr ~ "Yes" -->
<!--     ) -->
<!--   ) %>% -->
<!--   ungroup() -->
<!-- encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No" -->
<!-- # Ordered factors -->
<!-- encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr, -->
<!--   levels = c("No", "Yes"), ordered = T -->
<!-- ) -->
<!-- # Remove those missing age -->
<!-- encounter <- encounter %>% filter(!is.na(Age)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: tbl-incidence_cffpr -->
<!-- #| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone" -->
<!-- # For each year, pull everyone with and without CFRD that year -->
<!-- year_inc <- function(df = encounter, year, py = 1000) { -->
<!--   # Filter and count -->
<!--   inc <- df %>% -->
<!--     filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>% -->
<!--     group_by(eDWID) %>% -->
<!--     filter(row_number() == 1) %>% -->
<!--     select(eDWID, `Age Group`, cfrd_cffpr) -->
<!--   inc <- inc %>% -->
<!--     group_by(cfrd_cffpr) %>% -->
<!--     count(`Age Group`, .drop = F) -->
<!--   inc <- inc[!is.na(inc$`Age Group`), ] -->
<!--   # Calculations -->
<!--   new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"]) -->
<!--   total_at_risk <- sum(inc$n) -->
<!--   total_inc <- new_cases / (total_at_risk / py) -->

<!--   new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"]) -->
<!--   at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"]) -->
<!--   inc_10 <- new_cases_10 / (at_risk_10 / py) -->

<!--   new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"]) -->
<!--   at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
<!--   inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py) -->

<!--   new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"]) -->
<!--   at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
<!--   inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py) -->

<!--   new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"]) -->
<!--   at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
<!--   inc_30 <- new_cases_30 / (at_risk_30 / py) -->

<!--   # Nice results -->
<!--   list( -->
<!--     `Year` = year, -->
<!--     `Total New Cases` = new_cases, -->
<!--     `Total At Risk` = total_at_risk, -->
<!--     `Total Incidence` = total_inc, -->
<!--     `Total New Cases Age < 10` = new_cases_10, -->
<!--     `Total At Risk Age < 10` = at_risk_10, -->
<!--     `Total Incidence Age < 10` = inc_10, -->
<!--     `Total New Cases Age 10-18` = new_cases_10_18, -->
<!--     `Total At Risk Age 10-18` = at_risk_10_18, -->
<!--     `Total Incidence Age 10-18` = inc_10_18, -->
<!--     `Total New Cases Age 19-30` = new_cases_19_30, -->
<!--     `Total At Risk Age 19-30` = at_risk_19_30, -->
<!--     `Total Incidence Age 19-30` = inc_19_30, -->
<!--     `Total New Cases Age > 30` = new_cases_30, -->
<!--     `Total At Risk Age > 30` = at_risk_30, -->
<!--     `Total Incidence Age > 30` = inc_30 -->
<!--   ) -->
<!-- } -->
<!-- incidence <- lapply(2003:2018, function(y) { -->
<!--   year_inc(year = y) -->
<!-- }) -->
<!-- incidence <- data.frame(do.call(rbind, incidence), check.names = F) -->
<!-- incidence <- data.frame(lapply(incidence, as.numeric), check.names = F) -->
<!-- kable(incidence, digits = 3) -->
<!-- ``` -->

<!-- In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases. -->

<!-- ```{r} -->
<!-- #| label: tbl-prevalence_cffpr -->
<!-- #| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone" -->
<!-- encounter %>% -->
<!--   group_by(eDWID, reviewyear) %>% -->
<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
<!--   group_by(reviewyear) %>% -->
<!--   summarise( -->
<!--     Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   kable(., digits = 3) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: tbl-prevalence_cffpr_year -->
<!-- #| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone" -->
<!-- encounter %>% -->
<!--   group_by(eDWID, reviewyear, `Age Group`) %>% -->
<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
<!--   group_by(reviewyear, `Age Group`) %>% -->
<!--   summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>% -->
<!--   pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>% -->
<!--   kable(., digits = 3) -->
<!-- ``` -->

<!-- In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year. -->

<!-- # CFRD screening rates -->

<!-- ```{r} -->
<!-- #| label: tbl-screening_ogtt -->
<!-- #| tbl-cap: "Yearly OGTT Screening Rates" -->
<!-- year_screen <- function(df = encounter, year) { -->
<!--   # Filter and count -->
<!--   inc <- df %>% -->
<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
<!--     mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>% -->
<!--     group_by(eDWID) %>% -->
<!--     summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1]) -->
<!--   inc <- inc %>% -->
<!--     group_by(ogtt) %>% -->
<!--     count(`Age Group`) %>% -->
<!--     filter(`Age Group` != "< 10") -->
<!--   # Calculations -->
<!--   total_screened <- sum(inc$n[inc$ogtt == 1]) -->
<!--   total_seen <- sum(inc$n) -->
<!--   screen_rate <- total_screened / total_seen * 100 -->

<!--   screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"] -->
<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->

<!--   screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"] -->
<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->

<!--   screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"] -->
<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
<!--   rate_30 <- screened_30 / seen_30 * 100 -->

<!--   # Nice results -->
<!--   list( -->
<!--     `Year` = year, -->
<!--     `Total Screened` = total_screened, -->
<!--     `Total Seen` = total_seen, -->
<!--     `Screen Rate (%)` = screen_rate, -->
<!--     `Total Screened Age 10-18` = screened_10_18, -->
<!--     `Total Seen Age 10-18` = seen_10_18, -->
<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
<!--     `Total Screened Age 19-30` = screened_19_30, -->
<!--     `Total Seen Age 19-30` = seen_19_30, -->
<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
<!--     `Total Screened Age > 30` = screened_30, -->
<!--     `Total Seen Age > 30` = seen_30, -->
<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
<!--   ) -->
<!-- } -->
<!-- screen <- lapply(2003:2018, function(y) { -->
<!--   year_screen(year = y) -->
<!-- }) -->
<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
<!-- kable(screen, digits = 3) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: tbl-screening_a1c -->
<!-- #| tbl-cap: "Yearly HbA1c Screening Rates" -->
<!-- year_screen <- function(df = encounter, year) { -->
<!--   # Filter and count -->
<!--   inc <- df %>% -->
<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
<!--     mutate(a1c = !is.na(hgba1c)) %>% -->
<!--     group_by(eDWID) %>% -->
<!--     summarise(a1c = max(a1c), `Age Group` = `Age Group`[1]) -->
<!--   inc <- inc %>% -->
<!--     group_by(a1c) %>% -->
<!--     count(`Age Group`) %>% -->
<!--     filter(`Age Group` != "< 10") -->
<!--   # Calculations -->
<!--   total_screened <- sum(inc$n[inc$a1c == 1]) -->
<!--   total_seen <- sum(inc$n) -->
<!--   screen_rate <- total_screened / total_seen * 100 -->

<!--   screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"] -->
<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->

<!--   screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"] -->
<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->

<!--   screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"] -->
<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
<!--   rate_30 <- screened_30 / seen_30 * 100 -->

<!--   # Nice results -->
<!--   list( -->
<!--     `Year` = year, -->
<!--     `Total Screened` = total_screened, -->
<!--     `Total Seen` = total_seen, -->
<!--     `Screen Rate (%)` = screen_rate, -->
<!--     `Total Screened Age 10-18` = screened_10_18, -->
<!--     `Total Seen Age 10-18` = seen_10_18, -->
<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
<!--     `Total Screened Age 19-30` = screened_19_30, -->
<!--     `Total Seen Age 19-30` = seen_19_30, -->
<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
<!--     `Total Screened Age > 30` = screened_30, -->
<!--     `Total Seen Age > 30` = seen_30, -->
<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
<!--   ) -->
<!-- } -->
<!-- screen <- lapply(2003:2018, function(y) { -->
<!--   year_screen(year = y) -->
<!-- }) -->
<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
<!-- kable(screen, digits = 3) -->
<!-- ``` -->

<!-- In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded. Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year. -->

# Data cleaning

```{r}
# Import data and codebook
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("./Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$age_group <- cut(annual$Age_YrEnd,
  breaks = c(-Inf, 10, 18, 30, Inf),
  right = F
)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
# Approximate year of birth
by <- annual %>%
  mutate(birth_year = floor(ReviewYear - Age_YrEnd)) %>%
  group_by(eDWID) %>%
  count(birth_year) %>%
  group_by(eDWID) %>%
  slice_min(birth_year) %>%
  select(-n)
annual <- left_join(annual, by, by = join_by(eDWID))
# Get first year of CFRD
first_cfrd <- annual %>%
  filter(cfrd_status_annual == "CFRD+") %>%
  group_by(eDWID) %>%
  summarise(first_cfrd = min(ReviewYear))
annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
# BMI percentile to categorical
annual$bmi_perc_cat <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc_cat <- relevel(annual$bmi_perc_cat, ref = "(50,85]")
# FEV1 to categorical
annual$ppFEV1_cat <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F
)
annual$ppFEV1_cat <- relevel(annual$ppFEV1_cat, ref = "[100,Inf)")
```

Registry fields included in our liver disease yes/no variable were:

```{r}
data.frame(data_dict$code_meaning[data_dict$var %in%
  sub("A_", "", liver_fields)]) %>%
  gt() %>%
  tab_options(column_labels.hidden = TRUE)
```

Registry fields included in our tube feeding yes/no variable were:

```{r}
data.frame(data_dict$code_meaning[
  paste0(data_dict$field_name, data_dict$code_value) %in%
    sub("A_", "", tube_fields)
]) %>%
  gt() %>%
  tab_options(column_labels.hidden = TRUE)
```

# Survival analysis - first CFRD diagnosis

Based on CFFPR annualized data. These data are interval censored because we know the first visit at which people with CFRD were diagnosed, but we do not know the exact time that they developed CFRD. These data are also left truncated because not all participants enter the study at the same time.

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc_cat", "MutClass", "A_IsOnEnzymes",
  "ppFEV1_cat", "tube_feeding", "NumPulmExacerbation",
  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
```

## Kaplan Meier Analysis

```{r}
# # Whole cohort
# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
#   data = s, cluster = eDWID
# )
# ggsurvplot(km_fit,
#   data = s, risk.table = T, cumevents = T,
#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
# )
# kable(tidy(km_fit))
# # By age group
# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
#   data = s, cluster = eDWID
# )
# ggsurvplot(km_fit,
#   data = s, risk.table = T, cumevents = T,
#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
# )
# km_table <- tidy(km_fit)
# km_table <- split.data.frame(km_table, km_table$strata)
# invisible(lapply(km_table, function(t) {
#   name <- t$strata[1]
#   t <- t %>% select(-strata)
#   kable(t, caption = name)
# }))
```

## Cox PH model

### Base model

```{r}
# Formula for base model
x <- c(
  "age_group", "Gender", "bmi_perc_cat", "MutClass", "A_IsOnEnzymes",
  "tube_feeding", "liver_disease",
  "ppFEV1_cat", "NumPulmExacerbation",
  "A_pulmonarycomplications1", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
ggforest(cox_res, data = s)
# Diagnostics
# test.ph <- cox.zph(cox_res)
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

### Time varying coefficients

- `tgroup=1` indicates the years 2003 - 2008
- `tgroup=2` indicates the years 2009 - 2013
- `tgroup=3` indicates the years 2014 - 2018

```{r results='asis'}
s_split <- survSplit(Surv(tstart, tstop, cfrd_status_annual) ~ .,
  data = s, cut = c(2008, 2013, 2018), episode = "tgroup"
)
cox_res_split <- coxph(
  Surv(tstart, tstop, cfrd_status_annual) ~ bmi_perc_cat*strata(tgroup),
  data = s_split, cluster = eDWID
)

invisible(lapply(x, function(v) {
  new_f <- as.formula(paste0(
    y, "~", paste0(x[-which(x == v)], collapse = "+"),
    paste0("+", v, "+", v, ":strata(tgroup)")
  ))
  cox_res_split <- coxph(new_f, data = s_split, cluster = eDWID)
  t <- tidy(cox_res_split, exponentiate = T, conf.int = T)
  cat("\n")
  cat("####", v)
  cat("\n")
  print(kable(t, digits = 2))
  cat("\n")
  cat("\n")
  t <- t %>%
    filter(!is.na(estimate))
  p <- t %>%
    ggplot(aes(y = fct_rev(term))) +
    geom_point(aes(x = estimate), shape = 15, size = 3) +
    geom_linerange(aes(xmin = conf.low, xmax = conf.high)) +
    geom_vline(xintercept = 1, linetype = "dashed") +
    xlab("Hazard Ratio")
  cat("\n")
  print(p)
  cat("\n")
}))
```

<!-- #### P. aeruginosa -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Aspergillus -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### ABPA -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- ### Vx770 -->

<!-- ```{r} -->
<!-- f <- update(f, . ~ . + A_Vx770) -->
<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
<!-- # Plot survival -->
<!-- ggsurvplot(survfit(cox_res), -->
<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
<!--   xlim = c(2003, 2018), break.x.by = 1 -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- # Diagnostics -->
<!-- # test.ph= cox.zph(cox_res) -->
<!-- # ggcoxzph(test.ph) -->
<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
<!-- ``` -->

<!-- #### P. aeruginosa -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Aspergillus -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### ABPA -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Mycobacterium -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- ### Vx809 -->

<!-- ```{r} -->
<!-- f <- update(f, . ~ . + A_VX809comb) -->
<!-- cox_res <- coxph(f, data = s, cluster = eDWID) -->
<!-- # Plot survival -->
<!-- ggsurvplot(survfit(cox_res), -->
<!--   data = s, risk.table = T, cumevents = T, fun = "event", -->
<!--   xlim = c(2003, 2018), break.x.by = 1 -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- # Diagnostics -->
<!-- # test.ph= cox.zph(cox_res) -->
<!-- # ggcoxzph(test.ph) -->
<!-- # ggcoxdiagnostics(cox_res,type = "martingale") -->
<!-- # ggcoxdiagnostics(cox_res,type = "deviance") -->
<!-- # ggcoxdiagnostics(cox_res,type = "score") -->
<!-- # ggcoxdiagnostics(cox_res,type = "schoenfeld") -->
<!-- ``` -->

<!-- #### P. aeruginosa -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pseudomonasaeruginosa), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Aspergillus -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_fungalyeast1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### ABPA -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_pulmonarycomplications1), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- #### Mycobacterium -->

<!-- ```{r} -->
<!-- # Update formula -->
<!-- cox_res <- coxph(update(f, . ~ . + A_mycocultureresults3), -->
<!--   data = s, cluster = eDWID -->
<!-- ) -->
<!-- # Forest plot and coefficients table -->
<!-- ggforest(cox_res, data = s) -->
<!-- ``` -->

<!-- t -->