### 2024-03-13 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
updated plots for Yaron

---
title: "Pulmonary Exacerbations in CF"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/zotero.bib
csl: /Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Miscellaneous/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(lubridate)
library(gtsummary)
library(ggpubr)
library(rspiro)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF")
# Colors for plots
yellow <- "#f8b80a"
red <- "#e30022"
purple <- "#7b2670"
blue <- "#022451"
```

```{r data cleaning}
#| warning: false
# Import
df <- read.csv("./Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
  na.strings = ""
)
# Get admit year (some missing in admit_year field)
df$year <- year(df$admit_date)
# Rename centers
df$redcap_data_access_group <- factor(df$redcap_data_access_group,
  levels = c("chco", "childrens_national", "orange_county", "rady", "seattle"),
  labels = c("CHCO", "Children's National", "Orange County", "Rady", "Seattle")
)
# CFTR modulators
df$cftr_mod <- factor(df$cftr_mod, levels = 0:1, labels = c("No", "Yes"))
df$cftr_mod_spec <- factor(df$cftr_mod_spec,
  levels = 1:4, labels = c("Kalydeco", "Orkambi", "Symdeko", "Trikafta")
)
df$Trikafta <- factor(df$cftr_mod_spec == "Trikafta",
  levels = c(F, T), labels = c("No", "Yes")
)
# % predicted FEV1
df$sex <- factor(df$sex, levels = c(2, 1), labels = c("Male", "Female"))
df$fev1_baseline <- rowMeans(df[, paste0("fev1_baseline", 1:6)], na.rm = T)
df <- df %>%
  group_by(sid) %>%
  fill(sex) %>%
  ungroup()
df$ppfev1_admit <- pctpred_GLIgl(
  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
  FEV1 = df$fev1_admit
)
df$ppfev1_baseline <- pctpred_GLIgl(
  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
  FEV1 = df$fev1_baseline
)
df$ppfev1_discharge <- pctpred_GLIgl(
  age = df$age_at_admit, height = df$height / 100, gender = as.numeric(df$sex),
  FEV1 = df$fev1_discharge
)
# Labels
labels <- list(
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "NTM testing performed?" = "ntm_test",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "A. xylosoxidans" = "admit_cx_results___1",
  "B. cepacia complex" = "admit_cx_results___2",
  "H. influenzae" = "admit_cx_results___3",
  "MRSA" = "admit_cx_results___4",
  "Methicillin-susceptible Staph aureus" = "admit_cx_results___5",
  "P. aeruginosa" = "admit_cx_results___6",
  "P. aeruginosa mucoid" = "admit_cx_results___7",
  "S. maltophilia" = "admit_cx_results___8",
  "Other bacteria" = "admit_cx_results___9",
  "On CFTR Modulator?" = "cftr_mod",
  "CFTR Modulator Type" = "cftr_mod_spec",
  "Cultures Tested" = "redcap_data_access_group",
  "Length of Admission" = "days_admit",
  "Systemic Steroids" = "steroids_system",
  "Oral Steroids" = "steroids_oral",
  "IV Steroids" = "steroids_iv"
)
labels <- setNames(names(labels), labels)
# Format
df <- df %>%
  select(
    sid, redcap_event_name, redcap_data_access_group, year, days_admit,
    cftr_mod, cftr_mod_spec, Trikafta, fungal_test, fungi___1:fungi___5,
    ntm_test, ntm___1:ntm___3,
    admit_cx_results___1:admit_cx_results___9,
    ppfev1_admit, ppfev1_discharge, ppfev1_baseline,
    contains("steroids")
  ) %>%
  filter(!is.na(year))
```

# Fungi

The tables below are limited to patients who had fungal testing performed.

```{r}
#| label: tbl-fungi
#| tbl-cap: Fungi
df %>%
  filter(fungal_test == 1) %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-fungi_center
#| tbl-cap: Fungi by Center
#| message: false
df %>%
  filter(fungal_test == 1) %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi_cftr_mod
#| tbl-cap: Fungi by Modulator Status
#| message: false
df %>%
  filter(fungal_test == 1) %>%
  select(year, cftr_mod_spec, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod_spec, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi_trikafta
#| tbl-cap: Fungi by Trikafta Status
#| message: false
df %>%
  filter(fungal_test == 1) %>%
  select(year, Trikafta, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

# NTM

The tables below are limited to patients who had NTM testing performed.

```{r}
#| label: tbl-ntm
#| tbl-cap: NTM
df %>%
  filter(ntm_test == 1) %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-ntm_center
#| tbl-cap: NTM by Center
#| message: false
df %>%
  filter(ntm_test == 1) %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm_cftr_mod
#| tbl-cap: NTM by Modulator Status
#| message: false
df %>%
  filter(ntm_test == 1) %>%
  select(year, cftr_mod_spec, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod_spec, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm_trikafta
#| tbl-cap: NTM by Trikafta Status
#| message: false
df %>%
  filter(ntm_test == 1) %>%
  select(year, Trikafta, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

# Bacteria

```{r}
#| label: tbl-bacteria
#| tbl-cap: Bacteria
df %>%
  select(
    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
  ) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-bacteria_center
#| tbl-cap: Bacteria by Center
#| message: false
df %>%
  select(
    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
  ) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria_cftr_mod
#| tbl-cap: Bacteria by Modulator Status
#| message: false
df %>%
  select(year, cftr_mod_spec, admit_cx_results___1:admit_cx_results___9) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod_spec, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria_trikafta
#| tbl-cap: Bacteria by Trikafta Status
#| message: false
df %>%
  select(year, Trikafta, admit_cx_results___1:admit_cx_results___9) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

# Days of admission

```{r}
#| label: tbl-days
#| tbl-cap: Length of Admission
df %>%
  select(year, days_admit) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-days_center
#| tbl-cap: Length of Admission by Center
#| message: false
df %>%
  select(year, redcap_data_access_group, days_admit) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-days_cftr_mod
#| tbl-cap: Length of Admission by Modulator Status
#| message: false
df %>%
  select(year, cftr_mod_spec, days_admit) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing", label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-days_trikafta
#| tbl-cap: Length of Admission by Trikafta Status
#| message: false
df %>%
  select(year, Trikafta, days_admit) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

# Steroids

```{r}
#| label: tbl-steroids
#| tbl-cap: Steroids
df %>%
  select(year, contains("steroids")) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-steroids_center
#| tbl-cap: Steroids by Center
#| message: false
df %>%
  select(year, redcap_data_access_group, contains("steroids")) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-steroids_cftr_mod
#| tbl-cap: Steroids by Modulator Status
#| message: false
df %>%
  select(year, cftr_mod_spec, contains("steroids")) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing", label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-steroids_trikafta
#| tbl-cap: Steroids by Trikafta Status
#| message: false
df %>%
  select(year, Trikafta, contains("steroids")) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

# Plots

```{r}
# Number of hospitalizations
df %>%
  group_by(year) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = factor(year), y = n, fill = factor(year))) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_label(aes(label = n),nudge_y = -20, fill = "white") +
  ylab("Num. of Hospitalizations") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  scale_fill_manual(values = c(purple, yellow))
```

```{r}
# % hospitalized 2 or more times in both years
df %>%
  group_by(sid, year) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(year) %>%
  summarise(
    num = sum(n >= 2), total = n(),
    perc = mean(n >= 2) * 100, .groups = "drop"
  ) %>%
  ggplot(aes(x = factor(year), y = perc, fill = factor(year))) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_label(aes(label = scales::percent(perc, scale = 1, accuracy = 0.1)),
    nudge_y = -5, fill = "white"
  ) +
  geom_label(aes(label = paste0(num, "/", total)),
    nudge_y = -8, fill = "white"
  ) +
  geom_signif(
    xmin = 1, xmax = 2, y_position = 50, map_signif_level = T,
    annotation = "<0.001"
  ) +
  xlab("Site") +
  ylab("Percent With >= 2 Hospitalizations") +
  theme_bw() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c(purple, yellow))
```

```{r}
# Number hospitalized 2 or more times in both years
df %>%
  group_by(sid, year) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(year) %>%
  summarise(
    num = sum(n >= 2), total = n(),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = factor(year), y = num, fill = factor(year))) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_label(aes(label = num), nudge_y = -5, fill = "white") +
  xlab("Site") +
  ylab("Number With >= 2 Hospitalizations") +
  theme_bw() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c(purple, yellow))
```

```{r}
#| warning: false
df %>%
  group_by(year, sid) %>%
  rename(
    Admission = ppfev1_admit, Discharge = ppfev1_discharge,
    Baseline = ppfev1_baseline
  ) %>%
  pivot_longer(c(Admission, Discharge, Baseline)) %>%
  mutate(name = factor(name,
    levels = c("Baseline", "Admission", "Discharge")
  )) %>%
  ggplot(aes(
    x = name, y = value,
    fill = factor(year)
  )) +
  geom_boxplot() +
  geom_pwc(
    aes(group = year),
    tip.length = 0.01,
    method = "t_test", label = "p.format"
  ) +
  ylab("FEV1 % predicted") +
  theme_bw() +
  theme(
    legend.title = element_blank(), legend.position = "bottom",
    axis.title.x = element_blank()
  ) +
  scale_fill_manual(values = c(purple, yellow))
```

FEV1 percent predicted was calculated using the GLI global (2022) equations with the `rspiro` R package.

```{r}
# Proportions who returned to >= 90% and >= 100% at discharge
df %>%
  mutate(
    returned_90 = ppfev1_discharge >= ppfev1_baseline * 0.9,
    returned_100 = ppfev1_discharge >= ppfev1_baseline
  ) %>%
  group_by(year) %>%
  summarise(
    n_returned_90 = sum(returned_90, na.rm = T),
    n_returned_100 = sum(returned_100, na.rm = T),
    total = sum(!is.na(ppfev1_discharge) & !is.na(ppfev1_baseline)),
    returned_90 = mean(returned_90, na.rm = T) * 100,
    returned_100 = mean(returned_100, na.rm = T) * 100,
  ) %>%
  ggplot(aes(x = factor(year))) +
  geom_bar(aes(y = returned_90, color = "Returned to 90%"),
    position = "dodge", stat = "identity", fill = yellow
  ) +
  geom_label(
    aes(
      y = returned_90,
      label = scales::percent(returned_90, accuracy = 0.1, scale = 1)
    ),
    nudge_y = -5, fill = "white"
  ) +
  geom_label(aes(y = returned_90, label = paste0(n_returned_90, "/", total)),
    nudge_y = -11, fill = "white"
  ) +
  geom_bar(aes(y = returned_100, color = "Returned to 100%"),
    position = "dodge", stat = "identity", fill = purple
  ) +
  geom_label(
    aes(
      y = returned_100,
      label = scales::percent(returned_100, accuracy = 0.1, scale = 1)
    ),
    nudge_y = -10, fill = "white"
  ) +
  geom_label(aes(y = returned_100, label = paste0(n_returned_100, "/", total)),
    nudge_y = -16, fill = "white"
  ) +
  ylab("Percent") +
  xlab("Year") +
  theme_classic() +
  scale_color_manual(
    name = element_blank(), values = c(purple, yellow),
    labels = c("Returned to 100%", "Returned to 90%")
  )
```

The above plot represents the proportion of admissions where the patient returned to close to baseline lung function, and was not aggregated by patient. In other words, multiple admissions for the same patient were considered independent in this case.