### 2024-03-21 Edith Zemanick/Sweat Test/sweat_test_cf_conference.qmd
CF research conference presentation rough draft

---
title: "Using Sweat Na/Cl Ratio to Predict CFTR-RD or CRMS"
author: "Tim Vigers"
date: "today"
date-format: long
format: revealjs
editor: source
bibliography: Presentation.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
---

```{r setup}
#| include: false
library(DiagrammeR)
library(readxl)
library(arsenal)
library(tidyverse)
library(factoextra)
library(pROC)
library(knitr)
library(kableExtra)
library(caret)
library(rpart)
library(rpart.plot)
source("/Users/timvigers/GitHub/CF/genotype_class_severity.R")
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/timvigers/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
n1 <- nrow(df)
n2 <- length(unique(df$PatientID))
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
# Diagnosis per Angela
df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
)
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Numeric variables
num_vars <- c("Amount", "Na", "K", "Cl")
num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
# Filter per Maxie's code
exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
df <- df %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
  filter(
    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
    Modulator == "None"
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE")
# Genotype info
## By severity
df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "_")
df$geno_risk <- factor(df$geno_risk,
  levels = c(
    "Severe_Severe", "Mild_Severe", "Mild_Unknown", "Severe_Mild", "Mild_Mild",
    "Not CF_Not CF", "Severe_Unknown", "Unknown_Severe", "Unknown_Unknown"
  ),
  labels = c(
    "High", "Low", "Low", "Low", "Low", "Not CF", "Unknown",
    "Unknown", "Unknown"
  )
)
## By function
df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")

df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")

df$geno_class <- factor(paste0(
  pmin(df$geno1_class, df$geno2_class), "/",
  pmax(df$geno1_class, df$geno2_class)
))
# Which arm?
df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
  k <- sum(!is.na(r))
  if (k == 6) {
    return("Both")
  } else if (k == 3) {
    return("Arm Not Noted")
  } else {
    return(NA)
  }
})
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sodium to chloride ratio
df$`Na/Cl Ratio` <- df$Na / df$Cl
high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
high_cl <- unique(which(df$Cl > 160))
high_na <- unique(which(df$Na > 160))
df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
# Sweat chloride group
df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
# Table options (most variables are not normally distributed)
mycontrols <- tableby.control(
  numeric.test = "kwt",
  numeric.stats = c("Nmiss", "median", "q1q3")
)
labels <- list(
  Age.at.Test.in.Years = "Age (Years)",
  geno_risk = "Genotype Severity", geno_class = "Specific Genotype"
)
# First measure for everyone
t1_df <- df %>%
  group_by(PatientID) %>%
  filter(row_number() == 1)
# Order groups
df$Diagnosis <- factor(df$Diagnosis, levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF"))
```

# Background

# Our data

## Flowchart

![](flowchart.png)

## Based on visual inspection

```{r warning=FALSE}
# Visual inspection
ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_classic()
```

## Confusion matrix

```{r}
df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`,
  breaks = c(-Inf, 1, 1.25, Inf), right = F
)
t <- table(df$nacl_ratio_cat, df$Diagnosis)
t <- addmargins(t)
# colorize the diagonal elements in table with non-ints
for (i in 1:3) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t)
```

## Confusion matrix (proportions)

```{r}
t <- round(prop.table(table(df$nacl_ratio_cat, df$Diagnosis), margin = 2) * 100, 1)
# colorize the diagonal elements in table with non-ints
for (i in 1:(ncol(t))) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t, digits = 3)
```

::: {.smaller}

*Proportions in the table above are out of the column total (i.e. by diagnosis), not the total number of participants.*

:::

# ROC analyses

## Na/Cl Ratio for CF vs. CFTR-RD or CRMS

```{r}
roc_cf <- roc(
  response = df$`Diagnosis`,
  predictor = df$`Na/Cl Ratio`,
  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
  quiet = T
)
coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)
plot(roc_cf, print.auc = TRUE, col = "blue",print.thres=T)
```

## Na/Cl Ratio for Non-CF vs. CFTR-RD or CRMS

```{r}
roc_non_cf <- roc(
  response = df$`Diagnosis`,
  predictor = df$`Na/Cl Ratio`,
  levels = c("Non-CF", "CFTR-RD or CRMS"),
  quiet = T
)
coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)
plot(roc_non_cf, print.auc = TRUE, col = "red",print.thres=T)
```

## Histograms by group

Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(
  df[!is.na(df$`Diagnosis`), ],
  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
) +
  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
  ylab("Count") +
  xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```

# K-means clustering

## Brief overview

Detail here

# K-means clustering by Na and Cl

```{r warning=FALSE}
set.seed(1017)
# K means
km_df <- df %>%
  select(Diagnosis, Na, Cl) %>%
  drop_na()
km <- kmeans(km_df[, c("Na", "Cl")], centers = 3)
km_df$Cluster <- factor(km$cluster,
  levels = 1:3,
  labels = c("High", "Medium", "Low")
)
centers <- data.frame(km$centers)
centers$Cluster <- rownames(km$centers)
centers$Cluster <- factor(centers$Cluster,
  levels = 1:3,
  labels = c("High", "Medium", "Low")
)
```

## Na by Cl plot

```{r}
ggplot(km_df, aes(x = Na, y = Cl, color = Diagnosis)) +
  geom_point(alpha = 0.5) +
  theme_classic()
```

## Na by Cl plot with k-means centers

```{r}
ggplot(km_df, aes(x = Na, y = Cl, color = Cluster)) +
  geom_point(shape = 20, alpha = 0.1) +
  geom_point(data = centers, size = 5) +
  theme_classic()
```

## K-means centers

```{r}
rownames(centers) <- NULL
centers <- centers %>% column_to_rownames("Cluster")
kable(centers, row.names = T, digits = 1)
```

## Confusion matrix

```{r}
t <- table(km_df$Cluster, km_df$Diagnosis)
t <- addmargins(t)
# colorize the diagonal elements in table with non-ints
for (i in 1:3) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t, row.names = T)
```

## Confusion matrix (proportions)

```{r}
t <- table(km_df$Cluster, km_df$Diagnosis)
t <- round(prop.table(t, margin = 2) * 100, 1)
# colorize the diagonal elements in table with non-ints
for (i in 1:(ncol(t))) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t, digits = 3)
```

# Classification trees

```{r}
cart_df <- df %>%
  select(Diagnosis, Na, Cl) %>%
  drop_na()
```

## Brief overview

Detail here

## All 3 categories

```{r}
set.seed(1017)
cart_full <- rpart(Diagnosis ~ Na + Cl, data = cart_df, method = "class",
                   control = rpart.control(cp=0))
rpart.plot(cart_full)
```

## Confusion matrix

```{r}
t <- table(predict(cart_full,type = "class"),cart_df$Diagnosis)
t <- addmargins(t)
# colorize the diagonal elements in table with non-ints
for (i in 1:3) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t)
```

## Confusion matrix (proportions)

```{r}
t <- table(predict(cart_full,type = "class"),cart_df$Diagnosis)
t <- round(prop.table(t, margin = 2) * 100, 1)
# colorize the diagonal elements in table with non-ints
for (i in 1:(ncol(t))) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t, digits = 3)
```

## CF vs. CFTR-RD or CRMS

```{r}
set.seed(1017)
cf_crms <- cart_df[cart_df$Diagnosis != "Non-CF", ]
cf_crms$Diagnosis <- droplevels(cf_crms$Diagnosis)
cart_cf_crms <- rpart(Diagnosis ~ Na + Cl, data = cf_crms, method = "class")
rpart.plot(cart_cf_crms)
```

## Confusion matrix

```{r}
t <- table(predict(cart_cf_crms,type = "class"),cf_crms$Diagnosis)
t <- addmargins(t)
# colorize the diagonal elements in table with non-ints
for (i in 1:2) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t)
```

## Confusion matrix (proportions)

```{r}
t <- table(predict(cart_cf_crms,type = "class"),cf_crms$Diagnosis)
t <- round(prop.table(t, margin = 2) * 100, 1)
# colorize the diagonal elements in table with non-ints
for (i in 1:(ncol(t))) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t, digits = 3)
```

## Non-CF vs. CFTR-RD or CRMS

```{r}
set.seed(1017)
non_cf_crms <- cart_df[cart_df$Diagnosis != "Cystic Fibrosis", ]
non_cf_crms$Diagnosis <- droplevels(non_cf_crms$Diagnosis)
cart_non_cf_crms <- rpart(Diagnosis ~ Na + Cl, data = non_cf_crms,
                          method = "class")
rpart.plot(cart_non_cf_crms)
```

## Confusion matrix

```{r}
t <- table(predict(cart_non_cf_crms,type = "class"),non_cf_crms$Diagnosis)
t <- addmargins(t)
# colorize the diagonal elements in table with non-ints
for (i in 1:2) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t)
```

## Confusion matrix (proportions)

```{r}
t<- table(predict(cart_non_cf_crms,type = "class"),non_cf_crms$Diagnosis)
t <- round(prop.table(t, margin = 2) * 100, 1)
# colorize the diagonal elements in table with non-ints
for (i in 1:(ncol(t))) {
  t[i, i] <- cell_spec(t[i, i],
    bold = T, color = "white", background = "darkgreen"
  )
}
kable(t, digits = 3)
```

## Future directions

1. Additional work on the prediction approaches to see if we can improve the results in our current data.

  - Support vector machines may allow for more flexeible partitioning of the Na/Cl space.

2. Validate final prediction models on new data.

# Questions?

## References

---
nocite: |
  @jamesIntroductionStatisticalLearning2021, @robinPROCOpensourcePackage2011, @therneau1997introduction
---

::: {#refs}
:::