### 2024-04-01 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
Update amir_nacfc_2024.qmd

---
title: "Title"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gt)
library(gtsummary)
library(lmerTest)
library(emmeans)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import data cleaned for Katie's analysis
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
  na.strings = ""
)
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
# Remove those missing all OGTT values
df <- df[rowSums(is.na(df[, glucose])) < length(glucose), ]
# Hypoglycemia
df$Hypo70 <- factor(rowSums(df[, glucose] < 70, na.rm = T) > 0,
  levels = c(F, T), labels = c("No", "Yes")
)
df$Hypo60 <- factor(rowSums(df[, glucose] < 60, na.rm = T) > 0,
  levels = c(F, T), labels = c("No", "Yes")
)
df$Hypo50 <- factor(rowSums(df[, glucose] <= 50, na.rm = T) > 0,
  levels = c(F, T), labels = c("No", "Yes")
)
```

# Hypoglycemia defined as < 70 mg/dL at any point

## Participant demographics

```{r}
#| label: tbl-70
#| tbl-cap: BG < 70 at Any Point
df %>%
  select(
    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo70
  ) %>%
  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 70}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
gt(means)
```

## Figure

```{r}
#| label: fig-mixed-model70
#| fig-cap: Model means by timepoint
ggplot(means, aes(x = Time, y = emmean, color = Hypo70, group = Hypo70)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  theme_classic() +
  ylab("Mean Glucose")
```

# Hypoglycemia defined as < 60 mg/dL at any point

## Participant demographics

```{r}
#| label: tbl-60
#| tbl-cap: BG < 60 at Any Point
df %>%
  select(
    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo60
  ) %>%
  tbl_summary(by = Hypo60, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 60}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
gt(means)
```

## Figure

```{r}
#| label: fig-mixed-model60
#| fig-cap: Model means by timepoint
ggplot(means, aes(x = Time, y = emmean, color = Hypo60, group = Hypo60)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  theme_classic() +
  ylab("Mean Glucose")
```

# Hypoglycemia defined as <= 50 mg/dL at any point

## Participant demographics

```{r}
#| label: tbl-50
#| tbl-cap: BG <= 50 at Any Point
df %>%
  select(
    age_at_current_date:cftr_mutation_2, all_of(glucose), iAUC30:iAUC180, Hypo50
  ) %>%
  tbl_summary(by = Hypo50, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 50}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
gt(means)
```

## Figure

```{r}
#| label: fig-mixed-model50
#| fig-cap: Model means by timepoint
ggplot(means, aes(x = Time, y = emmean, color = Hypo50, group = Hypo50)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  theme_classic() +
  ylab("Mean Glucose")
```