### 2024-04-05 Christine Chan/EnVision CF/amir_nacfc_2024.qmd
Update amir_nacfc_2024.qmd

---
title: "Title"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gt)
library(gtsummary)
library(lmerTest)
library(emmeans)
switch(Sys.info()[["sysname"]],
  Windows = {
    home_dir <- "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
  },
  Darwin = {
    home_dir <- "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF"
  }
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import data cleaned for Katie's analysis
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
  na.strings = ""
)
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
insulin <- gsub("timepoint_", "insulin_", glucose)
# Remove those missing all OGTT values
df <- df[rowSums(is.na(df[, glucose])) < length(glucose), ]
```

# Demographics at first visit

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_at_current_date:origin_race, cftr_mutation_1,
    cftr_mutation_2
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

# Hypoglycemia defined as < 70 mg/dL at any point

## Participant demographics

```{r}
#| label: tbl-70
#| tbl-cap: BG < 70 at Any Point
df %>%
  select(
    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
    all_of(glucose), Hypo70, all_of(insulin), hypoglycemia_symptoms_yesno,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 70}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia")
```

```{r}
#| label: fig-mixed-model70
#| fig-cap: Model means by timepoint
ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  ylab("EM Mean") +
  theme_bw() +
  facet_wrap(~Hypo70)
```

# Hypoglycemia defined as < 60 mg/dL at any point

## Participant demographics

```{r}
#| label: tbl-60
#| tbl-cap: BG < 60 at Any Point
df %>%
  select(
    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
    all_of(glucose), Hypo60, all_of(insulin), hypoglycemia_symptoms_yesno,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = Hypo60, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 60}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo60 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo60 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo60 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo60) <- c("No Hypoglycemia", "Hypoglycemia")
```

```{r}
#| label: fig-mixed-model60
#| fig-cap: Model means by timepoint
ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  ylab("EM Mean") +
  theme_bw() +
  facet_wrap(~Hypo60)
```

# Hypoglycemia defined as <= 50 mg/dL at any point

## Participant demographics

```{r}
#| label: tbl-50
#| tbl-cap: BG <= 50 at Any Point
df %>%
  select(
    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
    all_of(glucose), Hypo50, all_of(insulin), hypoglycemia_symptoms_yesno,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = Hypo50, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

## LS mean glucose by hypoglycemia status

```{r mixed models 50}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo50 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo50 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo50 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo50) <- c("No Hypoglycemia", "Hypoglycemia")
```

```{r}
#| label: fig-mixed-model50
#| fig-cap: Model means by timepoint
ggplot(plot_df, aes(x = Time, y = emmean, color = Variable, group = Variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_line() +
  ylab("EM Mean") +
  theme_bw() +
  facet_wrap(~Hypo50)
```