### 2024-04-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
race neutral lung function

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(arsenal)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
# Baseline HbA1c
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
# Race neutral equations
df$ppFEV1gl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
)
df$ppFVCgl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
)
```

# Participant characteristics prior to CFTR modulator start

```{r results='asis'}
# Table
t1 <- tableby(
  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFEV1gl + ppFVC +
    ppFVCgl + `G-tube in past 12 months` + `Pancreatic Status` + Site +
    Genotype + `Age Group` + Race,
  data = df %>% filter(`Pre or Post` == "Pre")
)
summary(t1)
```

# Pre-post

## HbA1c and others

### All groups

```{r results='asis'}
p <- paired(
  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight +
    `BMI Percentile` + ppFEV1 + ppFEV1gl + ppFVC + ppFVCgl + TDD +
    `Long acting insulin dose`,
  data = df, id = `CFF ID`,
  signed.rank.exact = FALSE,
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

### Stratified by baseline HbA1c

```{r results='asis'}
p <- paired(
  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight + `BMI Percentile` +
    ppFEV1 + ppFVC,
  data = df, id = `CFF ID`, strata = `Baseline HbA1c`,
  signed.rank.exact = FALSE,
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

### By site

```{r results='asis'}
p <- paired(
  `Pre or Post` ~ `Num. A1c Values` + HbA1c + Time + Weight +
    `BMI Percentile` + ppFEV1 + ppFEV1gl + ppFVC + ppFVCgl + TDD,
  data = df, id = `CFF ID`, strata = Site,
  signed.rank.exact = FALSE, numeric.test = "notest",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

## CGM

Only 13 of those with CGM also had HbA1c values pre- and post-CFTR modulator.

### Demographics

```{r results='asis'}
cgm <- df[!is.na(df$average_sensor), ]
# Table
t1 <- tableby(
  ~ Age + `BMI Percentile` + Weight + Sex + ppFEV1 + ppFEV1gl + ppFVC +
    ppFVCgl + `G-tube in past 12 months` + `Pancreatic Status` + Site +
    Genotype + `Age Group` + Race,
  data = cgm %>% filter(`Pre or Post` == "Pre")
)
summary(t1)
```

### Overall CGM variables

```{r results='asis'}
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
cgm_formula <- update(cgm_formula, . ~ . + TDD + `Long acting insulin dose`)
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

#### Stratified by baseline HbA1c

```{r results='asis'}
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(cgm_vars, collapse = "+")))
cgm_formula <- update(cgm_formula, . ~ . + TDD + `Long acting insulin dose`)
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

### Daytime CGM variables

```{r results='asis'}
cgm$daytime_cv <- cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
day_cgm <- c(day_cgm, "daytime_cv")
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

#### Stratified by baseline HbA1c

```{r results='asis'}
cgm$daytime_cv <- cgm$daytime_sd / cgm$daytime_avg_sensor_glucose
day_cgm <- c(day_cgm, "daytime_cv")
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(day_cgm, collapse = "+")))
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

### Nighttime CGM variables

```{r results='asis'}
cgm$nighttime_cv <- cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
night_cgm <- c(night_cgm, "nighttime_cv")
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```

#### Stratified by baseline HbA1c

```{r results='asis'}
cgm$nighttime_cv <- cgm$nighttime_sd / cgm$nighttime_avg_sens_glucose
night_cgm <- c(night_cgm, "nighttime_cv")
cgm_formula <- as.formula(paste0("`Pre or Post`~", paste0(night_cgm, collapse = "+")))
p <- paired(cgm_formula,
  data = cgm, id = `CFF ID`, strata = `Baseline HbA1c`,
  signed.rank.exact = FALSE, numeric.test = "signed.rank",
  numeric.stats = c("Nmiss", "medianrange", "q1q3")
)
summary(p, pfootnote = T, digits = 1)
```