### 2024-05-17 Edith Zemanick/Sweat Test/sweat_test_analysis_chec.qmd
added pre-modulator correlations

---
title: "CHEC Study Sweat Chloride"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(ggpmisc)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

# CHCO data

```{r}
# Edith's data
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
  na.strings = ""
)
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
# Diagnosis per Angela
df$Diagnosis <- factor(diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)],
  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
)
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Numeric variables
num_vars <- c("Amount", "Na", "K", "Cl")
num_vars <- c(paste0(num_vars, "_Left"), paste0(num_vars, "_Right"))
df[, num_vars] <- suppressWarnings(lapply(df[, num_vars], as.numeric))
# Filter per Maxie's code
exclude2 <- c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793)
no_dem <- unique(df$PatientID[df$Demographics.Available != "TRUE" | is.na(df$Demographics.Available) | df$Sex == ""])
df <- df %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
  filter(
    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
    !(Diagnosis == "Non-CF" & PatientID %in% exclude2),
    Modulator == "None"
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE")
# Genotype info
## By severity
df$geno1 <- cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "/")
df$geno_risk <- factor(df$geno_risk)
## By function
df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
df$geno_class <- factor(paste0(
  pmin(df$geno1_class, df$geno2_class), "/",
  pmax(df$geno1_class, df$geno2_class)
))
levels(df$geno_class) <- c(
  "MF/MF", "MF/MF", "MF/RF", "MF/Unknown", "MF/MF",
  "MF/RF", "MF/Unknown", "Both RF or Unknown", "Both RF or Unknown"
)
# Which arm?
df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
  k <- sum(!is.na(r))
  if (k == 6) {
    return("Both")
  } else if (k == 3) {
    return("Arm Not Noted")
  } else {
    return(NA)
  }
})
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sodium to chloride ratio
df$`Na/Cl Ratio` <- df$Na / df$Cl
high_ratio <- unique(which(df$`Na/Cl Ratio` > 2))
high_cl <- unique(which(df$Cl > 160))
high_na <- unique(which(df$Na > 160))
df <- df[-unique(c(high_ratio, high_cl, high_na)), ]
# Sweat chloride group
df$`Sweat Chloride Group` <- cut(df$Cl, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
# First measure for everyone, remove those with no sweat chloride
df <- df %>%
  filter(!is.na(Cl)) %>%
  group_by(PatientID) %>%
  filter(row_number() == 1)
# Order groups
df$Diagnosis <- factor(df$Diagnosis, levels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "Non-CF"))
```

## By sweat chloride group

```{r}
#| warning: false
ggplot(df, aes(x = Na, y = Cl, colour = `Sweat Chloride Group`)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~`Sweat Chloride Group`, scales = "free", ncol = 2)
```

## By classification

```{r}
#| warning: false
ggplot(df, aes(x = Na, y = Cl, colour = Diagnosis)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~Diagnosis, scales = "free", ncol = 2)
```

## By genotype severity

```{r}
#| warning: false
ggplot(df, aes(x = Na, y = Cl, colour = geno_class)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(df, aes(x = Na, y = Cl)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~geno_class, scales = "free", ncol = 2)
```

# CHEC data

```{r data cleaning}
# Import CHEC data
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
# Chloride groups
chec$swc_group <- cut(chec$swcavg, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
chec$swc_group_pm <- cut(chec$swhcavgpm, c(-Inf, 30, 60, Inf),
  right = F,
  labels = c("< 30", "30-59", ">60")
)
```

# Pre-modulator

## By sweat chloride group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = swc_group_pm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~swc_group_pm, scales = "free", ncol = 2)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swhsavgpm, y = swhcavgpm)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
```

# Post-modulator

## By sweat chloride group

```{r}
#| warning: false
ggplot(chec, aes(x = swsavg, y = swcavg, colour = swc_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~swc_group, scales = "free", ncol = 2)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
```

## By CHEC genotype group

```{r}
#| warning: false
ggplot(chec, aes(x = swsavg, y = swcavg, colour = genotype_group)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~genotype_group, scales = "free", ncol = 2)
```

## By modulator type

```{r}
#| warning: false
ggplot(chec, aes(x = swsavg, y = swcavg, colour = modqual)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw()
ggplot(chec, aes(x = swsavg, y = swcavg)) +
  geom_point(alpha = 0.2) +
  stat_poly_line(se = F) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_bw() +
  facet_wrap(~modqual, scales = "free", ncol = 2)
```