### 2024-07-16 Jordana Hoppe/SSwAB-IT/process_data.qmd
started process data run charts

---
title: "SSwAB-IT Process Data"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(redcapAPI)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT Process Data"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Select columns and convert to long
df <- db_sswabit_process_data %>%
  select(cf_center, week, total_swabs_ordered:total_swabs_resulted) %>%
  pivot_longer(cols = starts_with("total_"))
# Nicer names
df$name <- gsub("_", " ", df$name)
df$name <- str_to_title(df$name)
```

# Run chart

```{r}
#| label: fig-run
#| fig-width: 12
#| fig-height: 9
ggplot(df, aes(x = week, y = value, colour = cf_center, group = cf_center)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_wrap(vars(name), scales = "free_y") +
  theme(
    legend.title = element_blank(), axis.title.y = element_blank(),
    axis.title.x = element_blank(), text = element_text(size = 20)
  )
```