### 2024-07-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_merge.Rmd

---
title: ""
author: "Tim Vigers & Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---

NOTE: I HAVE THE SHARED DRIVE AS S:, to change if mapped to another letter
```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Colleen Wood"
)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

Import the datasets (I'll be doing this one site at a time to look at the variables etc)
```{r data import}
###### Import CHCO data ###########

# bmi, pulm function 577 obs
# 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
chco_bmi_pft = read_csv("./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv")
chco_bmi_pft$Date = mdy(chco_bmi_pft$Date) 
# continuous glucose monitoring, 23 obs
# 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
chco_cgm = read_csv("./Prepost Trikafta/Data_Cleaned/chco_cgm.csv")
chco_cgm$Date = date(parse_date_time(chco_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))
# chco glycemic, 261 obs
# 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
chco_glycemic = read_csv("./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv")
chco_glycemic$Date = mdy(chco_glycemic$Date)
# chco registry 108 obs
# CFF ID and MRN, start date
chco_registry = read_csv("./Prepost Trikafta/Data_Cleaned/chco_registry.csv")
###################################

###### IMPORT MONTANA DATA ########

# NEED TO RENAME VARIABLES TO MATCH CHCO#
# bmi, pulm function 975 obs
montana_bmi_pft <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_bmi_pft.csv")

# cgm 16 obs
montana_cgm <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_cgm.csv")
montana_cgm$Date = date(parse_date_time(montana_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))

# registry: we need to combine demo (126), genotypes (126), pancreatic (241), gtubes (155)
montana_demo <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_demo.csv")
montana_genotypes <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_genotypes.csv")
montana_gtubes <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
montana_pancreatic <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
# ask tim:what do we need to do with gtubes and pancreatic dates? dont look like chco data

# glycemic 279
montana_glycemic_registry <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_glycemic_registry.csv")
montana_glycemic_registry$Date = mdy(montana_glycemic_registry$Date)
# looks identical
###################################

###### IMPORT UW DATA #############
uw_registry <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_registry.csv") #cffid and start
uw_bmi <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
uw_pft <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
uw_cgm <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_cgm.csv") # same code as montana
uw_demo <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_demo.csv") 
uw_genotypes <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_registry.csv") # no genotype data here?
uw_glycemic <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
uw_gtubes <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
uw_pancreatic <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though

###################################

###### IMPORT UM DATA ########
#um_all <- read_csv("./Prepost Trikafta/Data_Cleaned/um_all.csv")
###################################

```

Lets try and stitch these sets together by site (one site at a time as above)
note: by default, the dplyr join functions join by all variables in common. So we need to make sure the common variables are same format and name

```{r form stitching}
####### CHCO ############
# CGM Data Cleaning
# get a CFF ID and visit variable by splitting subjectid variable into CFF ID and Visit
chco_cgm$`CFF ID` = as.numeric(sub("_.*", "", chco_cgm$subject_id))
chco_cgm$Visit = sub(".*_", "", chco_cgm$subject_id)

# per tim, its ok to have more than 2 dates per subject,
# as it looks like a lot of this stuff was collected at different times for different subjects;
# i expect this will make it pretty long

chco_long = full_join(chco_glycemic, chco_bmi_pft) # will join by MRN, Start, Date

chco_long = full_join(chco_registry, chco_long) # joins by MRN, CFFID, Start
chco_long = full_join(chco_long, chco_cgm) # joins by cff id and date

# add in a site id variable
chco_long$site = "CHCO"

#########################

####### Montana ############

# Race labels
factor(chco_registry$Race) # Black, Caucasian, Other, Unknown
factor(montana_demo$Race) # American Indian or Alaska Native, Black or African American, Some other race, Two or more races, White

# keep original data, add in a new one to match CHCO
montana_demo = montana_demo %>% rename(race_original = Race)
montana_demo = montana_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# rename the bmi_pft to chco names
# montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
#                                              FEV1 = `Measure of FEV1`,
#                                              Height = `Patients height`,
#                                              Weight = `Weight of Patient`)
montana_bmi_pft$Date = mdy(montana_bmi_pft$Date)

# make registry dataframe
montana_registry = full_join(montana_demo, montana_genotypes)
montana_registry = full_join(montana_registry, montana_pancreatic)
montana_registry = full_join(montana_registry, montana_gtubes)
montana_registry$Date = mdy(montana_registry$Date)

# CGM looks like this is the same as CHCO (GROOVY)
# cgm data cleaning
montana_cgm$`CFF ID` = as.numeric(sub("_.*", "", montana_cgm$subject_id))
montana_cgm$Visit = sub(".*_", "", montana_cgm$subject_id)

# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
montana_cgm$Date = mdy(montana_cgm$Date)

# combine forms
montana_long = full_join(montana_glycemic_registry, montana_bmi_pft) # will join by cffid, Date

montana_long = full_join(montana_registry, montana_long) # joins by MRN, CFFID, Start
montana_long = full_join(montana_long, montana_cgm) # joins by cff id and date

#hba1c should be a numeric, one with a note : pre trikafta cffid 1433411
montana_long = montana_long %>% rename(HbA1c_original = HbA1c)
montana_long = montana_long %>% mutate(hba1c_spc = str_detect(HbA1c_original, " "),
                               HbA1c = as.numeric(ifelse(hba1c_spc,sub(" .*", "", montana_long$HbA1c_original), HbA1c_original)))

montana_long = montana_long %>% select(-c(hba1c_spc, HbA1c_original))
# add in a site id variable
montana_long$site = "Montana"
#########################

####### UM  ############
# to do: rename variables to match CHCO
# most variables have mult columns and no notation: FEV, A1C, FVC, CFRD Status, etc
#um_all %>% select(`CFF ID`, `CFRD Status...10`, `CFRD Status...34`) %>% filter(`CFRD Status...10` != `CFRD Status...34`)

########################

####### UW #############
# Race labels
factor(chco_registry$Race) # Black, Caucasian, Other, Unknown
factor(uw_demo$Race) # American Indian or Alaska Native, Black or African American, Some other race, Two or more races, White

# keep original data, add in a new one to match CHCO
uw_demo = uw_demo %>% rename(race_original = Race)
uw_demo = uw_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# make bmi pft
uw_bmi_pft = full_join(uw_bmi, uw_pft)
# rename the bmi_pft to chco names
# uw_bmi_pft = uw_bmi_pft %>% rename(FVC = `Measure of FVC`,
#                                    FEV1 = `Measure of FEV1`,
#                                    Height = `Patients height`,
#                                    BMI = `BMI Value`)

# CGM looks like this is the same as CHCO (GROOVY)
# cgm data cleaning
uw_cgm$`CFF ID` = as.numeric(sub("_.*", "", uw_cgm$subject_id))
uw_cgm$Visit = sub(".*_", "", uw_cgm$subject_id)
# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
uw_cgm = uw_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
                               Date = ifelse(datewithtime,sub(" .*", "", uw_cgm$date_cgm_placement), date_cgm_placement))

# combine registry (leaving pancreatic and genotypes out because only a id and date variable)
uw_registry = full_join(uw_registry, uw_demo)
# uw_registry = full_join(uw_registry, uw_genotypes)
# uw_registry = full_join(uw_registry, uw_pancreatic)
uw_registry = full_join(uw_registry, uw_gtubes)

# combine forms
uw_long = full_join(uw_glycemic, uw_bmi_pft) # will join by cffid, Date
uw_long = full_join(uw_registry, uw_long) # joins by MRN, CFFID, Start
uw_long = full_join(uw_long, uw_cgm) # joins by cff id and date
uw_long$Date = mdy(uw_long$Date)
# add in a site id variable
uw_long$site = "UW"
########################
```

TO ASK TIM: 
-significance of dates? adding a ton of different rows, and wondering which dates are significant
-why do UW and Montana files (pancreatic) just have dates? are these different dates/ significant? do we need this file/ rename the date? gtubes similar

bmi pft: didnt worry about percent predicted (included in uw and montana), wasnt in chco data

```{r analysis dataset}
# full join long datasets
analysis_data = full_join(chco_long, montana_long)
analysis_data = full_join(analysis_data, uw_long)

# export csv analysis_data
write.csv(analysis_data, file = "./Prepost Trikafta/Data_Cleaned/analysis_data.csv")
```