### 2024-08-22 Christine Chan/EnVision CF/christine_CGM_OGTT_category.qmd
EnVision CGM

---
title: "CGM and OGTT Categories"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(gtsummary)
library(dplyr)
library(pracma)
library(ggplot2)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
  na.strings = ""
)
# Filter out those with no CGM data and those with no associated OGTT
df <- df %>% filter(!is.na(mean_sensor), !is.na(Date))
# Scully cutoffs
df$scully140 <- cut(df$percent_time_over_140,
  breaks = c(-Inf, 17.5, Inf), right = F
)
df$scully180 <- cut(df$percent_time_over_180,
  breaks = c(-Inf, 3.4, Inf), right = F
)
```

# Demographics at first visit

`r sum(duplicated(df$study_id))` participants wore a CGM more than once.

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
    cftr_mutation_2, weight
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

## Time spent >140 mg/dL >=17.5% vs <17.5%

```{r}
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
    cftr_mutation_2, weight, scully140
  ) %>%
  tbl_summary(by = scully140, missing_text = "Missing") %>%
  add_p() %>%
  separate_p_footnotes() %>%
  add_overall()
```

## Time spent >180 mg/dL >=3.4% vs < 3.4%

```{r}
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_visit:origin_race, cftr_mutation_1,
    cftr_mutation_2, weight, scully180
  ) %>%
  tbl_summary(by = scully180, missing_text = "Missing") %>%
  add_p() %>%
  separate_p_footnotes() %>%
  add_overall()
```

# Glycemic metrics

iAUC was calculated as the area under the glucose curve from baseline to the timepoint of interest, excluding the area under the fasting glucose level.

## All visits

```{r}
#| label: tbl-2
#| tbl-cap: Participant diagnoses based on OGTT
df %>%
  select(
    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

## Time spent >140 mg/dL >=17.5% vs <17.5%

```{r}
df %>%
  select(
    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir, scully140
  ) %>%
  tbl_summary(by = scully140, missing_text = "Missing") %>%
  add_p() %>%
  separate_p_footnotes() %>%
  add_overall()
```

## Time spent >180 mg/dL >=3.4% vs < 3.4%

```{r}
df %>%
  select(
    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir, scully180
  ) %>%
  tbl_summary(by = scully180, missing_text = "Missing") %>%
  add_p() %>%
  separate_p_footnotes() %>%
  add_overall()
```

## Visit 1 only

```{r}
#| label: tbl-3
#| tbl-cap: Participant diagnoses based on OGTT (visit 1 only)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

## Time spent >140 mg/dL >=17.5% vs <17.5%

```{r}
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir, scully140
  ) %>%
  tbl_summary(by = scully140, missing_text = "Missing") %>%
  add_p() %>%
  separate_p_footnotes() %>%
  add_overall()
```

## Time spent >180 mg/dL >=3.4% vs < 3.4%

```{r}
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, Diagnosis, a1c_result, Glucose_0, Glucose_30,
    Glucose_60, Glucose_120, Glucose_180,
    iAUC30gluc:iAUC180ins, homa_ir, scully180
  ) %>%
  tbl_summary(by = scully180, missing_text = "Missing") %>%
  add_p() %>%
  separate_p_footnotes() %>%
  add_overall()
```

# Figure

```{r}
#| label: fig-1
#| fig-cap: Percentages by visit
#| fig-width: 12
#| fig-height: 8
df %>%
  group_by(study_id) %>%
  mutate(`Visit Num.` = row_number()) %>%
  group_by(`Visit Num.`) %>%
  count(Diagnosis) %>%
  mutate(total = sum(n), perc = n / total) %>%
  ggplot(., aes(x = `Visit Num.`, y = perc, fill = Diagnosis)) +
  geom_bar(position = "stack", stat = "identity") +
  geom_text(aes(label = paste0(scales::percent(perc), " (n = ", n, ")")),
    position = position_stack(vjust = 0.5)
  ) +
  theme_bw() +
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```