### 2024-09-06 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(readxl)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import data
chco <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",
  na.strings = ""
)
njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv")
# Numeric variables
num_vars <- c(
  "Na", "K", "Cl", "Amount_Left", "Amount_Right", "Na_Left", "Na_Right",
  "K_Left", "K_Right", "Cl_Left", "Cl_Right"
)
chco[, num_vars] <- lapply(chco[, num_vars], function(c) {
  suppressWarnings(as.numeric(c))
})
# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) | diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
chco <- chco[!(chco$PatientID %in% exclude), ]
chco$Diagnosis <- factor(
  diagnoses$Diagnosis[match(
    chco$PatientID,
    diagnoses$PatientID
  )],
  levels = c("CFTR-related disorder", "CRMS", "Cystic Fibrosis", "Non-CF"),
  labels = c("CFTR-RD or CRMS", "CFTR-RD or CRMS", "Cystic Fibrosis", "Non-CF")
)
```

```{r}
chco %>%
  group_by(PatientID) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(-PatientID, -Genotypes1, -Genotypes2, -T_Tract, -Notes) %>%
  tbl_summary(by = Diagnosis, missing_text = "(Missing)")%>%
  add_overall()
```