### 2025-02-05 Lilah Melzer/cough_monitor_analysis.qmd
need to add CFQ-R plots

---
title: "Home Nighttime Cough Monitoring"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
library(GLMMadaptive)
library(tidyverse)
library(ggstats)
library(corrplot)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
```

```{r data import}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "Cough Monitor 23-0382"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
df <- exportRecordsTyped(rcon)
```

```{r data cleaning}
#| message: false
#| warning: false
# Import use data
start_end <- read.csv("./Data_Clean/Start and end dates.csv",
  na.strings = ""
)
start_end$Start <- mdy(start_end$Start)
start_end$End <- mdy(start_end$End)
vacations <- read.csv("./Data_Clean/Vacation days.csv",
  na.strings = ""
)
vacations$Vacation.Start <- mdy(vacations$Vacation.Start)
vacations$Vacation.End <- mdy(vacations$Vacation.End)
# Last date should be either the end date in REDCap or the last day with data
pull_date <- mdy("11/05/24")
start_end$End[is.na(start_end$End)] <- pull_date
start_end$End[start_end$End > pull_date] <- pull_date
# Make a grid of everyone's possible dates
dates <- start_end %>%
  rowwise() %>%
  transmute(sid,
    Date = list(seq(Start, End, by = "day"))
  ) %>%
  unnest(Date)
# Import counts
agg_counts <- read.csv(
  "./Data_Raw/Cough_w_TotalCounts_1105_NMDcorrected - Cough_agg_w_count_122324.csv",
  na.strings = c("", "NMD"), header = F
)
# Format aggregated count data
agg_counts <- data.frame(t(agg_counts))
colnames(agg_counts) <- agg_counts[1, ]
agg_counts$`Date/Device` <- NULL
agg_counts$`#nights-monitoring` <- NULL
agg_counts <- agg_counts[-1, ]
colnames(agg_counts)[1] <- "sid"
agg_counts$sid <- sub("FN", "", agg_counts$sid)
# Wide to long, fix dates
agg_counts <- agg_counts %>%
  pivot_longer(
    cols = -one_of("sid"), names_to = "Date", values_to = "Coughs"
  ) %>%
  mutate(
    Date = mdy(Date),
    Coughs = as.numeric(Coughs)
  )
# Merge with all possible dates
agg_counts$sid <- as.numeric(agg_counts$sid)
agg_counts <- left_join(dates, agg_counts)
# Vacations
agg_counts$Vacation <- apply(agg_counts, 1, function(r) {
  id <- as.numeric(r["sid"])
  d <- r["Date"]
  any(d >= vacations$Vacation.Start[vacations$sid == id] &
    d <= vacations$Vacation.End[vacations$sid == id])
})
agg_counts <- agg_counts %>%
  filter(!Vacation | is.na(Vacation)) %>%
  arrange(sid, Date)
agg_counts$Vacation <- NULL
# Sick
sick <- df %>%
  select(sid, pulm_ex_v2, pulm_ex_spec_v2) %>%
  filter(pulm_ex_v2 == "Yes")
sick$pulm_ex_abx_start <- mdy(c("4/17/24", "5/8/24"))
sick$pulm_ex_start <- sick$pulm_ex_abx_start - 7
sick$pulm_ex_end <- sick$pulm_ex_abx_start + 28
# Make a grid of everyone's sick days and merge
sick <- sick %>%
  select(sid, pulm_ex_start, pulm_ex_end) %>%
  rowwise() %>%
  transmute(sid,
    Date = list(seq(pulm_ex_start, pulm_ex_end, by = "day"))
  ) %>%
  unnest(Date) %>%
  mutate(Sick = T)
sick$sid <- as.numeric(sick$sid)
agg_counts <- left_join(agg_counts, sick)
# Format sick column
agg_counts$Sick[is.na(agg_counts$Sick)] <- F
agg_counts$Sick <- factor(agg_counts$Sick,
  levels = c(F, T),
  labels = c("No", "Yes")
)
# Count number of nights
agg_counts <- agg_counts %>%
  group_by(sid) %>%
  mutate(Night = row_number()) %>%
  ungroup()
# Demographics data
df$Diagnosis <- factor(df$redcap_event_name,
  levels = c(
    "Visits (Arm 1: CF)", "Interim (Arm 1: CF)", "Visits (Arm 2: Control)",
    "Interim (Arm 2: Control)"
  ),
  labels = c("CF", "CF", "Non-CF", "Non-CF")
)
df$Diagnosis <- relevel(df$Diagnosis, ref = "Non-CF")
demo <- df %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  select(
    sid, Diagnosis, age_at_consent, height, weight, pre_fev, post_fev,
    nih_sex:nih_ethnicity, mutation1, mutation2,
    cftr_mod, cftr_mod_type, mucolytic:antih
  )
# Calculate FEV1% - rspiro not accepting NA values for gender, which is really
# annoying
fev1 <- demo %>%
  select(sid, age_at_consent, height, weight, nih_sex, pre_fev, post_fev) %>%
  filter(nih_sex %in% c("Male", "Female"))
fev1$pre_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
  factor(fev1$nih_sex, levels = c("Male", "Female")),
  FEV1 = fev1$pre_fev
)
fev1$post_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
  factor(fev1$nih_sex, levels = c("Male", "Female")),
  FEV1 = fev1$post_fev
)
fev1$height_perc <- sds(as.numeric(fev1$height),
  age = fev1$age_at_consent, sex = fev1$nih_sex,
  male = "Male", female = "Female", item = "height2_20", ref = cdc.ref,
  type = "perc"
) * 100
fev1$weight_perc <- sds(as.numeric(fev1$weight),
  age = fev1$age_at_consent, sex = fev1$nih_sex,
  male = "Male", female = "Female", item = "weight2_20", ref = cdc.ref,
  type = "perc"
) * 100
demo <-
  left_join(
    demo,
    fev1[, c("sid", "height_perc", "weight_perc", "pre_fev_pp", "post_fev_pp")]
  )
label(demo$pre_fev_pp) <- "Pre-bronchodilator FEV1 % Predicted"
label(demo$pre_fev_pp) <- "Post-bronchodilator FEV1 % Predicted"
label(demo$height_perc) <- "Height %ile"
label(demo$weight_perc) <- "Weight %ile"
# CF genotype
demo$Genotype <-
  factor(rowSums(demo[, c("mutation1", "mutation2")] == "F508del"),
    levels = c(2, 1, 0),
    labels = c("F508del homozygous", "F508del heterozygous", "No F508del")
  )
demo[, grep("mutation", colnames(demo))] <- NULL
# Add group to counts
agg_counts$Diagnosis <-
  demo$Diagnosis[match(agg_counts$sid, as.numeric(as.character(demo$sid)))]
# Import and merge sleep times from Lilah
sleep <- read.csv("./Data_Clean/Total sleep times.csv", na.strings = "")
sleep <- sleep %>%
  select(
    Participant.First.Name..or.any.uniq.identifier.,
    Total.sleep.time..hours.
  ) %>%
  rename(
    sid = Participant.First.Name..or.any.uniq.identifier.,
    total_sleep = Total.sleep.time..hours.
  )
sleep$sid <- as.character(as.numeric(sub("FN", "", sleep$sid)))
label(sleep$total_sleep) <- "Total sleep time (hours)"
demo <- left_join(demo, sleep)
# Order columns
demo <- demo %>%
  select(sid, height_perc, weight_perc, pre_fev_pp, total_sleep, everything()) %>%
  select(-height, -weight, -pre_fev)
# Set table options for the document
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct"),
    numeric.test = "kwt"
  )
# Save
write.csv(agg_counts,
  file = "./Data_Clean/cleaned_aggregated_counts.csv",
  na = "", row.names = F
)
```

# Demographics

```{r results='asis'}
#| label: tbl-mean-demo
#| tbl-cap: Participant characteristics
f <- as.formula(paste0("Diagnosis ~", paste0(colnames(demo), collapse = "+")))
f <- update(f, . ~ . - Diagnosis - sid)
t1 <- tableby(f, data = demo, control = mycontrols)
summary(t1, pfootnote = T)
```

# Summary data

```{r results='asis'}
per_person <- agg_counts %>%
  group_by(sid, Diagnosis) %>%
  summarise(
    `Nights` = n(),
    `Nights With Data` = sum(!is.na(Coughs)),
    `Prop. Nights With Data` = sum(!is.na(Coughs)) / n(),
    `Median Coughs per Night` = median(Coughs, na.rm = T),
    .groups = "drop"
  )
# Exclude the people with 0 nights of data
per_person <- per_person %>% filter(`Nights With Data` != 0)
t2 <- tableby(
  Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
    `Median Coughs per Night`,
  data = per_person, control = mycontrols
)
summary(t2, pfootnote = T)
```

# Distribution of coughs

```{r}
#| label: fig-coughs-dist
#| fig-cap: Histogram of Coughs Across All Nights
#| warning: false
ggplot(agg_counts, aes(x = Coughs)) +
  geom_histogram(binwidth = 1)
```

The distribution of coughs per night is extremely skewed, so coughs per night, etc. are reported as medians rather than means.

# Coughs over time (aggregated)

## Plot

### All participants

```{r}
#| label: fig-coughs-by-night
#| fig-cap: Coughs by nights from start
#| warning: false
ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  geom_smooth(aes(group = Diagnosis)) +
  theme_bw()
```

### By participant

```{r results='asis'}
#| label: fig-coughs-by-night-2
#| fig-cap: Coughs by nights from start
#| warning: false
invisible(lapply(unique(agg_counts$sid), function(s) {
  d <- agg_counts[agg_counts$sid == s, ]
  diag <- d$Diagnosis[1]
  p <- ggplot(d, aes(
    x = Night, y = Coughs, group = factor(sid), color = Sick
  )) +
    geom_point() +
    geom_line() +
    theme_bw() +
    ylim(0, 50) +
    xlim(0, 150) +
    scale_color_manual(values = c("#377EB8", "#E41A1C")) +
    ggtitle(paste0("Participant ", s, " (", diag, ")"))
  if (s == 38) {
    p <- p + ylim(0, 270)
  }
  cat("\n")
  cat("\n")
  print(p)
  cat("\n")
  cat("\n")
}))
```

# Basic statistics excluding 0s

## Aggregated by person

The comparisons below exclude any nights with missing data OR 0 coughs, as well as periods of sickness. This table could potentially be a little bit confusing because the numbers presented are the median of multiple medians, but the data are still so skewed without 0s that this seemed the best way to aggregate the data.

```{r results='asis'}
#| warning: false
no_0 <- agg_counts %>%
  filter(Sick == "No", !is.na(Coughs)) %>%
  group_by(sid, Diagnosis) %>%
  mutate(`Prop. Non-0 Coughs` = mean(Coughs > 0, na.rm = T)) %>%
  filter(Coughs > 0) %>%
  summarise(
    `Prop. Non-0 Coughs` = `Prop. Non-0 Coughs`[1],
    `Median Coughs` = median(Coughs),
    `Cough IQR` = IQR(Coughs),
    .groups = "drop"
  )
no_0 <- left_join(
  no_0,
  fev1 %>%
    mutate(sid = as.numeric(sid)) %>%
    select(sid, pre_fev_pp, post_fev_pp),
  by = join_by(sid)
)
no_0_t <- tableby(
  Diagnosis ~
    kwt(`Prop. Non-0 Coughs`, "medianq1q3", "range") +
    kwt(`Median Coughs`, "medianq1q3", "range"),
  data = no_0
)
summary(no_0_t, pfootnote = T)
M <- cor.test(no_0$`Median Coughs`, no_0$pre_fev_pp, method = "spearman")
corrplot(cor(no_0 %>% select(`Median Coughs`, pre_fev_pp),
  method = "spearman", use = "pairwise.complete.obs"
))
ggplot(no_0, aes(x = `Median Coughs`, y = pre_fev_pp)) +
  geom_point()
```

Spearman correlation between median coughs and pre-bronchodilator FEV1 was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,eps = 0.001,digits = 3)`).

## Not aggregated

```{r results='asis'}
#| warning: false
no_0 <- agg_counts %>% filter(Sick == "No", !is.na(Coughs), Coughs > 0)
no_0 <- left_join(
  no_0,
  fev1 %>%
    mutate(sid = as.numeric(sid)) %>%
    select(sid, pre_fev_pp, post_fev_pp),
  by = join_by(sid)
)
no_0_t <- tableby(Diagnosis ~ kwt(Coughs, "medianq1q3", "range"), data = no_0)
summary(no_0_t, pfootnote = T)
```

Spearman correlation between median coughs and pre-bronchodilator FEV1 was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,eps = 0.001,digits = 3)`).

# CFQ-R

```{r}
# CFQ-R Caregivers 6-13
# SAS code
# if nmiss (phys1, phys2, phys3, phys4, phys5, phys6) <= 3 then
# physical = (mean (phys1, phys2, phys3, phys4, phys5, phys6)-1)/3*100;
#
# if nmiss (emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14) <= 4 then
# emotion = (mean (emot7, emot8, emot9, emot10, emot11, emot12, emot13, emot14)-1)/3*100;
#
# if nmiss (social20, social21, social22, social23, social24, social25, social26) <= 3 then
# social = (mean (social20, social21, social22, social23, social24, social25, social26)-1)/3*100;
#
# if nmiss (eat15, eat17, eat19) <= 1 then
# eat = (mean(eat15, eat17, eat19)-1)/3*100;
#
# if nmiss (body27, body28, body29) <= 1 then
# body = (mean(body27, body28, body29)-1)/3*100;
#
# if nmiss (treat16, treat18, treat30) <= 1 then
# treat = (mean (treat16, treat18, treat30)-1)/3*100;
#
# if nmiss (resp31, resp32, resp33, resp34) <= 2 then
# respirat = (mean (resp31, resp32, resp33, resp34)-1)/3*100;
#
# if nmiss (digest35) = 0 then
# digest = (mean (digest35)-1)/3*100;

# Caregivers of children ages 6-13
# Rename columns so we can use the code from the CFF
cfqr_parent_6_13_names <- list(
  "cfq_r_p_1" = "phys1", "cfq_r_p_2" = "phys2", "cfq_r_p_3" = "phys3",
  "cfq_r_p_4" = "phys4", "cfq_r_p_5" = "phys5", "cfq_r_p_6" = "emot6",
  "cfq_r_p_7" = "emot7", "cfq_r_p_8" = "vital8", "cfq_r_p_9" = "vital9",
  "cfq_r_p_10" = "vital10", "cfq_r_p_11" = "vital11", "cfq_r_p_12" = "vital12",
  "cfq_r_p_13" = "school13", "cfq_r_p_14" = "phys14", "cfq_r_p_15" = "phys15",
  "cfq_r_p_16" = "phys16", "cfq_r_p_17" = "eat17", "cfq_r_p_18" = "treat18",
  "cfq_r_p_19" = "body19", "cfq_r_p_20" = "body20", "cfq_r_p_21" = "body21",
  "cfq_r_p_22" = "health22", "cfq_r_p_23" = "emot23", "cfq_r_p_24" = "health24",
  "cfq_r_p_25" = "emot25", "cfq_r_p_26" = "emot26", "cfq_r_p_27" = "school27",
  "cfq_r_p_28" = "school28", "cfq_r_p_29" = "school29",
  "cfq_r_p_30" = "treat30", "cfq_r_p_31" = "treat31", "cfq_r_p_32" = "health32",
  "cfq_r_p_33" = "weight33", "cfq_r_p_34" = "resp34", "cfq_r_p_35" = "resp35",
  "cfq_r_p_36" = "resp36", "cfq_r_p_37" = "resp37", "cfq_r_p_38" = "resp38",
  "cfq_r_p_39" = "resp39", "cfq_r_p_40" = "resp40", "cfq_r_p_41" = "digest41",
  "cfq_r_p_42" = "digest42", "cfq_r_p_43" = "digest43", "cfq_r_p_44" = "eat44"
)
cfqr_parent_6_13_names <-
  setNames(names(cfqr_parent_6_13_names), cfqr_parent_6_13_names)
# Some items need to be reverse scored
cfqr_parent_6_13_reverse <- c(
  "emot6", "vital10", "vital12", "phys15", "treat31", "health22", "health24",
  "health32", "school28", "resp37"
)
# Do it!
cfqr_parent_6_13 <- df %>%
  select(
    sid, redcap_event_name:redcap_repeat_instance,
    cfq_r_p_1:cfq_r_p_44
  ) %>%
  mutate(across(cfq_r_p_1:cfq_r_p_44, ~ as.numeric(.x))) %>%
  rename(all_of(unlist(cfqr_parent_6_13_names))) %>%
  mutate(across(all_of(cfqr_parent_6_13_reverse), ~ 5 - .x)) %>%
  rowwise() %>%
  mutate(
    physical =
      ifelse(
        sum(is.na(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16))) <= 4,
        (mean(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16)) - 1)
        / 3 * 100, NA
      ),
    emotion =
      ifelse(
        sum(is.na(c(emot6, emot7, emot23, emot25, emot26))) <= 2,
        (mean(c(emot6, emot7, emot23, emot25, emot26)) - 1)
        / 3 * 100, NA
      ),
    vitality =
      ifelse(
        sum(is.na(c(vital8, vital9, vital10, vital11, vital12))) <= 2,
        (mean(c(vital8, vital9, vital10, vital11, vital12)) - 1)
        / 3 * 100, NA
      ),
    school =
      ifelse(
        sum(is.na(c(school13, school27, school28, school29))) <= 2,
        (mean(c(school13, school27, school28, school29)) - 1)
        / 3 * 100, NA
      ),
    eat =
      ifelse(
        sum(is.na(c(eat17, eat44))) == 0,
        (mean(c(eat17, eat44)) - 1)
        / 3 * 100, NA
      ),
    body =
      ifelse(
        sum(is.na(c(body19, body20, body21))) <= 1,
        (mean(c(body19, body20, body21)) - 1)
        / 3 * 100, NA
      ),
    treat =
      ifelse(
        sum(is.na(c(treat18, treat30, treat31))) <= 1,
        (mean(c(treat18, treat30, treat31)) - 1)
        / 3 * 100, NA
      ),
    health =
      ifelse(
        sum(is.na(c(health22, health24, health32))) <= 1,
        (mean(c(health22, health24, health32)) - 1)
        / 3 * 100, NA
      ),
    respirat =
      ifelse(
        sum(is.na(c(resp34, resp35, resp36, resp38, resp39, resp40))) <= 3,
        (mean(c(resp34, resp35, resp36, resp38, resp39, resp40)) - 1)
        / 3 * 100, NA
      ),
    digest =
      ifelse(
        sum(is.na(c(digest41, digest42, digest43))) <= 1,
        (mean(c(digest41, digest42, digest43)) - 1)
        / 3 * 100, NA
      ),
    weight =
      ifelse(
        sum(is.na(c(weight33))) == 0,
        (mean(c(weight33)) - 1)
        / 3 * 100, NA
      )
  )
check <- left_join(
  df %>%
    select(
      sid, redcap_event_name:redcap_repeat_instance,
      cfq_r_p_1:cfq_r_p_44
    ),
  cfqr_parent_6_13 %>%
    select(sid, redcap_event_name:redcap_repeat_instance, physical:weight),
  join_by(
    sid, redcap_event_name, redcap_repeat_instrument,
    redcap_repeat_instance
  )
)
write.csv(check, file = "./Data_Clean/cfqr_parent_6_13_check.csv", na = "", row.names = F)

cfqr_children_6_11 <- list(
  "cfq_r_y_1" = "phys1", "cfq_r_y_2" = "phys2", "cfq_r_y_3" = "phys3",
  "cfq_r_y_4" = "phys4", "cfq_r_y_5" = "phys5", "cfq_r_y_6" = "phys6",
  "cfq_r_y_7" = "emot7", "cfq_r_y_8" = "emot8", "cfq_r_y_9" = "emot9",
  "cfq_r_y_10" = "emot10", "cfq_r_y_11" = "emot11", "cfq_r_y_12" = "emot12",
  "cfq_r_y_13" = "emot13", "cfq_r_y_14" = "emot14", "cfq_r_y_15" = "eat15",
  "cfq_r_y_16" = "treat16", "cfq_r_y_17" = "eat17", "cfq_r_y_18" = "treat18",
  "cfq_r_y_19" = "eat19", "cfq_r_y_20" = "social20", "cfq_r_y_21" = "social21",
  "cfq_r_y_22" = "social22", "cfq_r_y_23" = "social23",
  "cfq_r_y_24" = "social24", "cfq_r_y_25" = "social25",
  "cfq_r_y_26" = "social26", "cfq_r_y_27" = "body27", "cfq_r_y_28" = "body28",
  "cfq_r_y_29" = "body29", "cfq_r_y_30" = "treat30", "cfq_r_y_31" = "resp31",
  "cfq_r_y_32" = "resp32", "cfq_r_y_33" = "resp33", "cfq_r_y_34" = "resp34",
  "cfq_r_y_35" = "digest35 "
)
cfqr_children_6_11 <-
  setNames(names(cfqr_children_6_11), cfqr_children_6_11)
cfqr_teens_adults <- list(
  "cfq_r_a_1" = "phys1", "cfq_r_a_2" = "phys2", "cfq_r_a_3" = "phys3",
  "cfq_r_a_4" = "phys4", "cfq_r_a_5" = "phys5", "cfq_r_a_6" = "vital6",
  "cfq_r_a_7" = "emot7", "cfq_r_a_8" = "emot8", "cfq_r_a_9" = "vital9",
  "cfq_r_a_10" = "vital10", "cfq_r_a_11" = "vital11", "cfq_r_a_12" = "emot12",
  "cfq_r_a_13" = "phys13", "cfq_r_a_14" = "eat14", "cfq_r_a_15" = "treat15",
  "cfq_r_a_16" = "treat16", "cfq_r_a_17" = "treat17", "cfq_r_a_18" = "health18",
  "cfq_r_a_19" = "phys19", "cfq_r_a_20" = "phys20", "cfq_r_a_21" = "eat21",
  "cfq_r_a_22" = "social22", "cfq_r_a_23" = "social23", "cfq_r_a_24" = "body24",
  "cfq_r_a_25" = "body25", "cfq_r_a_26" = "body26", "cfq_r_a_27" = "social27",
  "cfq_r_a_28" = "social28", "cfq_r_a_29" = "social29",
  "cfq_r_a_30" = "social30", "cfq_r_a_31" = "emot31", "cfq_r_a_32" = "health32",
  "cfq_r_a_33" = "emot33", "cfq_r_a_34" = "health34", "cfq_r_a_35" = "role35",
  "cfq_r_a_36" = "role36", "cfq_r_a_37" = "role37", "cfq_r_a_38" = "role38",
  "cfq_r_a_39" = "weight39", "cfq_r_a_40" = "resp40", "cfq_r_a_41" = "resp41",
  "cfq_r_a_42" = "resp42", "cfq_r_a_43" = "resp43", "cfq_r_a_44" = "resp44",
  "cfq_r_a_45" = "resp45", "cfq_r_a_46" = "resp46", "cfq_r_a_47" = "digest47",
  "cfq_r_a_48" = "digest48", "cfq_r_a_49" = "digest49", "cfq_r_a_50" = "eat50"
)
cfqr_teens_adults <-
  setNames(names(cfqr_teens_adults), cfqr_teens_adults)
```