### 2025-02-11 Angela Metcalf/quantifying_cl_cf_carriers.qmd
CF carrier interim power

---
title: "Quantifying Sweat Chloride in CF Carriers"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(arsenal)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers")
```

```{r data cleaning}
#| message: false
# Import from REDCap
unlockREDCap(c(rcon = "Sweat Electrolytes in CF Carriers 24-0691"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
df <- exportRecordsTyped(rcon)
# Format
df <- df %>%
  mutate(
    `Cl Right` = as.numeric(cl_right), `Cl Left` = as.numeric(cl_left),
    `Na Right` = as.numeric(na_right), `Na Left` = as.numeric(na_left)
  )
df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")])
df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")])
```

# Table 1

```{r}
t1 = tableby(carrier~`Mean Cl` +`Mean Na`,data = df )
summary(t1)
```

# Plots

```{r}
ggplot(df, aes(x = carrier, y = `Mean Cl`)) +
  geom_boxplot()
```