### 2025-02-12 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
clean up SSwAB-IT tables

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(gt)
library(gtsummary)
library(caret)
library(eulerr)
```

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Exclude
exclude <- c("01-030_W", "01-038_W", "02-0038", "02-0039", "04-001")
# Combine collection results
df <- full_join(db_cos_collection_results, db_sos_collection_results)
df <- full_join(df,
  db_selfobtained_sample[, c("record_id", "who_collected_the_sos")],
  by = join_by(record_id)
)
# Add dates to demographics
db_demographics <- left_join(db_demographics,
  db_clinic_obtained_values %>% select(record_id, cos_collection_date),
  by = join_by(record_id)
)
db_demographics$age_at_cos <-
  as.numeric(difftime(db_demographics$cos_collection_date,
    db_demographics$dob,
    units = "days"
  )) / 365.25
# 1-5, 6-11, 12-17 and 18+ years
db_demographics$age_group <- cut(db_demographics$age_at_cos,
  c(-Inf, 6, 12, 18, Inf),
  right = F
)
# Format
df <- df %>%
  select(
    record_id, who_collected_cos, who_collected_the_sos,
    staphylococcus_aureus_cos:other_result_cos,
    staphylococcus_aureus_sos:other_result_sos
  ) %>%
  filter(!record_id %in% exclude)
# Remove empty demographics rows
db_demographics <- db_demographics %>%
  filter(
    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
    !record_id %in% exclude
  )
# Excluded participants
db_medications <- db_medications %>% filter(!record_id %in% exclude)
# Combine race and ethnicity columns
races <- c(
  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
  "Black or African American", "White", "Unknown", "Other"
)
ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
db_demographics$Race <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("race", names(r))] == "Checked")
  if (length(w) == 1) {
    return(races[w])
  } else if (length(w) > 1) {
    return("Multiple Races")
  } else if (length(w) == 0) {
    return(NA)
  }
})
db_demographics$Race <- factor(db_demographics$Race,
  levels = c(
    "White", "Other", "Multiple Races", "American Indian or Alaskan Native"
  )
)
db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("ethnicity", names(r))] == "Checked")
  if (length(w) == 1) {
    return(ethnicities[w])
  } else if (length(w) > 1) {
    return(paste(ethnicities[w], sep = "/"))
  } else if (length(w) == 0) {
    return(NA)
  }
})
db_demographics$Ethnicity <- factor(db_demographics$Ethnicity,
  levels = c("Non-Hispanic", "Hispanic", "Unknown/Not Reported")
)
# Drop empty levels
db_demographics <- droplevels(db_demographics)
# Add groups to other dataframes
db_medications <- left_join(db_medications,
  db_demographics %>% select(record_id, group),
  by = join_by(record_id)
)
# Bug list
bugs <- c(
  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
  "other"
)
bug_names <- c(
  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
)
# Convert yes/no to positive/negative
df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
# Create overall columns (any bug detected in sample)
df$overall_cos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_cos")] == "Positive")
})
df$overall_cos <- factor(df$overall_cos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
df$overall_sos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_sos")] == "Positive")
})
df$overall_sos <- factor(df$overall_sos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
# Add overall to list
bugs <- c("overall", bugs)
bug_names <- c("Across all bugs", bug_names)
bug_list <- as.list(bug_names)
names(bug_list) <- bugs
# Add groups to dataframe
df <- left_join(df, db_demographics %>% select(record_id, group, age_group),
  by = join_by(record_id)
)
df <- left_join(df, db_medications, by = join_by(record_id, group))
# Feasibility
db_feasibility <- db_feasibility %>% filter(!record_id %in% exclude)
# Drop unused levels
df$who_collected_cos <- droplevels(df$who_collected_cos)
# Shorter group labels for cleaner tables
levels(df$group) <- c("COS/SOS/IP", "COS/SOS/DP", "SOS/COS/IP", "SOS/COS/DP")
levels(db_demographics$group) <-
  c("COS/SOS/IP", "COS/SOS/DP", "SOS/COS/IP", "SOS/COS/DP")
levels(db_medications$group) <-
  c("COS/SOS/IP", "COS/SOS/DP", "SOS/COS/IP", "SOS/COS/DP")
```

# Data cleaning

Participants 01-030_W, 01-038_W, 02-0038, 02-0039, and 04-001 were excluded from all analyses. Participants missing either the COS or SOS sample were excluded from agreement analyses, but retained in feasibility and other analyses.

# Demographics

## By randomization group

```{r}
db_demographics %>%
  select(group, age_at_cos, age_group, sex, Race, Ethnicity) %>%
  tbl_summary(
    by = group, missing_text = "Missing",
    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
    label = list(
      "sex" = "Sex", "age_at_cos" = "Age at COS", "age_group" = "Age Group"
    )
  ) %>%
  add_overall() %>%
  add_p(
    test = list(c("Race", "Ethnicity") ~ "fisher.test"),
    test.args = age_group ~ list(workspace = 2e9)
  ) %>%
  separate_p_footnotes() %>%
  modify_footnote(everything() ~ NA) %>%
  as_gt()
```

# Medications

```{r}
db_medications %>%
  select(group, azithromycin:please_specify_mucolytic) %>%
  select(
    -inhaled_colisitin_status, -inhaled_vancomycin_status,
    -other_inhaled_status, -please_specify_mucolytic
  ) %>%
  tbl_summary(by = group, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes() %>%
  modify_footnote(everything() ~ NA) %>%
  as_gt()
```

# Acceptability and feasibility

```{r}
db_feasibility %>%
  select(
    previous_home_sample:if_no_why_not___4, -barrier_other,
    -barriers___5, -if_no_why_not___4
  ) %>%
  tbl_summary(
    missing_text = "Missing",
    type = all_dichotomous() ~ "categorical"
  )
```

# Agreement

For these agreement analyses, we reported accuracy metrics that include all samples. Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative). Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement. A Cohen's kappa of NaN or NA indicates a lack of variability between the two tests.

| Cohen's Kappa | Interpretation         |
|---------------|------------------------|
| 0             | No agreement           |
| 0.10 - 0.20   | Slight agreement       |
| 0.21 - 0.40   | Fair agreement         |
| 0.41 - 0.60   | Moderate agreement     |
| 0.61 - 0.80   | Substantial agreement  |
| 0.81 - 0.99   | Near perfect agreement |
| 1             | Perfect agreement      |

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  colnames(d) <- c("COS", "SOS")
  d <- d[complete.cases(d), ]
  # Agreement metrics
  c <- confusionMatrix(d$SOS,
    reference = d$COS,
    positive = "Positive", mode = "everything",
    dnn = c("SOS", "COS")
  )
  # Print
  cat("\n")
  cat("\n")
  cat("####", "2x2 Table")
  cat("\n")
  cat("\n")
  d %>%
    tbl_cross(percent = "cell") %>%
    print()
  cat("\n")
  cat("\n")
  cat("####", "2x2 Table (With CIs)")
  cat("\n")
  cat("\n")
  d %>%
    tbl_summary(by = COS, percent = "cell") %>%
    add_ci(method = list(all_categorical() ~ "wald")) %>%
    print()
  cat("\n")
  cat("\n")
  cat("Overall accuracy (both negative/negative and positive/positive) was ",
    scales::percent(c$overall["Accuracy"], accuracy = 0.1), " (95% CI: ",
    scales::percent(c$overall["AccuracyLower"], accuracy = 0.1), " - ",
    scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1), ").",
    sep = ""
  )
  cat("\n")
  cat("\n")
  cat("Kappa was ", round(c$overall["Kappa"], 3), ".", sep = "")
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of negative results")
  cat("\n")
  cat("\n")
  print(plot(euler(d == "Negative"),
    quantities = list(cex = 3),
    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
  ))
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of positive results")
  cat("\n")
  cat("\n")
  print(plot(euler(d == "Positive"),
    quantities = list(cex = 3),
    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
  ))
  cat("\n")
  cat("\n")
}
agreement_test <- function(bug, group = "group", data = df) {
  # Prepare data
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos, group)]
  d <- d[complete.cases(d), ]
  colnames(d) <- c("COS", "SOS", "Group")
  # Agreement and detailed agreement
  d$agree <- d$COS == d$SOS
  d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
  d$detailed_agree <- factor(d$detailed_agree,
    levels = c(
      "COS Negative/SOS Negative", "COS Negative/SOS Positive",
      "COS Positive/SOS Negative", "COS Positive/SOS Positive"
    ),
    labels = c("COS-/SOS-", "COS-/SOS+", "COS+/SOS-", "COS+/SOS+")
  )
  # Print
  cat("\n")
  cat("\n")
  d %>%
    select(agree, detailed_agree, Group) %>%
    tbl_summary(by = Group, missing_text = "Missing", label = list(
      agree ~ "Agree (+/+ and -/-)",
      detailed_agree ~ "Detailed Agreement"
    )) %>%
    add_overall() %>%
    add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e9)) %>%
    separate_p_footnotes() %>%
    modify_footnote(everything() ~ NA) %>%
    print()
  cat("\n")
  cat("\n")
}
```

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  print(b)
  cat("\n")
  cat("\n")
  cat("##", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("###", "Overall agreement")
  cat("\n")
  cat("\n")
  agreement(b)
  cat("\n")
  cat("\n")
  cat("###", "Agreement by group")
  cat("\n")
  cat("\n")
  agreement_test(b)
  cat("\n")
  cat("\n")
}))
```

# Sub-group comparisons

## Age group

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by age group")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
  cat("\n")
}))
```

## Sex

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by gender assigned at birth")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
  cat("\n")
}))
```

## Azithromycin

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by Azithromycin status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "azithromycin")
  cat("\n")
  cat("\n")
}))
```

## Inhaled Tobramycin

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by inhaled Tobramycin status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "inhaled_tobramycin")
  cat("\n")
  cat("\n")
}))
```

## Inhaled Aztreonam

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by inhaled Aztreonam status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "inhaled_aztreonam")
  cat("\n")
  cat("\n")
}))
```

## Inhaled Colisitin

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by inhaled Colisitin status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "inhaled_colisitin")
  cat("\n")
  cat("\n")
}))
```

## Inhaled Vancomycin

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by inhaled Vancomycin status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "inhaled_vancomycin")
  cat("\n")
  cat("\n")
}))
```

## Other inhaled antibiotics

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by other inhaled antibiotics status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "other_inhaled_antibiotics")
  cat("\n")
  cat("\n")
}))
```

## Chronic oral antibiotic

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by chronic oral antibiotic status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "chronic_oral_antibiotic")
  cat("\n")
  cat("\n")
}))
```

## CFTR modulator status

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by CFTR modulator status")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "cftr_modulator")
  cat("\n")
  cat("\n")
}))
```

## Who collected the COS

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by who collected the COS")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "who_collected_cos")
  cat("\n")
  cat("\n")
}))
```

## Who collected the SOS

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("\n")
  cat("####", "Agreement by who collected the SOS")
  cat("\n")
  cat("\n")
  agreement_test(b, group = "who_collected_the_sos")
  cat("\n")
  cat("\n")
}))
```