### 2025-02-26 Angela Metcalf/quantifying_cl_cf_carriers.qmd
mostly file path changes

---
title: "Quantifying Sweat Chloride in CF Carriers"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/Documents/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(arsenal)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Angela Metcalf/Quantifying Sweat Chloride in CF Carriers")
```

```{r data cleaning}
#| message: false
# Import from REDCap
# unlockREDCap(c(rcon = "Sweat Electrolytes in CF Carriers 24-0691"),
#   keyring = "API_KEYs",
#   envir = 1,
#   url = "https://redcap.ucdenver.edu/api/"
# )
# df <- exportRecordsTyped(rcon)
# # Format
# df <- df %>%
#   mutate(
#     `Cl Right` = as.numeric(cl_right), `Cl Left` = as.numeric(cl_left),
#     `Na Right` = as.numeric(na_right), `Na Left` = as.numeric(na_left)
#   )
# df$`Mean Cl` <- rowMeans(df[, c("Cl Right", "Cl Left")])
# df$`Mean Na` <- rowMeans(df[, c("Na Right", "Na Left")])
```

# Confidence interval power calculations

```{r}
ns <- 35:45
power <- lapply(ns, function(n) {
  se <- qt(0.05 / 2, lower.tail = F, df = n - 1) * (12 / sqrt(n))
  return(c(
    "Confidence level" = 0.95, "Sample size (N)" = n,
    "Actual Distance from Mean to Limits" = se,
    "Standard deviation of sweat chloride" = 12
  ))
})
power <- do.call(rbind, power)
knitr::kable(power)
```