### 2025-03-04 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
joint model errors

---
title: "Using HbA1c and Fasting BG to Predict Development of CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(survival)
library(nlme)
library(JM)
library(dplyr)
library(tidyr)
knitr::opts_knit$set(root.dir = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry")
```

```{r data cleaning}
# Import data
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
# Demographic data
demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
  na.strings = c("", "N/A")
)
# Race
races <- c(
  "White", "Black or African American", "American Indian or Alaska Native",
  "Asian", "Native Hawaiian or Other Pacific Islander", "Some other race"
)
demo$Race <- apply(demo, 1, function(r) {
  w <- which(r[paste0("Race", 1:6)] == 1)
  if (length(w) == 0) {
    "Unknown"
  } else if (length(w) > 1) {
    "More than one race"
  } else {
    races[w]
  }
})
demo$Race <- factor(demo$Race,
  levels = c(
    "White", "Black or African American", "American Indian or Alaska Native",
    "Asian", "More than one race", "Native Hawaiian or Other Pacific Islander",
    "Some other race"
  ),
  labels = c(
    "White", "Black or African American", "Other race", "Other race",
    "More than one race", "Other race", "Other race"
  )
)
demo$Hispanicrace <- factor(demo$Hispanicrace,
  levels = 1:2,
  labels = c("Yes", "No")
)
# F508
demo$F508 <- factor(demo$F508,
  levels = 1:3,
  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
)
# Merge
annual <- left_join(
  annual,
  demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
  by = join_by(eDWID)
)
# CFRD by OGTT two hour
annual$cfrd_ogtt <- factor(annual$A_twohour_bloodglucose >= 200,
  levels = c(F, T), labels = c("CFRD-", "CFRD+")
)
# CFRD by registry definition
annual$cfrd_registry <- factor(annual$cfrd_status_annual,
  levels = c(1:5), labels = c("CFRD-", "CFRD-", "CFRD+", "CFRD-", "CFRD-")
)
# Assume that missing data for CFRD means "CFRD-" (some of these will get filled in
# later)
annual$cfrd_registry[is.na(annual$cfrd_registry)] <- "CFRD-"
annual$cfrd_ogtt[is.na(annual$cfrd_ogtt)] <- "CFRD-"
# Modulators
annual$A_Vx770 <- factor(annual$A_Vx770, levels = 0:1, labels = c("No", "Yes"))
annual$A_VX809comb <- factor(annual$A_VX809comb,
  levels = 0:1, labels = c("No", "Yes")
)
# Get age at first CFRD (by OGTT and registry)
annual <- annual %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_registry_age = first(na.omit(Age_YrEnd[cfrd_registry == "CFRD+"])),
    cfrd_ogtt_age = first(na.omit(Age_YrEnd[cfrd_ogtt == "CFRD+"]))
  )
# Fill in CFRD diagnoses
annual$cfrd_registry_age[is.na(annual$cfrd_registry_age)] <- Inf
annual$cfrd_ogtt_age[is.na(annual$cfrd_ogtt_age)] <- Inf
annual$cfrd_registry[annual$Age_YrEnd < annual$cfrd_registry_age] <- "CFRD-"
annual$cfrd_registry[annual$Age_YrEnd >= annual$cfrd_registry_age] <- "CFRD+"
annual$cfrd_ogtt[annual$Age_YrEnd < annual$cfrd_ogtt_age] <- "CFRD-"
annual$cfrd_ogtt[annual$Age_YrEnd >= annual$cfrd_ogtt_age] <- "CFRD+"
# Create start times
annual$Age_YrStart <- annual$Age_YrEnd - 1
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
# Labels
labels <- list(
  Age_YrStart = "Age at Year Start", Age_YrEnd = "Age at Year End",
  Hispanicrace = "Hispanic", A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_FEV1_pct_predicted = "FEV1 percent predicted", MutClass = "Mutation class",
  NumPulmExacerbation = "Admissions for PEx"
)
# Select columns
annual <- annual %>%
  select(
    eDWID, Race, Hispanicrace, Gender, F508, MutClass,
    ReviewYear, Age_YrStart, Age_YrEnd, A_Vx770, A_VX809comb, A_bmivalue,
    A_FEV1_pct_predicted, NumPulmExacerbation, A_IsOnEnzymes,
    A_fast_bloodglucose, A_hgba1c, cfrd_registry, cfrd_ogtt,
    A_twohour_bloodglucose, cfrd_status_annual
  )
```

# Joint models

```{r}
# Complete cases
model_df <- annual %>%
  select(eDWID, Gender, Age_YrStart, Age_YrEnd, A_hgba1c, cfrd_registry) %>%
  drop_na()
# Fit mixed model and Cox model separately
lme_fit_a1c <- lme(A_hgba1c ~ Age_YrStart + Gender,
  random = ~ Age_YrStart | eDWID,
  data = model_df
)
cox_fit <- coxph(
  Surv(
    time = Age_YrStart, time2 = Age_YrEnd,
    event = as.numeric(cfrd_registry == "CFRD+")
  ) ~ Gender,
  cluster = eDWID,
  data = model_df, x = T, model = T
)
# Combine
joint_fit <- jointModel(lme_fit_a1c, cox_fit,
  timeVar = "Age_YrStart", method = "piecewise-PH-aGH"
)
```

# By registry CFRD status

## Participant demographics

Categorical variables are from either the first year with CFRD or the most recent year for those without CFRD.

```{r results='asis'}
v1 <- registry %>%
  group_by(eDWID) %>%
  slice_max(ReviewYear)
t1 <- tableby(cfrd_registry ~ Race + Hispanicrace + Gender + F508 + MutClass +
  Age_YrEnd + A_bmipercentile + A_hgba1c, data = v1)
summary(t1, pfootnote = T)
```

## Kaplan-Meier plots

```{r}
# Plot
survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~ 1,
  data = registry
) %>%
  ggsurvfit() +
  labs(
    x = "Age",
    y = "Overall CFRD- probability"
  ) +
  add_confidence_interval() +
  add_risktable()
# Median "survival" time
survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_registry)) ~ 1,
  data = registry
) %>%
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median time to CFRD+ (95% CI)**"
  )
```

## Cox PH model

```{r}
cph <- coxph(
  update(f, Surv(time = Age_YrStart, time2 = Age_YrEnd, event = as.numeric(cfrd_registry) == 2) ~ .),
  data = registry, cluster = eDWID
)
cph %>% tbl_regression(exponentiate = T)
plot(cox.zph(cph), resid = F)
cox.zph(cph)
ggcoef_table(cph,
  exponentiate = T, show_p_values = F, add_reference_rows = T, signif_stars = F,
  colour = F
)

d <- model.frame(cph)
d$pred <- predict(cph, type = "survival")
```

# By CFRD status based on 2 hour OGTT >= 200

## Participant demographics

Categorical variables are from either the first year with CFRD or the most recent year for those without CFRD.

```{r results='asis'}
v1 <- ogtt %>%
  group_by(eDWID) %>%
  slice_max(ReviewYear)
t1 <- tableby(cfrd_ogtt ~ Race + Hispanicrace + Gender + F508 + MutClass +
  Age_YrEnd + A_bmipercentile + A_hgba1c, data = v1)
summary(t1, pfootnote = T)
```

## Kaplan-Meier plots

```{r}
# Plot
survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~ 1,
  data = ogtt
) %>%
  ggsurvfit() +
  labs(
    x = "Age",
    y = "Overall CFRD- probability"
  ) +
  add_confidence_interval() +
  add_risktable()
# Median "survival" time
survfit2(Surv(Age_YrStart, Age_YrEnd, as.numeric(cfrd_ogtt)) ~ 1,
  data = ogtt
) %>%
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median time to CFRD+ (95% CI)**"
  )
```

## Cox PH model

```{r}
cph <- coxph(
  update(f, Surv(time = Age_YrStart, time2 = Age_YrEnd, event = as.numeric(cfrd_ogtt)) ~ .),
  data = ogtt, cluster = eDWID
)
cph %>% tbl_regression(exponentiate = T)
plot(cox.zph(cph), resid = F)
cox.zph(cph)
ggcoef_table(cph,
  exponentiate = T, show_p_values = F, add_reference_rows = T, signif_stars = F,
  colour = F
)
```