### 2025-05-02 Emily Muther/mental_health_screening_psc.qmd
file paths

---
title: "PSC Screening in Cystic Fibrosis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Emily Muther",
  "Linux" = "/home/timvigers/OneDrive"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import
df <- read_excel("./Muther Research Group - General/PSC manuscript 2025/PSC_9-20-19 thru 4-18-25_pulled 4-21-25.xlsx")
# Fill down MRN, select columns, remove duplicates
df <- df %>%
  fill(`Pat Mrn Id`, .direction = "down") %>%
  select(
    `Pat Mrn Id`, `Payor Name`, `AGE AT CONTACT`, `Internalizing subscale`,
    `Attention subscale`, `Externalizing subscale`, Score
  ) %>%
  distinct(`Pat Mrn Id`, `AGE AT CONTACT`, .keep_all = T)
```

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets
```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets
```