### 2025-05-02 Emily Muther/mental_health_screening_psc.qmd
PSC EDA for Emily

---
title: "PSC Screening in Cystic Fibrosis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(arsenal)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Emily Muther",
  "Linux" = "/home/timvigers/OneDrive"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import
df <- read_excel("./Muther Research Group - General/PSC manuscript 2025/PSC_9-20-19 thru 4-18-25_pulled 4-21-25.xlsx")
# Fill down MRN, select columns, remove duplicates
df <- df %>%
  fill(`Pat Mrn Id`, .direction = "down") %>%
  select(
    `Pat Mrn Id`, `Payor Name`, `AGE AT CONTACT`, `Internalizing subscale`,
    `Attention subscale`, `Externalizing subscale`, Score
  ) %>%
  # pivot_wider(names_from = `Payor Name`,values_from = `Payor Name`)
  distinct(`Pat Mrn Id`, `AGE AT CONTACT`, .keep_all = T)
# Convert to numeric
df$`AGE AT CONTACT` <- as.numeric(df$`AGE AT CONTACT`)
df$`Internalizing subscale` <- as.numeric(df$`Internalizing subscale`)
df$`Attention subscale` <- as.numeric(df$`Attention subscale`)
df$`Externalizing subscale` <- as.numeric(df$`Externalizing subscale`)
df$Score <- as.numeric(df$Score)
# Make age group
df$`Age Group` <- cut(df$`AGE AT CONTACT`,
  breaks = c(-Inf, 6, 10, Inf),
  labels = c("Age < 6", "Age >= 6 and < 10", "Age >= 10"), right = F
)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct")
  )
```

# Insurance providers

MEDICAID = Public
MEDICAID NEW MEXICO = Public
MEDICAID WYOMING = Public
UNITED HEALTHCARE = Private
CIGNA = Private
ANTHEM BCBS = Private
ROCKY MOUNTAIN HEALTH PLAN = ?
AETNA = Private
MEDICAID OTHER STATES = Public
TRICARE = Tricare
CICP = ?
CHP+COLORADO ACCESS = ?
CU GME HEALTH = ?
MIDLANDS CHOICE = ?
BRIGHT HEALTH = ?
FIRST HEALTH GROUP = ?
ROCKY MOUNTAIN MEDICAID = Public
KAISER = Kaiser
MEDICAID NEBRASKA = Public
MEDICAID OKLAHOMA = Public
EBMS = ?
COFINITY = ?
SIMPLIFIED BENEFITS ADMINISTRATORS = ?
FRIDAY HEALTH PLANS = ?
SELECTHEALTH = INC. = Private?

# Score summary

## At first visit

```{r results='asis'}
first_visit <- df %>%
  group_by(`Pat Mrn Id`) %>%
  slice_min(`AGE AT CONTACT`)
t1 <- tableby(
  `Age Group` ~ `Internalizing subscale` + `Attention subscale` +
    `Externalizing subscale` + Score,
  data = first_visit
)
summary(t1, pfootnote = T)
```

## All visits

The table below treats multiple visits from the same participant as independent.

```{r results='asis'}
t1 <- tableby(
  `Age Group` ~ `Internalizing subscale` + `Attention subscale` +
    `Externalizing subscale` + Score,
  data = df
)
summary(t1, pfootnote = T)
```

# Figures

## Histograms

```{r}
hist(df$`Internalizing subscale`)
hist(df$`Attention subscale`)
hist(df$`Externalizing subscale`)
hist(df$Score)
```

## Scores by age

```{r}
#| warning: false
ggplot(df, aes(x = `AGE AT CONTACT`, y = `Internalizing subscale`)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))

ggplot(df, aes(x = `AGE AT CONTACT`, y = `Attention subscale`)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))

ggplot(df, aes(x = `AGE AT CONTACT`, y = `Externalizing subscale`)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))

ggplot(df, aes(x = `AGE AT CONTACT`, y = Score)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))
```