### 2025-05-15 Emily Muther/mental_health_screening_psc.qmd

---
title: "PSC Screening in Cystic Fibrosis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(arsenal)
library(GLMMadaptive)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Emily Muther",
  "Linux" = "/home/timvigers/OneDrive"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import
df <- read_excel("./Muther Research Group - General/PSC manuscript 2025/PSC_9-20-19 thru 4-18-25_pulled 4-21-25.xlsx")
# Classify insurance per Emily
df$Insurance <- factor(df$`Payor Name`,
  levels = c(
    "MEDICAID", "MEDICAID NEW MEXICO", "MEDICAID WYOMING", "UNITED HEALTHCARE",
    "CIGNA", "ANTHEM BCBS", "ROCKY MOUNTAIN HEALTH PLAN", "AETNA",
    "MEDICAID OTHER STATES", "TRICARE", "CICP", "CHP+COLORADO ACCESS",
    "CU GME HEALTH", "MIDLANDS CHOICE", "BRIGHT HEALTH", "FIRST HEALTH GROUP",
    "ROCKY MOUNTAIN MEDICAID", "KAISER", "MEDICAID NEBRASKA",
    "MEDICAID OKLAHOMA", "EBMS", "COFINITY", "SIMPLIFIED BENEFITS ADMINISTRATORS",
    "FRIDAY HEALTH PLANS", "SELECTHEALTH, INC."
  ),
  labels = c(
    "Public", "Public", "Public", "Private", "Private", "Private", "Public",
    "Private", "Public", "Private", "Public", "Public", "Private", "Private",
    "Private", "Private", "Public", "Private", "Public", "Public", "Private",
    "Private", "Private", "Private", "Private"
  )
)
# Fill down MRN and select columns
df <- df %>%
  fill(`Pat Mrn Id`, .direction = "down") %>%
  select(
    `Pat Mrn Id`, Insurance, `AGE AT CONTACT`, `Internalizing subscale`,
    `Attention subscale`, `Externalizing subscale`, Score
  ) %>%
  distinct() %>%
  pivot_wider(names_from = Insurance, values_from = Insurance) %>%
  unite("Insurance", Public, Private, `NA`, remove = T, na.rm = T)
df$Insurance <- factor(df$Insurance,
  levels = c("", "Private", "Public", "Public_Private"),
  labels = c("None", "Private", "Public", "Both")
)
# Convert to numeric
df$`AGE AT CONTACT` <- as.numeric(df$`AGE AT CONTACT`)
df$`Internalizing subscale` <- as.numeric(df$`Internalizing subscale`)
df$`Attention subscale` <- as.numeric(df$`Attention subscale`)
df$`Externalizing subscale` <- as.numeric(df$`Externalizing subscale`)
df$Score <- as.numeric(df$Score)
# Make age group
df$`Age Group` <- cut(df$`AGE AT CONTACT`,
  breaks = c(-Inf, 6, 10, Inf),
  labels = c("Age < 6", "Age >= 6 and < 10", "Age >= 10"), right = F
)
# Make variables for 0 scores on each scale
df$`Internalizing subscale symptoms` <- factor(df$`Internalizing subscale` == 0,
  levels = c(F, T),
  labels = c(
    "No internalizing symptoms (score = 0)",
    "Any internalizing symptoms (score > 0)"
  )
)
df$`Attention subscale symptoms` <- factor(df$`Attention subscale` == 0,
  levels = c(F, T),
  labels = c(
    "No attention symptoms (score = 0)",
    "Any attention symptoms (score > 0)"
  )
)
df$`Externalizing subscale symptoms` <- factor(df$`Externalizing subscale`==0,
  levels = c(F, T),
  labels = c(
    "No externalizing symptoms (score = 0)",
    "Any externalizing symptoms (score > 0)"
  )
)
df$`Any symptoms` <- factor(df$Score == 0,
  levels = c(F, T),
  labels = c(
    "No symptoms (total score = 0)",
    "Any symptoms (total score > 0)"
  )
)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "medianq1q3", "range"),
    cat.stats = c("N", "Nmiss", "countpct")
  )
# Basic formula
f <- ~ Insurance + `Internalizing subscale` + `Internalizing subscale symptoms` +
  `Attention subscale` + `Attention subscale symptoms` +
  `Externalizing subscale` + `Externalizing subscale symptoms` +
  Score + `Any symptoms`
```

# Score summary at first visit

## By age group

```{r results='asis'}
first_visit <- df %>%
  group_by(`Pat Mrn Id`) %>%
  slice_min(`AGE AT CONTACT`)
t1 <- tableby(update(f, `Age Group` ~ .), data = first_visit)
summary(t1, pfootnote = T)
```

## By insurance type

```{r results='asis'}
first_visit <- df %>%
  group_by(`Pat Mrn Id`) %>%
  slice_min(`AGE AT CONTACT`)
t1 <- tableby(update(f, Insurance ~ .), data = first_visit)
summary(t1, pfootnote = T)
```

# Figures

## Histograms

```{r}
hist(df$`Internalizing subscale`)
hist(df$`Attention subscale`)
hist(df$`Externalizing subscale`)
hist(df$Score)
```

## Scores by age

```{r}
#| warning: false
ggplot(df, aes(x = `AGE AT CONTACT`, y = `Internalizing subscale`)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))

ggplot(df, aes(x = `AGE AT CONTACT`, y = `Attention subscale`)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))

ggplot(df, aes(x = `AGE AT CONTACT`, y = `Externalizing subscale`)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))

ggplot(df, aes(x = `AGE AT CONTACT`, y = Score)) +
  geom_point() +
  geom_line(aes(group = `Pat Mrn Id`))
```

# Mixed models

## Total score

```{r}
# First determine whether we need Poisson or negative binomial
pm1 <- mixed_model(Score ~ `AGE AT CONTACT`,
  random = ~ 1 | `Pat Mrn Id`, data = df,
  family = zi.poisson(), zi_fixed = ~`AGE AT CONTACT`
)
nbm1 <- mixed_model(Score ~ `AGE AT CONTACT`,
  random = ~ 1 | `Pat Mrn Id`, data = df,
  family = zi.negative.binomial(), zi_fixed = ~`AGE AT CONTACT`
)
# anova(pm1, nbm1, test = FALSE) - NB better by AIC and BIC
summary(nbm1)
```

# Questions for Emily

- If payor is blank, does that indicate no insurance? Or is it missing?
- Should we ask Elin for gender and lung function data? Not sure how much work that would be for her.