### 2025-05-27 Christine Chan/EnVision CF/cfrd_roc.qmd
Envision CFRD all thresholds

---
title: "Using CGM to Predict CFRD in the Envision Cohort"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
library(pROC)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import cleaned dataset
df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
# Limit to those with CGM data (based on mean glucose)
df <- df %>% filter(!is.na(mean_sensor))
```

- **All tables and analyses below treat CGM wears as independent**, but we will need to think about how to handle this in future analyses.

# Participant demographics

- In the cleaned EnVision dataset we have a total of `r nrow(df)` CGMs from `r length(unique(df$study_id))` unique individuals. Of these CGM wears, `r sum(is.na(df$Diagnosis))` are missing a diagnosis due to missing OGTT timepoints.
- BMI percentile was calculated with age = 20 for adults.

```{r results='asis'}
t1 <- tableby(
  CFRD ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t1, pfootnote = T)
```

# ROC Analyses

## Mean sensor glucose

```{r}
r <- roc(CFRD ~ mean_sensor, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## Percent time spent >140 mg/dL

```{r}
r <- roc(CFRD ~ percent_time_over_140, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## Percent time spent >180 mg/dL

```{r}
r <- roc(CFRD ~ percent_time_over_180, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## GMI

```{r}
r <- roc(CFRD ~ gmi, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## SD

```{r}
r <- roc(CFRD ~ standard_deviation, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## CV

```{r}
r <- roc(CFRD ~ cv, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

# Questions

- Are we including fasting glucose >= 126 in the definition of CFRD for this analysis?