### 2025-06-09 Christine Chan/EnVision CF/cfrd_roc.qmd
updated EnVision ROC curves

---
title: "Using CGM to Predict CFRD in the Envision Cohort"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(arsenal)
library(pROC)
library(gt)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
# Import cleaned dataset
load("./Data_Clean/envision_analysis_dataset.RData")
df <- final_df
# Limit to those with CGM data (based on mean glucose) and OGTT diagnosis
df <- df %>% filter(!is.na(mean_sensor))
# Convert CV units to match Scully et al.
df$cv <- df$cv * 100
```

- **All tables and analyses below treat CGM wears as independent**, but we will need to think about how to handle this in future analyses.

# Participant demographics

- In the cleaned EnVision dataset we have a total of `r nrow(df)` CGMs from `r length(unique(df$study_id))` unique individuals. Of these CGM wears, `r sum(is.na(df$Diagnosis))` were missing a diagnosis due to missing OGTT timepoints. Those without a diagnosis were excluded from all analyses.
- BMI percentile was calculated with age = 20 for adults.

```{r}
df <- df %>% filter(!is.na(Diagnosis))
```

## By OGTT categories

### Detailed OGTT category

```{r results='asis'}
t1 <- tableby(
  Diagnosis ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t1, pfootnote = T)
```

### General OGTT category

```{r results='asis'}
t2 <- tableby(
  Dysglycemia ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t2, pfootnote = T)
```

### CFRD status

```{r results='asis'}
t3 <- tableby(
  CFRD ~ redcap_data_access_group + sex + origin_race + ethnicity +
    pancreatic_status + CFTR + fev1 + fvc + bmi_perc + a1c_result,
  data = df
)
summary(t3, pfootnote = T)
```

# ROC analyses

## Mean sensor glucose

```{r}
r <- roc(CFRD ~ mean_sensor, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## Percent time spent >140 mg/dL

```{r}
r <- roc(CFRD ~ percent_time_over_140, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## Percent time spent >180 mg/dL

```{r}
r <- roc(CFRD ~ percent_time_over_180, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## GMI

```{r}
r <- roc(CFRD ~ gmi, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## SD

```{r}
r <- roc(CFRD ~ standard_deviation, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## CV

```{r}
r <- roc(CFRD ~ cv, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

## HbA1c

```{r}
r <- roc(CFRD ~ a1c_result, data = df, quiet = T)
plot(r, print.auc = T, print.thres = T, legacy.axes = T)
gt(coords(r, "all", ret = "all")) %>% fmt_number(decimals = 3)
```

# Pulmonary function comparisons

## By OGTT categories

### Detailed OGTT category

```{r results='asis'}
t1 <- tableby(Diagnosis ~ fev1 + fvc, data = df)
summary(t1, pfootnote = T)
```

### General OGTT category

```{r results='asis'}
t2 <- tableby(Dysglycemia ~ fev1 + fvc, data = df)
summary(t2, pfootnote = T)
```

### CFRD status

```{r results='asis'}
t3 <- tableby(CFRD ~ fev1 + fvc, data = df)
summary(t3, pfootnote = T)
```

## By CGM category

```{r}
df$scully_ag <- cut(df$mean_sensor, c(-Inf, 112, Inf))
df$our_ag <- cut(df$mean_sensor, c(-Inf, 101, Inf))
df$scully_140 <- cut(df$percent_time_over_140, c(-Inf, 17.5, Inf))
df$our_140 <- cut(df$percent_time_over_140, c(-Inf, 11.3, Inf))
df$scully_180 <- cut(df$percent_time_over_180, c(-Inf, 3.4, Inf))
df$our_180 <- cut(df$percent_time_over_180, c(-Inf, 3.1, Inf))
df$scully_sd <- cut(df$standard_deviation, c(-Inf, 30.8, Inf))
df$our_sd <- cut(df$standard_deviation, c(-Inf, 28.8, Inf))
df$scully_cv <- cut(df$cv, c(-Inf, 30.6, Inf))
df$our_cv <- cut(df$cv, c(-Inf, 27.7, Inf))
```

### Mean glucose

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_ag ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_ag ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### Percent time spent >140 mg/dL

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_140 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_140 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### Percent time spent >180 mg/dL

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_180 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_180 ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### SD

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_sd ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_sd ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

### CV

#### Scully et al. cutoff

```{r results='asis'}
tb <- tableby(scully_cv ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

#### Our cutoff

```{r results='asis'}
tb <- tableby(our_cv ~ fev1 + fvc, data = df)
summary(tb, pfootnote = T)
```

# Figure

In the figure below, red lines indicate the cutoffs identified in Scully et al. and purple lines indicate the cutoffs identified in EnVision.

```{r}
#| message: false
# Dotplot
box_p1 <- ggplot(
  df,
  aes(x = Dysglycemia, y = percent_time_over_140, colour = Dysglycemia)
) +
  geom_jitter() +
  geom_hline(yintercept = 17.5, color = "red") +
  geom_hline(yintercept = 11.3, color = "purple") +
  ggtitle("% Time Over 140 (mg/dL)") +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(), axis.title.y = element_blank()
  ) +
  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
box_p2 <- ggplot(
  df,
  aes(x = Dysglycemia, y = percent_time_over_180, colour = Dysglycemia)
) +
  geom_jitter() +
  geom_hline(yintercept = 3.4, color = "red") +
  geom_hline(yintercept = 3.1, color = "purple") +
  ggtitle("% Time Over 180 (mg/dL)") +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(), axis.title.y = element_blank()
  ) +
  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
box_p3 <- ggplot(
  df,
  aes(x = Dysglycemia, y = mean_sensor, colour = Dysglycemia)
) +
  geom_jitter() +
  geom_hline(yintercept = 112, color = "red") +
  geom_hline(yintercept = 101, color = "purple") +
  ggtitle("Mean glucose (mg/dL)") +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(), axis.title.y = element_blank()
  ) +
  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
box_p4 <- ggplot(
  df,
  aes(x = Dysglycemia, y = cv, colour = Dysglycemia)
) +
  geom_jitter() +
  geom_hline(yintercept = 30.6, color = "red") +
  geom_hline(yintercept = 27.7, color = "purple") +
  ggtitle("Coefficient of Variation (%)") +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(), axis.title.y = element_blank()
  ) +
  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
box_p5 <- ggplot(
  df,
  aes(x = Dysglycemia, y = a1c_result, colour = Dysglycemia)
) +
  geom_jitter() +
  geom_hline(yintercept = 5.8, color = "red") +
  geom_hline(yintercept = 5.75, color = "purple") +
  ggtitle("HbA1c (%)") +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(), axis.title.y = element_blank()
  ) +
  scale_color_manual(values = c("#4DA8DA", "#80D8C3", "#FFD66B"))
box_p <- box_p1 + box_p2 + box_p4 + box_p3 + box_p5 + plot_layout(ncol = 5)
box_p
```