### 2025-06-15 Christine Chan/CF Registry/WNAR/wnar_presentation.qmd
WNAR slides done

---
title: "The Changing Landscape of Cystic Fibrosis Related Diabetes Risk Factors Between 2003-2018"
author:
  - name: Tim Vigers
    orcid: 0000-0001-7087-8570
    email: timothy.vigers@cuanschutz.edu
    affiliations:
      - name: University of Colorado Denver, Anschutz Medical Campus
date: "June 17, 2025"
date-format: long
format: 
  revealjs:
    scrollable: true
    embed-resources: true
editor: source
bibliography: wnar_references.bib
csl: american-medical-association.csl
---

```{r setup}
#| include: false
library(tidyverse)
library(survival)
library(survminer)
library(gtsummary)
library(ggstats)
knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry")
```

# Background

## The *CFTR* Gene

- Cystic fibrosis (CF) is an autosomal recessive disease caused by mutations in the cystic fibrosis transmembrane conductance regulator (*CFTR*) gene.[@grasemannCysticFibrosis2023]
- The CFTR protein is an ion channel that regulates electrolyte balance across cell membranes.[@stoltzOriginsCysticFibrosis2015]
- As of 2023, over 2000 *CFTR* variants have been described. Of these, we are certain that over 700 are disease-causing.[@grasemannCysticFibrosis2023]
- The *CFTR* gene is expressed in the upper and lower airways, **pancreas**, intestine, biliary tree, cervix, sweat glands, and other organ systems.[@grasemannCysticFibrosis2023]

## Cystic Fibrosis-Related Diabetes

- Because *CFTR* is highly expressed in the pancreas, *CFTR* dysfunction commonly leads to pancreatic insufficiency.[@moheetNewConceptsPathogenesis2022]
- The pathophysiology of CFRD is not fully understood, but we think that $\beta$-cell loss related to pancreas destruction is one of the major drivers of diabetes in people with CF.[@moheetNewConceptsPathogenesis2022]
- CFRD is the most common non-pulmonary complication in CF. A diagnosis before the age of 10 is unusual, but 50% of adults over 40 have CFRD.[@moheetNewConceptsPathogenesis2022]
- CFRD is associated with a decline in nutritional status, lung function, and early mortality. [@moranClinicalCareGuidelines2010]

## Advancements in Treatment

- *CFTR* modulator therapies have completely transformed CF care.
- Median predicted survival increased from 34 years in 2004 to 46 years in 2019.[@cysticfibrosisfoundationCysticFibrosisFoundation2024]
- We expect to see even greater gains with the approval of new highly effective *CFTR* modulator therapies. [@Bessonova731; @middletonElexacaftorTezacaftorIvacaftor2019]
- However, despite these advances CFRD remains prevalent and there have been few large-scale studies examining changes in the epidemiology of CFRD over time.[@moranClinicalCareGuidelines2010; @marshallEpidemiologyCysticFibrosisrelated2005; @adlerHyperglycemiaDeathCystic2011; @chamnanDiabetesDeterminantMortality2010]

# How Have CFRD Risk Factors Changed With the Introduction of Highly Effective *CFTR* Modulators?

## Known Risk Factors {.smaller}

- Earlier CFRD studies [@marshallEpidemiologyCysticFibrosisrelated2005; @moranCysticFibrosisRelated2009] found positive associations between CFRD risk and:
  - Age
  - Female sex
  - Race/ethnicity from marginalized groups
  - BMI
  - Genotype
  - Pancreatic insufficiency
  - Nutritional status
  - Liver disease
  - Worse pulmonary disease
  - Use of corticosteroids

## The CFF-PR

- These variables were also readily available in the United States Cystic Fibrosis Foundation Patient Registry (CFFPR) and have been collected consistently since 2003.
- The CFF-PR is national database that aims to collect information on everyone with CF who receive care in CF Foundation-accredited care centers.

## The CFF-PR

- In 2023 there were 33,288 people in the registry (about 95%).
- The study P.I. wanted to use annualized data for this analysis.
  - Easier to get permission for this because there are fewer PHI concerns than encounter-level data.
- Very few model selection issues (yay!)

# The Fun Part!

## New Diagnoses in 2018 {.scrolling}

```{r}
#| message: false
load("./Data_Clean/survival_dataset.RData")
annual %>%
  filter(ReviewYear == 2018) %>%
  select(
    CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_cat, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
  ) %>%
  tbl_summary(
    by = CFRD, missing_text = "Missing",
    type = list(A_corticosteroids1 ~ "categorical"),
    statistic = list(c(
      "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
      "A_FVC_pct_predicted"
    ) ~ "{mean} ({sd})")
  ) %>%
  add_overall()
```

## Kaplan-Meier Plot

Every person in the CFF-PR from 2003-2018

```{r}
#| fig-width: 16
#| fig-height: 9
# KM plot
fit <- survfit(
  Surv(time = tstop, event = as.numeric(CFRD) == 2) ~ age_group,
  data = annual %>%
    group_by(eDWID) %>%
    slice_max(ReviewYear)
)
# Visualize with survminer
ggsurvplot(fit,
  data = annual, risk.table = T, cumevents = T, palette = "simpsons",
  conf.int = T
) + xlab("Days From 2003-01-01")
```

## Data Setup {.smaller}

- Because we don't have an exact date of CFRD diagnosis, the only information we have is that the diagnosis occurred between January 1st of one year and January 1st of the next.
  - Participants had to have two consecutive rows with CFRD, but first instance was considered onset.
- Data are interval censored.
- Intervals are on a discrete grid (i.e. start and stop times are the same for all participants in a given year).
- Data include time-varying covariates such as age group, BMI category, etc.
- We categorized variables such as BMI and FEV1 mostly for the sake of interpretability for a clinical audience.
  - For BMI in particular though, we were not sure that the effect would be linear and categorization allowed for more flexible models.

## Data Setup

```{r}
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
annual %>%
  select(eDWID, ReviewYear, tstart, tstop, all_of(xvars), CFRD) %>%
  head(20) %>%
  gt::gt()
```

## Our First Model {.scrolling}

```{r}
#| fig-width: 12
#| fig-height: 9
# Formula for base model
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

## Proportional Hazards Assumption

```{r}
cox.zph(cox_res)
```

## Proportional Hazards Assumption

```{r}
plot(cox.zph(cox_res)[2], resid = F)
```

## Proportional Hazards Assumption

```{r}
plot(cox.zph(cox_res)[7], resid = F)
```

# What do we do when the proportional hazards assumption is violated?

## Two Good Options

- Joint modeling of longitudinal and survival data.[@hendersonJointModellingLongitudinal2000]
  - The gold standard, do this if you can.
  - Complex, requires somewhat specialized knowledge.
  - Ask Elizabeth for help with this.

- Time-dependent Cox models
  - More appropriate for external covariates (measures that vary independent of the failure time).
  - Or baseline covariates where the effect changes over time.[@zhangTimevaryingCovariatesCoefficients2018]

## Time Dependent Cox Model

- The hazard function for the Cox proportional hazards model with time dependent coefficients is[@therneauUsingTimeDependent2024]:

$$
\lambda(t)=\lambda_0(t)exp(\beta(t)X)
$$

- Where the proportional hazard assumption is[@therneauUsingTimeDependent2024]

$$
\beta(t)=\beta
$$

## Time Dependent Coefficients

- Another way to write this model:

$$
\lambda(t)=\lambda_0(t)exp(g(\beta,t)X)
$$

- Where $g(\beta,t)$ is a function of time that we can define.[@zhangTimevaryingCovariatesCoefficients2018]

## Time as a Step Function

```{r}
#| echo: true
data(cancer)
vet2 <- survSplit(Surv(time, status) ~ .,
  data = cancer, cut = c(90, 180), episode = "tgroup", id = "id"
)
vfit2 <- coxph(Surv(tstart, time, status) ~ age + sex + ph.ecog +
  pat.karno:strata(tgroup), data = vet2)
vfit2
```

# Applying This to Our CFRD Data

## The Code

```{r}
#| echo: true
annual_split <- survSplit(
  Surv(
    time = tstart, time2 = tstop,
    event = as.numeric(CFRD) == 2
  ) ~ .,
  data = annual, cut = c(2192, 4018), episode = "tgroup"
)
cox_res2 <- coxph(
  Surv(time = tstart, time2 = tstop, event) ~
    age_group:strata(tgroup) + Gender + Race + Hispanicrace + bmi_cat +
    MutClass:strata(tgroup) + A_IsOnEnzymes:strata(tgroup) +
    tube_feeding:strata(tgroup) + liver_disease:strata(tgroup) +
    ppFEV1_cat:strata(tgroup) + NumPulmExacerbation:strata(tgroup) +
    A_corticosteroids1:strata(tgroup),
  data = annual_split, cluster = eDWID
)
```

## The Results

```{r}
cox_res2 %>%
  tbl_regression(exponentiate = T)
```

## The Interpretation

![](homer.jpg)

# An Alternative Approach

## Eras of CFTR Modulators

- We split our data into three different eras: 2003-2008, 2009-2013, and 2014-2018.
  - Ivacaftor (VX-770) was approved January 2012 and lumacaftor/ivacaftor (combination VX-809) was approved July 2015.
  - Three eras of equal lengths of time, the first with no modulators, second with one modulator, and the third with combination therapy.

## Eras of CFTR Modulators

- Per the authors of the `survival` package, likelihood equations use only one copy of any subject, and the program picks out the correct row of data at each time.[@therneauUsingTimeDependent2024]
  - This is why we don't need to worry about correlated data even though subjects have multiple observations.
  - Not losing much information by splitting our data up.

## Eras of CFTR Modulators

```{r}
annual$time_period <- cut(annual$ReviewYear,
  breaks = c(-Inf, 2009, 2014, Inf), right = F
)
era1 <- coxph(f,
  data = annual[as.numeric(annual$time_period) == 1, ],
  cluster = eDWID
)
t1 <- era1 %>% tbl_regression(exponentiate = T)
era2 <- coxph(f,
  data = annual[as.numeric(annual$time_period) == 2, ],
  cluster = eDWID
)
t2 <- era2 %>% tbl_regression(exponentiate = T)
era3 <- coxph(f,
  data = annual[as.numeric(annual$time_period) == 3, ],
  cluster = eDWID
)
t3 <- era3 %>% tbl_regression(exponentiate = T)
tbl_merge(
  tbls = list(t1, t2, t3),
  tab_spanner = c("2003 - 2008", "2009 - 2013", "2014 - 2018")
)
```

## Eras of CFTR Modulators

```{r}
#| fig-width: 12
#| fig-height: 9
comb_plot <- ggcoef_compare(
  list("2003 - 2008" = era1, "2009 - 2013" = era2, "2014 - 2018" = era3),
  add_reference_rows = T, exponentiate = T, shape_guide = F, colour = F,
  type = "faceted"
)
comb_plot
```

# The Effect of *CFTR* Modulators

## Ivacaftor

```{r}
#| fig-width: 12
#| fig-height: 9
f_vx770 <- update(f, . ~ . + A_Vx770 - A_corticosteroids1 -
  NumPulmExacerbation - age_group - Hispanicrace)
vx770_cox_res <- coxph(f_vx770,
  data = annual[annual$ReviewYear >= 2011, ],
  cluster = eDWID
)
ggcoef_model(vx770_cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

## Lumacaftor/ivacaftor

```{r}
#| fig-width: 12
#| fig-height: 9
f_vx809comb <- update(f, . ~ . + A_VX809comb - A_corticosteroids1 -
  NumPulmExacerbation - age_group - Hispanicrace)
vx809comb_cox_res <- coxph(f_vx809comb,
  data = annual[annual$ReviewYear >= 2015, ],
  cluster = eDWID
)
ggcoef_model(vx809comb_cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

## Conclusions

- The middle age group is no longer at increased risk compared to the youngest.
- Female and Black people with CF remain at increased risk of developing CFRD.
- People with CF who are in the "undernourished" BMI category are no longer at increased risk, but those who are overweight are.

## Conclusions

- Other risk factors appear relatively stable over time.
- Ivcaftor may reduce the risk of CFRD, although combination therapy does not appear to have an effect.
- Time-varying Cox models strike a nice balance between clinical interpretability and statistical rigor.
- Everyone should take a survival class!

## Limitations and Future Directions

- Methods available for interval censored data in the `survival` R package[@therneauPackageSurvivalAnalysis2024] are slightly limited.
  - Other packages such as `icenReg`[@anderson-bergmanIcenRegRegressionModels2017] may be able to fit more flexible models.
- Categorizing variables is not always ideal, and there is some concern about loss of information.
  - Currently we are working on extending this analysis using joint modeling for more flexibility.
    - We will be able to use encounter-level data for this.

## Limitations and Future Directions

- Data end in 2018 but Trikafta (a triple combination therapy of Elexacaftor, Ivacaftor, and Tezacaftor) was approved in 2019.
  - Need to study later generations of CFTR modulators.
- Registry data has natural limitations because each center must track and report their own data.

## Special Thanks

- The Cystic Fibrosis Foundation for the use of CF Foundation Patient Registry data.
- The patients, care providers, and clinic coordinators at CF centers.
- 2025 WNAR/IMS
- Elizabeth Juarez-Colunga, PhD
- Kristen Miller, MS

## References