### 2025-07-14 Emily Muther/body_image_trikafta.qmd
raw BMIs for body image and trikafta

---
title: "Body Image and Trikafta"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

# Data Cleaning Steps

1. Participants who answered "I don't know" or "no" on the variable `TakingTrikafta` were NOT excluded from these analyses.

```{r warning=FALSE}
#| include: false
library(tidyverse)
library(gtsummary)
library(kableExtra)
library(foreign)
library(plotly)
library(broom)
library(ggstats)
library(Hmisc)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
#| warning: false
# Read in data
df <- read.spss("/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Emily Muther/Muther Research Group - General/BodyImage_Trikafta/Data/TrikaftaBodyImage3.6.23.sav",
  to.data.frame = T
)
# Filter
# df <- df %>%
#   filter(TakingTrikafta == "yes")
# Trim white space
df <- df %>%
  mutate(across(where(is.character), trimws))
# Replace blanks with NA
df[df == ""] <- NA
# Categorical variables
df$FRS_Category <- relevel(df$FRS_Category, ref = "neutral")
levels(df$FRS_Category) <- c(
  "Neutral", "Want to be larger", "Want to be smaller"
)
levels(df$WeightBody_Adherence) <- c("Agree", "Agree", "Neutral", "Disagree", "Disagree")
# Raw BMI values from Emma
df$bmi_raw <- NA
df$bmi_raw[df$ID == 1] <- 24.21
df$BMI[df$ID == 1] <- NA
df$bmi_raw[df$ID == 8] <- 22.33
df$BMI[df$ID == 8] <- NA
df$bmi_raw[df$ID == 11] <- 21.1
df$BMI[df$ID == 11] <- NA
df$bmi_raw[df$ID == 31] <- 23.64
df$BMI[df$ID == 31] <- NA
df$bmi_raw[df$ID == 39] <- 20.73
df$BMI[df$ID == 39] <- NA
# Labels
label(df$FRS_Category) <- "FRS Category"
label(df$BE_APPEARANCE_MEAN) <- "Mean Body Esteem Scale (Appearance)"
label(df$BE_ATTRIBUTION_MEAN) <- "Mean Body Esteem Scale (Attribution)"
label(df$BE_WEIGHT_MEAN) <- "Mean Body Esteem Scale (Weight)"
label(df$BMI) = "BMI %ile"
label(df$bmi_raw) = "BMI (kg/m2)"
```

# Descriptive analyses

## Demographics

```{r}
#| label: demographics
normal_dist <- c("FEV1", "Age", "BMI")
df %>%
  select(
    PHQ9_Total, PHQ9_category, GAD7_Total, GAD7_category, FRStotal,
    FRS_IdealBody, FRS_ThinkyouLook, FRS_Category, Age, Gender, Race,
    Ethnicity, InsuranceType, BMI, bmi_raw,FEV1
  ) %>%
  tbl_summary(
    by = Gender,
    type = list(
      c("FRStotal", "FRS_IdealBody", "FRS_ThinkyouLook","bmi_raw") ~ "continuous2",
      all_continuous() ~ "continuous2"
    ),
    statistic = list(
      all_continuous() ~ c(
        "{median} [{p25}, {p75}]",
        "{min}, {max}"
      ),
      all_of(normal_dist) ~ c("{mean} ({sd})", "{min}, {max}")
    ),
    missing_text = "(Missing)"
  ) %>%
  add_overall() %>%
  add_p(
    test = list(all_of(normal_dist) ~ "t.test"),
    test.args = all_tests("wilcox.test") ~ list(exact = F)
  ) %>%
  separate_p_footnotes()
```

## Frequency of reported changes

```{r}
#| label: changes
df %>%
  select(ChangestoBreathing:MoodChangesSideEf, FeelTrikafta) %>%
  tbl_summary(
    type = list(
      all_categorical() ~ "categorical"
    ),
    missing_text = "(Missing)"
  )
```

## Adherence by FRS category

```{r}
#| label: adherence
df %>%
  select(contains("Adherence"), FRS_Category) %>%
  mutate(
    PercentAdherence = as.numeric(sub("%", "", as.character(PercentAdherence))),
    over80 = PercentAdherence >= 80
  ) %>%
  tbl_summary(
    by = "FRS_Category",
    type = list("PercentAdherence" ~ "continuous"),
    missing_text = "(Missing)"
  ) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

# Mental health scores by FRS category

## PHQ9

### Odds of an elevated score

Results are reported as odds ratios, with the FRS "neutral" category as the reference group.

```{r}
mod <- glm(PHQ9_elevated ~ FRS_Category, data = df, family = "binomial")
tbl_regression(mod)
```

### Association with FRS category in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod2 <- glm(PHQ9_Total ~ FRS_Category,
  data = df[df$PHQ9_elevated == "yes", ],
  family = "poisson"
)
tbl_regression(mod2)
```

## GAD7

### Odds of an elevated score

Results are reported as odds ratios, with the FRS "neutral" category as the reference group.

```{r}
mod3 <- glm(GAD7_elevated ~ FRS_Category, data = df, family = "binomial")
tbl_regression(mod3)
```

### Association with FRS category in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod4 <- glm(GAD7_Total ~ FRS_Category, data = df[df$GAD7_elevated == "yes", ], family = "poisson")
tbl_regression(mod4)
```

# Mental health scores by BES

## PHQ9

### BE_APPEARANCE_MEAN

#### Odds of an elevated score

Results are reported as odds ratios for a one-unit change in the predictor.

```{r}
mod5 <- glm(PHQ9_elevated ~ BE_APPEARANCE_MEAN, data = df, family = "binomial")
tbl_regression(mod5)
```

#### Association with BE_APPEARANCE_MEAN in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod6 <- glm(PHQ9_Total ~ BE_APPEARANCE_MEAN,
  data = df[df$PHQ9_elevated == "yes", ],
  family = "poisson"
)
tbl_regression(mod6)
```

### BE_ATTRIBUTION_MEAN

#### Odds of an elevated score

Results are reported as odds ratios for a one-unit change in the predictor.

```{r}
mod7 <- glm(PHQ9_elevated ~ BE_ATTRIBUTION_MEAN, data = df, family = "binomial")
tbl_regression(mod7)
```

#### Association with BE_ATTRIBUTION_MEAN in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod8 <- glm(PHQ9_Total ~ BE_ATTRIBUTION_MEAN,
  data = df[df$PHQ9_elevated == "yes", ],
  family = "poisson"
)
tbl_regression(mod8)
```

### BE_WEIGHT_MEAN

#### Odds of an elevated score

Results are reported as odds ratios for a one-unit change in the predictor.

```{r}
mod9 <- glm(PHQ9_elevated ~ BE_WEIGHT_MEAN, data = df, family = "binomial")
tbl_regression(mod9)
```

#### Association with BE_WEIGHT_MEAN in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod10 <- glm(PHQ9_Total ~ BE_WEIGHT_MEAN,
  data = df[df$PHQ9_elevated == "yes", ],
  family = "poisson"
)
tbl_regression(mod10)
```

### Forest plot of logistic regression results

## GAD7

### BE_APPEARANCE_MEAN

#### Odds of an elevated score

Results are reported as odds ratios for a one-unit change in the predictor.

```{r}
mod11 <- glm(GAD7_elevated ~ BE_APPEARANCE_MEAN, data = df, family = "binomial")
tbl_regression(mod11)
```

#### Association with BE_APPEARANCE_MEAN in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod12 <- glm(GAD7_Total ~ BE_APPEARANCE_MEAN, data = df[df$GAD7_elevated == "yes", ], family = "poisson")
tbl_regression(mod12)
```

### BE_ATTRIBUTION_MEAN

#### Odds of an elevated score

Results are reported as odds ratios for a one-unit change in the predictor.

```{r}
mod13 <- glm(GAD7_elevated ~ BE_ATTRIBUTION_MEAN, data = df, family = "binomial")
tbl_regression(mod13)
```

#### Association with BE_ATTRIBUTION_MEAN in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod14 <- glm(GAD7_Total ~ BE_ATTRIBUTION_MEAN,
  data = df[df$GAD7_elevated == "yes", ],
  family = "poisson"
)
tbl_regression(mod14)
```

### BE_WEIGHT_MEAN

#### Odds of an elevated score

Results are reported as odds ratios for a one-unit change in the predictor.

```{r}
mod15 <- glm(GAD7_elevated ~ BE_WEIGHT_MEAN, data = df, family = "binomial")
tbl_regression(mod15)
```

#### Association with BE_WEIGHT_MEAN in those with elevated scores.

Results are reported as expected counts, with the FRS "neutral" category as the reference group.

```{r}
mod16 <- glm(GAD7_Total ~ BE_WEIGHT_MEAN,
  data = df[df$GAD7_elevated == "yes", ],
  family = "poisson"
)
tbl_regression(mod16)
```

# Nicely formatted stuff

## Table

```{r}
tbl_stack(
  list(
    mod %>%
      tbl_regression(exponentiate = T, intercept = T),
    mod5 %>%
      tbl_regression(exponentiate = T, intercept = T),
    mod7 %>%
      tbl_regression(exponentiate = T, intercept = T),
    mod9 %>%
      tbl_regression(exponentiate = T, intercept = T),
    mod3 %>%
      tbl_regression(exponentiate = T, intercept = T),
    mod11 %>%
      tbl_regression(exponentiate = T, intercept = T),
    mod13 %>%
      tbl_regression(exponentiate = T, intercept = T),
    mod15 %>%
      tbl_regression(exponentiate = T, intercept = T)
  ),
  group_header = c(
    rep("Elevated PHQ9", times = 4),
    rep("Elevated GAD7", times = 4)
  )
)
```

This table shows the odds ratios and confidence intervals for an elevated score on the PHQ9 or GAD7. Odds ratios for FRS category represent the multiplicative difference in odds compared the reference group ("Neutral"), and estimates for BES mean scores represent the average multiplicative change in odds for a one unit increase in score. For example, the odds of an elevated score on the GAD for someone who endorsed wanting to be larger is 0.75 times those of someone who endorsed feeling neutral about their body (i.e. a decrease of 25%).

## Forest plot of logistic regression results

```{r}
# Plot models together
p <- ggcoef_compare(list(
  "PHQ9" = mod,
  "PHQ9" = mod5, "PHQ9" = mod7, "PHQ9" = mod9, "GAD7" = mod3,
  "GAD7" = mod11, "GAD7" = mod13, "GAD7" = mod15
), exponentiate = T, add_reference_rows = F) +
  theme(text = element_text(size = 18))
p
# Save
# ggsave("./Emily Muther/General/BodyImage_Trikafta/Reports/forest_plot.png",
#   plot = p, width = 9, height = 6, units = "in"
# )
```

Figure 1 shows the unadjusted odds ratios and confidence intervals for an elevated score on the PHQ9 or GAD7. Filled circles indicate an odds ratio significantly different from 1. Estimates to the left of the vertical bar indicate lower odds of an elevated score.