### 2025-07-23 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
both study IDs for EnVision and GEM fsOGTT

---
title: "fsOGTT Pre- and Post-ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(lmerTest)
library(performance)
library(gtsummary)
library(emmeans)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import cleaned EnVision data
df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
# Import GEM data
gem <- read.csv(
  "./Data_Raw/CGMInHealthyControls-FsOGTTDataForMonique_DATA_2025-04-10_0903.csv",
  na.strings = ""
)
# ID linking from Monique
ids <- read_excel("./Data_Raw/ETI participant IDs_TV.xlsx")
ids <- ids[!is.na(ids$`GEM ID`), ]
ids$`EnVision ID` <- sub("-2", "", ids$`EnVision ID`)
# Get pre-/post-ETI
df$eti_start <- ifelse(df$corrector___4 == 1, df$corrector_start_date, NA)
df <- df %>%
  group_by(study_id) %>%
  fill(eti_start, .direction = "downup")
df$pre_post <-
  factor(df$Date >= df$eti_start, levels = c(F, T), labels = c("Pre", "Post"))
# Select columns for EnVision per Monique and combine race and ethnicity
df <- df %>%
  mutate(gem_id = NA, study = "EnVision CF") %>%
  select(
    study_id, gem_id, study, Date, age_visit, sex, origin_race, ethnicity, CFTR,
    pancreatic_status, corrector___4, pre_post, bmi, fev1,
    Diagnosis, Glucose_0:Glucose_180, Insulin_0:Insulin_180,
    C.Peptide_0:C.Peptide_180
  ) %>%
  unite(race_ethnicity, c(origin_race, ethnicity)) %>%
  rename(ETI = corrector___4)
df$race_ethnicity <- factor(df$race_ethnicity,
  levels = c(
    "American Indian or Alaska Native (N. S. & Central America)_Hispanic/Latino",
    "American Indian or Alaska Native (N. S. & Central America)_Not Hispanic/Latino",
    "Black or African American (Africa)_N/A",
    "Black or African American (Africa)_Not Hispanic/Latino",
    "Declined_Hispanic/Latino",
    "Declined_Not Hispanic/Latino",
    "NA_NA",
    "Native Hawaiian or Other Pacific Islander (Hawaii, Guam, Samoa, Pacific Islands)_Not Hispanic/Latino",
    "White (Europe, Middle East, North Africa)_Hispanic/Latino",
    "White (Europe, Middle East, North Africa)_N/A",
    "White (Europe, Middle East, North Africa)_Not Hispanic/Latino"
  ),
  labels = c(
    "Hispanic", "Native American", "Black/African American", "Black/African American",
    "Hispanic", "Other/Multiple", "Other/Multiple", "Other/Multiple", "Hispanic", "White", "White"
  )
)
df$pancreatic_status <- factor(df$pancreatic_status,
  levels = 1:2,
  labels = c("Sufficient", "Insufficient")
)
df$ETI <- factor(df$ETI, levels = 0:1, labels = c("No", "Yes"))
# Select the same columns from GEM
gem <- gem %>%
  mutate(ETI = 0, pre_post = "Pre", study_id = NA, study = "GEM-CF") %>%
  select(
    study_id, subject_id, study, date_of_visit, age_at_visit_1, gender,
    race_ethnicity, cf_genotype, cf_pancreatic, ETI, pre_post, average_bmi,
    fev1_percent, group_category, lab_ogtt_fasting, min_10_glucose,
    min_30_glucose, lab_ogtt_1_hour_glucose, min_90_glucose,
    lab_ogtt_2_hour_glucose, min_150_glucose, min_180_glucose,
    min_0_insulin, min_10_insulin, min_30_insulin, min_60_insulin,
    min_90_insulin, min_120_insulin, min_150_insulin, min_180_insulin,
    min_0_c_peptide, min_10_c_peptide, min_30_c_peptide, min_60_c_peptide,
    min_90_c_peptide, min_120_c_peptide, min_150_c_peptide, min_180_c_peptide
  ) %>%
  rename(
    gem_id = subject_id, Date = date_of_visit, age_visit = age_at_visit_1,
    sex = gender, CFTR = cf_genotype, pancreatic_status = cf_pancreatic,
    bmi = average_bmi, fev1 = fev1_percent, Diagnosis = group_category,
    min_0_glucose = lab_ogtt_fasting, min_60_glucose = lab_ogtt_1_hour_glucose,
    min_120_glucose = lab_ogtt_2_hour_glucose
  ) %>%
  rename_with(
    ~ paste0("Glucose_", parse_number(.)), min_0_glucose:min_180_glucose
  ) %>%
  rename_with(
    ~ paste0("Insulin_", parse_number(.)), min_0_insulin:min_180_insulin
  ) %>%
  rename_with(
    ~ paste0("C.Peptide_", parse_number(.)), min_0_c_peptide:min_180_c_peptide
  )
# GEM factors
gem$sex <- factor(gem$sex, levels = 1:2, labels = c("Male", "Female"))
gem$race_ethnicity <- factor(gem$race_ethnicity,
  levels = 1:6,
  labels = c(
    "White", "Black/African American", "Asian", "Native American",
    "Hispanic", "Other/Multiple"
  )
)
gem$CFTR <- factor(gem$CFTR,
  levels = c(
    "2184delA/M1101K", "2585delT/G551D", "3849+10kb C>T/F508del",
    "508del/508del", "F508del/1078delA", "f508DEL/1154INStc",
    "F508del/1154insTC", "F508del/1811 + 1634A > G", "F508del/1811+1634>G",
    "F508del/2183delAA->G", "F508del/2183delAA>G", "F508del/2184insA",
    "F508del/2711delT", "F508del/2789+5G>A", "F508del/27895GA",
    "F508del/3659delC", "F508del/3849+10kbC>T", "F508del/394delTT",
    "F508del/5492F", "F508del/A561E", "F508del/E60X",
    "f508del/F508del", "F508del/F508del", "F508del/G542X",
    "F508del/G551D", "F508del/N1303K", "F508del/Q98R",
    "F508del/R1162X", "F508del/R117H 9T", "F508del/R117H 9T/5T",
    "F508del/R553X", "F508del/R709X", "F508del/S492F",
    "F508del/W1282X", "F5o8del/F508del", "G542X/1078delT",
    "G542X/W1089X", "G551D/W846X", "I507del/G551D",
    "N1303K/1154insTC", "N1303K/2789 + 5G > A", "N1303K/2789+5G>A",
    "R117H 5t/G576A and R668C on exon 13", "R117H5t/G576A and R668C on exon 13",
    "V520F/3791delC", "W1089X/S492F", "W1090X/S492F"
  ),
  labels = c(
    "Other/Other", "Other/Other", "Other/Other", "F508del homozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del homozygous",
    "F508del homozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del homozygous", "Other/Other", "Other/Other", "Other/Other",
    "Other/Other", "Other/Other", "Other/Other", "Other/Other", "Other/Other",
    "Other/Other", "Other/Other", "Other/Other", "Other/Other"
  )
)
gem$pancreatic_status <- factor(gem$pancreatic_status,
  levels = 1:2,
  labels = c("Sufficient", "Insufficient")
)
gem$ETI <- factor(gem$ETI, levels = 0:1, labels = c("No", "Yes"))
gem$Diagnosis <- factor(gem$Diagnosis,
  levels = 1:4,
  labels = c("Healthy Control", "NGT", "AGT", "CFRD")
)
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "IGT", "INDET", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "CFRD")
)
# IDs
gem$study_id <- ids$`EnVision ID`[match(gem$gem_id, ids$`GEM ID`)]
df$gem_id <- ids$`GEM ID`[match(df$study_id, ids$`EnVision ID`)]
# Limit GEM per Monique's list
gem <- gem %>% filter(gem_id %in% ids$`GEM ID`)
# Merge, keep only pre/post, sort by ID and date
df <- rbind(df, gem) %>% filter(!is.na(pre_post))
df$pre_post <- factor(df$pre_post, levels = c("Pre", "Post"))
df <- df %>% arrange(study_id, Date)
# Write for checking
write.csv(df,
  file = "./Data_Clean/fsOGTT_pre_post_ETI.csv",
  row.names = F, na = ""
)
```

# Plot

```{r}
df %>%
  group_by(study_id) %>%
  count(pre_post) %>%
  ggplot(., aes(y = study_id, x = n)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  facet_wrap(~pre_post)
```

- Adjust models for age
- Paneled figure with results for glucose, c-peptide, and insulin
- Look at acute insulin response (change in 30 minute c peptide)/(change in 30 minute glucose)
  - Can also do across the whole time period and can use insulin in the numerator also
- See Teams folder document for detailed plan

# Glucose

```{r}
# Make a long DF for models
long_df <- df %>% pivot_longer(Glucose_0:Glucose_180,
  names_to = c(".value", "timepoint"),
  names_pattern = "(.*)_(.*)",
  values_to = "Glucose"
)
long_df$timepoint <- factor(long_df$timepoint,
  levels = c("0", "10", "30", "60", "90", "120", "150", "180")
)
mod1 <- lmer(
  Glucose ~ timepoint * pre_post + age_visit + (1 | study_id),
  data = long_df
)
mod1 %>% tbl_regression(.)
means <- data.frame(emmeans(mod1, ~ timepoint + pre_post))
knitr::kable(means)
ggplot(means, aes(
  x = as.numeric(as.character(timepoint)), y = emmean,
  color = pre_post, group = pre_post
)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1) +
  geom_line() +
  theme_bw() +
  xlab("Timepoint") +
  ylab("Glucose (mg/dL)") +
  scale_x_continuous(breaks = c(0, 10, 30, 60, 90, 120, 150, 180))
```