### 2025-07-28 Christine Chan/EnVision CF/cgm_and_a1c.qmd
Envision CGM and HbA1c

---
title: "The relationship between HbA1c, GMI, and average glucose in adults with CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(redcapAPI)
library(gtsummary)
library(labelled)
library(ggpmisc)
library(blandr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "EnVision CF HbA1c and CGM"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ahc.umn.edu/redcap/api/"
)
df <- exportRecordsTyped(rcon)
# Start building analysis dataset
df <- df %>%
  # Format columns
  mutate(
    # Numeric
    across(c(age, bmi, cfrd_diagnosis_age, average_sensor_glucose), ~
      as.numeric(.)),
    across(c(hb_a1c_value, gmi), ~
      as.numeric(sub("%", "", .))),
    # Categorical
    combined_race = factor(race,
      levels = c(
        "White", "Black or African American", "Asian", "More Than One Race",
        "Unknown / Not Reported"
      ),
      labels = c(
        "White", "Black or African American", "Other/Unknown", "Other/Unknown",
        "Other/Unknown"
      )
    )
  ) %>%
  # Drop unnecessary columns
  select(
    subject_id, age, cfrd_diagnosis_age, bmi, combined_race, sex, pi,
    cftr_modulator_use, type_of_modulator, type_of_cgm, hb_a1c_value, gmi,
    average_sensor_glucose
  ) %>%
  # Labels
  set_variable_labels(
    age = "Age (years)", cfrd_diagnosis_age = "Age at CFRD Diagnosis (years)",
    bmi = "BMI (kg/m2)", combined_race = "Race", sex = "Sex at Birth",
    type_of_modulator = "Type of Modulator", hb_a1c_value = "HbA1c (%)",
    gmi = "GMI (%)", average_sensor_glucose = "Mean Sensor Glucose (mg/dL)",
    cftr_modulator_use = "CFTR Modulator", pi = "Pancreatic Insufficiency"
  )
```

# Participant Characteristics

```{r}
#| label: tbl-1
#| tbl-cap: Table 1
df %>%
  select(-subject_id) %>%
  tbl_summary(
    missing_text = "Missing",
    type = all_continuous() ~ "continuous2",
    statistic = list(
      all_continuous() ~ c("{median} [{p25}, {p75}]", "{min} - {max}"),
      average_sensor_glucose ~ c("{mean} ({sd})", "{min} - {max}")
    ),
  )
```

# GMI and HbA1c

```{r}
#| warning: false
ggplot(df, aes(x = hb_a1c_value, y = gmi)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$hb_a1c_value, df$gmi, ciDisplay = F, annotate = T)
```

# GMI and HbA1c

```{r}
#| warning: false
ggplot(df, aes(x = hb_a1c_value, y = average_sensor_glucose)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$hb_a1c_value, df$average_sensor_glucose,
  ciDisplay = F, annotate = T
)
```