### 2025-08-07 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
genotype severity analysis

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import CHCO, filter out those with no demo data, rename columns, etc.
chco <- read.csv("./Data_Cleaned/chco_reidentified.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing", "Please reorder testing and re",
    "The received specimen cannot"
  )
)
chco <- chco %>%
  group_by(PatientID) %>%
  filter(Demographics.Available == "TRUE") %>%
  select(
    PatientID, Name, DOB, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
    Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
  ) %>%
  rename(Age = Age.at.Test.in.Years) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
# Fix CHCO diagnoses per Angela
chco_diagnoses <- read.csv("./Data_Cleaned/chco_diagnoses.csv")
chco <- full_join(chco, chco_diagnoses, by = join_by(PatientID))
chco$PatientID <- NULL
# Manually add some Na and Cl values that Angela was able to find (using
# numbers so that no PHI is on GitHub)
chco$Na_Right[791] <- 61.3
chco$Na_Right[826] <- 72.4
chco$Na_Left[826] <- 72.4
chco$Na_Right[790] <- 21.1
chco$Na_Left[790] <- 24.1
# For "LESS THAN 10", replace with missing
chco[chco == "LESS THAN 10"] <- NA
chco[, c("Cl", "Cl_Left", "Cl_Right")] <-
  lapply(chco[, c("Cl", "Cl_Left", "Cl_Right")], as.numeric)
# Import NJH and add diagnoses from Katie
njh <- read.csv("./Data_Cleaned/njh_full_list.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing"
  )
)
njh_diagnoses <- read.csv("./Data_Cleaned/njh_diagnoses.csv", na.strings = "")
njh_diagnoses <- njh_diagnoses %>% filter(!is.na(Diagnosis))
njh <- full_join(njh, njh_diagnoses)
# Filter out those with no demo data, rename columns, etc.
njh <- njh %>%
  group_by(Patient.ID) %>%
  filter(Demographics.Available == "TRUE") %>%
  select(
    Patient.ID, Patient.Name, Birth.Date, Genotype.1, Genotype.2,
    Sex, Age.at.test..in.years., Sweat.sodium, Sweat.potassium, Sweat.chloride,
    Sweat.sodium.left.arm, Sweat.sodium.right.arm, Sweat.potassium.left.arm,
    Sweat.potassium.right.arm, Sweat.chloride.left.arm,
    Sweat.chloride.right.arm, Diagnosis
  ) %>%
  rename(
    Name = Patient.Name, Genotypes1 = Genotype.1, DOB = Birth.Date,
    Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
    K = Sweat.potassium, Cl = Sweat.chloride, Na_Left = Sweat.sodium.left.arm,
    Na_Right = Sweat.sodium.right.arm, K_Left = Sweat.potassium.left.arm,
    K_Right = Sweat.potassium.right.arm, Cl_Left = Sweat.chloride.left.arm,
    Cl_Right = Sweat.chloride.right.arm
  ) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
njh$Patient.ID <- NULL
# Row bind, fill diagnosis, then remove dupes
df <- rbind(chco, njh)
df <- df %>%
  group_by(Name) %>%
  fill(Diagnosis) %>%
  distinct() %>%
  select(Name, Diagnosis, everything()) %>%
  arrange(Name, Age)
# For those with left and right, average them
df$Na_mean <- rowMeans(df[, c("Na_Left", "Na_Right")], na.rm = T)
df$Na <- coalesce(df$Na, df$Na_mean)
df$Cl_mean <- rowMeans(df[, c("Cl_Left", "Cl_Right")], na.rm = T)
df$Cl <- coalesce(df$Cl, df$Cl_mean)
df$K_mean <- rowMeans(df[, c("K_Left", "K_Right")], na.rm = T)
df$K <- coalesce(df$K, df$K_mean)
# Change Cl under 10 or over 160 to missing
df$Cl[df$Cl < 10] <- NA
df$Cl[df$Cl > 160] <- NA
# Remove Na under 10 as well
df$Na[df$Na < 10] <- NA
# Calculate Na:Cl ratio
df$`Na:Cl` <- df$Na / df$Cl
# Use each person's first sweat test with both sodium and chloride
df <- df %>%
  filter(!is.na(`Na:Cl`)) %>%
  group_by(Name) %>%
  slice_min(Age)
# If there are still duplicates, exclude them
dups <- df$Name[duplicated(df$Name)]
df <- df[!df$Name %in% dups, ]
# Get rid of unnecessary columns, remove those with no diagnosis
df <- df %>%
  select(
    Name, DOB, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl, `Na:Cl`
  ) %>%
  filter(!is.na(Diagnosis))
# Combine CFTR-related disorder and CRMS
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("Cystic Fibrosis", "CFTR-related disorder", "CRMS", "Non-CF"),
  labels = c("Cystic Fibrosis", "CFTR-RD or CRMS", "CFTR-RD or CRMS", "Non-CF")
)
# Additional changes based on Angela's review
df$Diagnosis[which(is.na(df$Name))] <- "Cystic Fibrosis"
df$Diagnosis[400] <- "Cystic Fibrosis"
df$Na[541] <- 69
# Angela reviewed a random set of 75 patients and requested the changes below
# See the "random_pull.csv" document for details
# Changes done by row number to avoid PHI in GitHub
df$Cl[1473] <- 103.5
df$Na[505] <- 42.6
df$Na[238] <- 73
df$Na[541] <- 69
df$Na[747] <- 88.9
# Remove people per Angela's recommendation and duplicates based on DOB review
df <- df[-c(
  78, 79, 183, 257, 259, 270, 314, 342, 691, 719, 724, 737, 812, 903, 1012,
  1066, 1126, 1151, 1202, 1209, 1222, 1224, 1340, 1428, 1456
), ]
# Get genotype severity
## By severity
df$geno_risk_1 <- cf_genotype_class_severity(df$Genotypes1)
df$geno_risk_2 <- cf_genotype_class_severity(df$Genotypes2)
df$geno_risk_1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk_2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno_risk_1, df$geno_risk_2, sep = "/")
df$geno_risk <- factor(df$geno_risk,
  levels = c(
    "Mild/Mild", "Mild/Severe", "Mild/Unknown", "Severe/Mild", "Severe/Severe",
    "Severe/Unknown", "Unknown/Severe", "Unknown/Unknown", "Not CF/Not CF"
  ),
  labels = c("M/M", "M/S", "M/U", "M/S", "S/S", "S/U", "S/U", "U", "Not CF")
)
## By function
df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
df$geno1_class[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2_class[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_class <- paste(df$geno1_class, df$geno2_class, sep = "/")
df$geno_class <- factor(df$geno_class,
  levels = c(
    "F508del/F508del", "F508del/MF", "F508del/RF", "F508del/Unknown",
    "MF/F508del", "MF/MF", "MF/RF", "MF/Unknown", "RF/F508del", "RF/MF",
    "RF/Unknown", "Unknown/F508del", "Unknown/MF", "Unknown/Unknown",
    "Not CF/Not CF"
  ),
  labels = c(
    "MF/MF", "MF/MF", "MF/RF", "MF/U", "MF/MF", "MF/MF", "MF/RF", "MF/U",
    "MF/RF", "MF/RF", "RF/U", "MF/U", "MF/U", "U", "Not CF"
  )
)
# Organize columns
df <- df %>%
  select(Name:Genotypes2, geno_risk, geno_class, everything()) %>%
  select(-geno_risk_1, -geno_risk_2, -geno1_class, -geno2_class) %>%
  rename(`Genotype Risk` = geno_risk, `Genotype Class` = geno_class)
# Set table options
mycontrols <- tableby.control(
  numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range")
)
# Save this for Angela et al. to check
write.csv(df,
  file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
  row.names = F, na = ""
)
```

# Full Cohort

## Table 1: Participant Demographics by Diagnosis

```{r results='asis'}
summary(tableby(Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

## Table 2: Participant Demographics by Genotype Class

In the table below "MF" stands for minimal function, "RF" for residual function, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(`Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

## Table 3: Participant Demographics by Genotype Risk

In the table below "M" stands for mild, "S" for severe, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(`Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

## Boxplots

### Chloride

#### By diagnosis

```{r}
#| warning: false
clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

#### By genotype class

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Class`, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

#### By genotype risk

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

### Sodium

#### By diagnosis

```{r}
#| warning: false
nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

#### By genotype class

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Class`, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

#### By genotype risk

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

### Na:Cl

#### By diagnosis

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

#### By genotype class

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

#### By genotype risk

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

## Na:Cl ROC Curves

```{r}
# Fit a multiclass ROC
mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
roc_list <- list(
  "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
)
ggroc(roc_list, legacy.axes = T) +
  theme_bw() +
  theme(legend.title = element_blank())
roc_table <- lapply(roc_list, function(r) {
  coords(r, "best",
    best.method = "youden",
    ret = c(
      "threshold", "sensitivity", "specificity", "accuracy", "tpr", "fpr",
      "tnr", "fnr"
    )
  )
})
roc_table <- do.call(rbind, roc_table)
kable(roc_table, digits = 3)
```

- The multiclass AUC for Na:Cl is `r round(as.numeric(auc(mc_roc)),3)`.

### Individual ROC Curves

```{r}
plot(roc_list[[2]],
  print.auc = T, print.thres = T,
  main = "Cystic Fibrosis vs. CFTR-RD or CRMS"
)
plot(roc_list[[1]],
  print.auc = T, print.thres = T,
  main = "Cystic Fibrosis vs. Non-CF"
)
plot(roc_list[[3]],
  print.auc = T, print.thres = T,
  main = "CFTR-RD or CRMS vs. Non-CF"
)
```

# Excluding Cl $\leq$ 30

```{r}
df <- df %>% filter(Cl > 30)
```

## Table 1: Participant Demographics by Diagnosis

```{r results='asis'}
summary(tableby(Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

## Table 2: Participant Demographics by Genotype Class

In the table below "MF" stands for minimal function, "RF" for residual function, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(`Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

## Table 3: Participant Demographics by Genotype Risk

In the table below "M" stands for mild, "S" for severe, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(`Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

## Boxplots

### Chloride

#### By diagnosis

```{r}
#| warning: false
clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

#### By genotype class

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Class`, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

#### By genotype risk

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

### Sodium

#### By diagnosis

```{r}
#| warning: false
nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

#### By genotype class

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Class`, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

#### By genotype risk

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

### Na:Cl

#### By diagnosis

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

#### By genotype class

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

#### By genotype risk

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

## Na:Cl ROC Curves

```{r}
# Fit a multiclass ROC
mc_roc <- multiclass.roc(df$Diagnosis, df$`Na:Cl`, quiet = T)
roc_list <- list(
  "Cystic Fibrosis vs. Non-CF" = mc_roc$rocs[[2]],
  "CFTR-RD or CRMS vs. Cystic Fibrosis" = mc_roc$rocs[[1]],
  "CFTR-RD or CRMS vs. Non-CF" = mc_roc$rocs[[3]]
)
ggroc(roc_list, legacy.axes = T) +
  theme_bw() +
  theme(legend.title = element_blank())
roc_table <- lapply(roc_list, function(r) {
  coords(r, "best",
    best.method = "youden",
    ret = c(
      "threshold", "sensitivity", "specificity", "accuracy", "tpr", "fpr",
      "tnr", "fnr"
    )
  )
})
roc_table <- do.call(rbind, roc_table)
kable(roc_table, digits = 3)
```

- The multiclass AUC for Na:Cl is `r round(as.numeric(auc(mc_roc)),3)`.

### Individual ROC Curves

```{r}
plot(roc_list[[2]],
  print.auc = T, print.thres = T,
  main = "Cystic Fibrosis vs. CFTR-RD or CRMS"
)
plot(roc_list[[1]],
  print.auc = T, print.thres = T,
  main = "Cystic Fibrosis vs. Non-CF"
)
plot(roc_list[[3]],
  print.auc = T, print.thres = T,
  main = "CFTR-RD or CRMS vs. Non-CF"
)
```