### 2025-08-12 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
additional ETI date fixes

---
title: "fsOGTT Pre- and Post-ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(lmerTest)
library(performance)
library(gtsummary)
library(emmeans)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import cleaned EnVision data
df <- read.csv("./Data_Clean/envision_analysis_dataset.csv", na.strings = "")
# Import GEM data
gem <- read.csv(
  "./Data_Raw/CGMInHealthyControls-FsOGTTDataForMonique_DATA_2025-04-10_0903.csv",
  na.strings = ""
)
# ID linking from Monique
ids <- read.csv("./Data_Clean/TV_ETI logitudinal patient list.csv", na.strings = "")
ids <- ids %>%
  select(EnVision.ID, GEM.ID) %>%
  rename(`EnVision ID` = EnVision.ID, `GEM ID` = GEM.ID)
ids <- ids[!is.na(ids$`GEM ID`), ]
# Get pre-/post-ETI
df$eti_start <- ifelse(df$corrector___4 == 1, df$corrector_start_date, NA)
df <- df %>%
  group_by(study_id) %>%
  fill(eti_start, .direction = "downup")
# Manually correct a few per Monique
df$eti_start[df$study_id == "CC0001"] <- "2021-11-24"
df$eti_start[df$study_id == "CC0010"] <- "2023-01-16"
df$eti_start[df$study_id == "CC0022"] <- "2019-12-06"
df$eti_start[df$study_id == "CC0023"] <- "2020-01-09"
df$eti_start[df$study_id == "CC0026"] <- "2020-12-04"
df$eti_start[df$study_id == "CC0059"] <- "2021-06-23"
df$pre_post <-
  factor(df$Date >= df$eti_start, levels = c(F, T), labels = c("Pre", "Post"))
df$pre_post[df$study_id == "CC0012"] <- "Pre"
df$pre_post[df$study_id == "CC0017"] <- "Pre"
df$pre_post[df$study_id == "CC0028"] <- "Pre"
# Select columns for EnVision per Monique and combine race and ethnicity
df <- df %>%
  mutate(gem_id = NA, study = "EnVision CF") %>%
  select(
    study_id, gem_id, study, Date, age_visit, sex, origin_race, ethnicity, CFTR,
    pancreatic_status, corrector___4, pre_post, bmi, fev1,
    Diagnosis, Glucose_0:Glucose_180, Insulin_0:Insulin_180,
    C.Peptide_0:C.Peptide_180
  ) %>%
  unite(race_ethnicity, c(origin_race, ethnicity)) %>%
  rename(ETI = corrector___4)
df$race_ethnicity <- factor(df$race_ethnicity,
  levels = c(
    "American Indian or Alaska Native (N. S. & Central America)_Hispanic/Latino",
    "American Indian or Alaska Native (N. S. & Central America)_Not Hispanic/Latino",
    "Black or African American (Africa)_N/A",
    "Black or African American (Africa)_Not Hispanic/Latino",
    "Declined_Hispanic/Latino",
    "Declined_Not Hispanic/Latino",
    "NA_NA",
    "Native Hawaiian or Other Pacific Islander (Hawaii, Guam, Samoa, Pacific Islands)_Not Hispanic/Latino",
    "White (Europe, Middle East, North Africa)_Hispanic/Latino",
    "White (Europe, Middle East, North Africa)_N/A",
    "White (Europe, Middle East, North Africa)_Not Hispanic/Latino"
  ),
  labels = c(
    "Hispanic", "Native American", "Black/African American", "Black/African American",
    "Hispanic", "Other/Multiple", "Other/Multiple", "Other/Multiple", "Hispanic", "White", "White"
  )
)
df$pancreatic_status <- factor(df$pancreatic_status,
  levels = 1:2,
  labels = c("Sufficient", "Insufficient")
)
df$ETI <- factor(df$ETI, levels = 0:1, labels = c("No", "Yes"))
# Convert C-peptide level from pmol/L to ng/mL (multiply by 0.003)
df[, grep("C.Peptide_.*", colnames(df))] <-
  lapply(df[, grep("C.Peptide_.*", colnames(df))], function(c) {
    c * 0.003
  })
# Select the same columns from GEM
gem <- gem %>%
  mutate(ETI = 0, pre_post = "Pre", study_id = NA, study = "GEM-CF") %>%
  select(
    study_id, subject_id, study, date_of_visit, age_at_visit_1, gender,
    race_ethnicity, cf_genotype, cf_pancreatic, ETI, pre_post, average_bmi,
    fev1_percent, group_category, lab_ogtt_fasting, min_10_glucose,
    min_30_glucose, lab_ogtt_1_hour_glucose, min_90_glucose,
    lab_ogtt_2_hour_glucose, min_150_glucose, min_180_glucose,
    min_0_insulin, min_10_insulin, min_30_insulin, min_60_insulin,
    min_90_insulin, min_120_insulin, min_150_insulin, min_180_insulin,
    min_0_c_peptide, min_10_c_peptide, min_30_c_peptide, min_60_c_peptide,
    min_90_c_peptide, min_120_c_peptide, min_150_c_peptide, min_180_c_peptide
  ) %>%
  rename(
    gem_id = subject_id, Date = date_of_visit, age_visit = age_at_visit_1,
    sex = gender, CFTR = cf_genotype, pancreatic_status = cf_pancreatic,
    bmi = average_bmi, fev1 = fev1_percent, Diagnosis = group_category,
    min_0_glucose = lab_ogtt_fasting, min_60_glucose = lab_ogtt_1_hour_glucose,
    min_120_glucose = lab_ogtt_2_hour_glucose
  ) %>%
  rename_with(
    ~ paste0("Glucose_", parse_number(.)), min_0_glucose:min_180_glucose
  ) %>%
  rename_with(
    ~ paste0("Insulin_", parse_number(.)), min_0_insulin:min_180_insulin
  ) %>%
  rename_with(
    ~ paste0("C.Peptide_", parse_number(.)), min_0_c_peptide:min_180_c_peptide
  )
# GEM factors
gem$sex <- factor(gem$sex, levels = 1:2, labels = c("Male", "Female"))
gem$race_ethnicity <- factor(gem$race_ethnicity,
  levels = 1:6,
  labels = c(
    "White", "Black/African American", "Asian", "Native American",
    "Hispanic", "Other/Multiple"
  )
)
gem$CFTR <- factor(gem$CFTR,
  levels = c(
    "2184delA/M1101K", "2585delT/G551D", "3849+10kb C>T/F508del",
    "508del/508del", "F508del/1078delA", "f508DEL/1154INStc",
    "F508del/1154insTC", "F508del/1811 + 1634A > G", "F508del/1811+1634>G",
    "F508del/2183delAA->G", "F508del/2183delAA>G", "F508del/2184insA",
    "F508del/2711delT", "F508del/2789+5G>A", "F508del/27895GA",
    "F508del/3659delC", "F508del/3849+10kbC>T", "F508del/394delTT",
    "F508del/5492F", "F508del/A561E", "F508del/E60X",
    "f508del/F508del", "F508del/F508del", "F508del/G542X",
    "F508del/G551D", "F508del/N1303K", "F508del/Q98R",
    "F508del/R1162X", "F508del/R117H 9T", "F508del/R117H 9T/5T",
    "F508del/R553X", "F508del/R709X", "F508del/S492F",
    "F508del/W1282X", "F5o8del/F508del", "G542X/1078delT",
    "G542X/W1089X", "G551D/W846X", "I507del/G551D",
    "N1303K/1154insTC", "N1303K/2789 + 5G > A", "N1303K/2789+5G>A",
    "R117H 5t/G576A and R668C on exon 13", "R117H5t/G576A and R668C on exon 13",
    "V520F/3791delC", "W1089X/S492F", "W1090X/S492F"
  ),
  labels = c(
    "Other/Other", "Other/Other", "Other/Other", "F508del homozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del homozygous",
    "F508del homozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del heterozygous", "F508del heterozygous", "F508del heterozygous",
    "F508del homozygous", "Other/Other", "Other/Other", "Other/Other",
    "Other/Other", "Other/Other", "Other/Other", "Other/Other", "Other/Other",
    "Other/Other", "Other/Other", "Other/Other", "Other/Other"
  )
)
gem$pancreatic_status <- factor(gem$pancreatic_status,
  levels = 1:2,
  labels = c("Sufficient", "Insufficient")
)
gem$ETI <- factor(gem$ETI, levels = 0:1, labels = c("No", "Yes"))
gem$Diagnosis <- factor(gem$Diagnosis,
  levels = 1:4,
  labels = c("Healthy Control", "NGT", "AGT", "CFRD")
)
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "IGT", "INDET", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "CFRD")
)
# IDs
gem$study_id <- ids$`EnVision ID`[match(gem$gem_id, ids$`GEM ID`)]
df$gem_id <- ids$`GEM ID`[match(df$study_id, ids$`EnVision ID`)]
# Limit GEM per Monique's list
gem <- gem %>% filter(gem_id %in% ids$`GEM ID`)
# Merge, keep only pre/post, sort by ID and date
df <- rbind(df, gem) %>% filter(!is.na(pre_post))
df$pre_post <- factor(df$pre_post, levels = c("Pre", "Post"))
df <- df %>% arrange(study_id, Date)
# Calculate ODI, etc.
# Write for checking
write.csv(df,
  file = "./Data_Clean/fsOGTT_pre_post_ETI.csv",
  row.names = F, na = ""
)
```