### 2025-08-12 Christine Chan/EnVision CF/cgm_and_a1c.qmd
formatting for HbA1c and CGM

---
title: "The relationship between HbA1c, GMI, and average glucose in adults with CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(keyring)
library(tidyverse)
library(redcapAPI)
library(gtsummary)
library(labelled)
library(ggpmisc)
library(blandr)
home_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
    "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
    "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Documents/GitHub",
    "Windows" = "C:/Users/Tim/Documents/GitHub",
    "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# Import from REDCap
token <- key_get("envision_cf_hba1c_cgm")
rcon <- redcapConnection(
    url = "https://redcap.ahc.umn.edu/redcap/api/",
    token = token
)
df <- exportRecordsTyped(rcon)
# Start building analysis dataset
df <- df %>%
    # Format columns
    mutate(
        # Numeric
        across(
            c(age, bmi, cfrd_diagnosis_age, average_sensor_glucose),
            ~ as.numeric(.)
        ),
        across(c(hb_a1c_value, gmi), ~ as.numeric(sub("%", "", .))),
        # Categorical
        combined_race = factor(
            race,
            levels = c(
                "White",
                "Black or African American",
                "Asian",
                "More Than One Race",
                "Unknown / Not Reported"
            ),
            labels = c(
                "White",
                "Black or African American",
                "Other/Unknown",
                "Other/Unknown",
                "Other/Unknown"
            )
        )
    ) %>%
    # Drop unnecessary columns
    select(
        subject_id,
        age,
        cfrd_diagnosis_age,
        bmi,
        combined_race,
        sex,
        pi,
        cftr_modulator_use,
        type_of_modulator,
        type_of_cgm,
        hb_a1c_value,
        gmi,
        average_sensor_glucose
    ) %>%
    # Labels
    set_variable_labels(
        age = "Age (years)",
        cfrd_diagnosis_age = "Age at CFRD Diagnosis (years)",
        bmi = "BMI (kg/m2)",
        combined_race = "Race",
        sex = "Sex at Birth",
        type_of_modulator = "Type of Modulator",
        hb_a1c_value = "HbA1c (%)",
        gmi = "GMI (%)",
        average_sensor_glucose = "Mean Sensor Glucose (mg/dL)",
        cftr_modulator_use = "CFTR Modulator",
        pi = "Pancreatic Insufficiency"
    )
```

# Participant Characteristics

```{r}
#| label: tbl-1
#| tbl-cap: Table 1
df %>%
    select(-subject_id) %>%
    tbl_summary(
        missing_text = "Missing",
        type = all_continuous() ~ "continuous2",
        statistic = list(
            all_continuous() ~ c("{median} [{p25}, {p75}]", "{min} - {max}"),
            average_sensor_glucose ~ c("{mean} ({sd})", "{min} - {max}")
        ),
    )
```

# GMI and HbA1c

```{r}
#| warning: false
ggplot(df, aes(x = gmi, y = hb_a1c_value)) +
    geom_point() +
    geom_smooth(method = "lm", se = F) +
    stat_poly_line() +
    stat_poly_eq(use_label(c("eq", "R2", "p", "n")))

blandr.draw(df$gmi, df$hb_a1c_value, ciDisplay = F, annotate = T)
```

# Mean sensor glucose and HbA1c

```{r}
#| warning: false
ggplot(df, aes(x = average_sensor_glucose, y = hb_a1c_value)) +
    geom_point() +
    geom_smooth(method = "lm", se = F) +
    stat_poly_line() +
    stat_poly_eq(use_label(c("eq", "R2", "p", "n")))
```