### 2025-08-15 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
air formatting

---
title: "Racial and Ethnic Differences in ETI: Outcomes"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-height: 10
    fig-width: 10
    page-layout: full
editor: source
---

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(multcomp)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(broom)
library(emmeans)
home_dir <- switch(Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
    "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
    "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Documents/GitHub",
    "Windows" = "C:/Users/Tim/Documents/GitHub",
    "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load(
    "./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData"
)
# Change age group reference level
encounter$age_group <- relevel(encounter$age_group, ref = "[18,30)")
```

# Data cleaning

```{r flowchart}
ggflowchart(flow_chart, text_size = 3, y_nudge = 0.4, x_nudge = 0.45)
```

# Table 1

## By person

Continuous variables are reported at the most recent visit prior to ETI start for each person, and categorical variables are reported as the most common value.

```{r}
t1_participant %>%
    select(-eDWID) %>%
    tbl_summary(
        by = Race,
        type = list(`Years With Data` ~ "continuous"),
        statistic = list(
            c(gli_fev1_ppred_rn, gli_fvc_ppred_rn) ~ "{mean} [{sd}]"
        ),
        missing_text = "Missing"
    ) %>%
    add_overall()
```

# Encounter-level

```{r}
encounter %>%
    select(
        Race,
        encounterlocation,
        bacterialculturedone:othermicroorganisms1
    ) %>%
    tbl_summary(by = Race, missing_text = "Missing")
```

# Segmented models

```{r}
# Create change in slope variable
encounter$slope_c_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_c_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_c_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_c_years) <- "Change in Slope"
# Prior modulator use
prior_mods <- encounter %>%
    select(eDWID, Days, vx661comb, vx445comb, Vx770, Vx809comb) %>%
    filter(Days < 0) %>%
    group_by(eDWID) %>%
    summarise(
        prior_modulator = any(vx661comb == 1) |
            any(vx445comb == 1) |
            any(Vx770 == 1) |
            any(Vx809comb)
    )
prior_mods$prior_modulator <-
    fct_na_value_to_level(factor(prior_mods$prior_modulator), level = "No")
prior_mods$prior_modulator <- factor(
    prior_mods$prior_modulator,
    levels = c(T, "No"),
    labels = c("Yes", "No")
)
encounter <- left_join(encounter, prior_mods, by = join_by(eDWID))
# On enzymes for pancreatic insufficiency
encounter$isOnEnzymes <-
    factor(encounter$isOnEnzymes, levels = 0:1, labels = c("No", "Yes"))
# Pseudomonas
encounter$pseudomonasaeruginosa[is.na(encounter$pseudomonasaeruginosa)] <- 0
encounter$pseudomonasaeruginosa <- factor(
    encounter$pseudomonasaeruginosa,
    levels = 0:1,
    labels = c("No", "Yes")
)
```

The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject (random slope did not improve the model AIC).

## FEV1 % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
    Years = -3:2,
    Race = unique(encounter$Race),
    age_group = "[18,30)",
    Insurance = "Private",
    prior_modulator = "Yes",
    isOnEnzymes = "Yes",
    pseudomonasaeruginosa = "No"
))
pop_dat$slope_c_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("FEV1 % Predicted") +
    theme_classic(base_size = 15)
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt(caption = "Predicted Values at Various Timepoints") %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

## FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fvc_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fvc_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FVC_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FVC_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fvc_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fvc_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FVC_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FVC_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fvc_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fvc_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FVC_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FVC_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

## FEV1/FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1fvc_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1fvc_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1FVC_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        age_group +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1FVC_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1fvc_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1fvc_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1FVC_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1FVC_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    gli_fev1fvc_ppred_rn ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("gli_fev1fvc_ppred_rn") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    GLI_FEV1FVC_pct_predicted ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("GLI_FEV1FVC_pct_predicted") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(GLI_FEV1FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

## Number of Pulmonary Exacerbations

For some reason, exacerbation models with the encounter-level data are just not working. I think it may be because there are often several pulmonary exacerbation visits very close together, so including the "Days" variable is causing problems. In the end I used a very simple logistic regression model on annualized PEx data. See `/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R` for details on making the exacerbation dataset.

### Full cohort

```{r}
# Zero-inflated models
# mod_p <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.poisson(), zi_fixed = ~1
# )
# mod_nb <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.negative.binomial(), zi_fixed = ~1
# )
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data,
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

The table above shows the predicted probability of a pulmonary exacerbation pre- and post-ETI. So for example, NHW pwCF had a PEx probability of `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` post-ETI. Conversely, Black pwCF had a PEx probability of `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` post-ETI.

### Age < 18 at ETI initiation

```{r}
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data[pex_data$age_eti_group == "[-Inf,18)", ],
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Age >= 18 at ETI initiation

```{r}
pex_logistic <- glmer(
    any_pex ~ ETI * Race + (1 | eDWID),
    data = pex_data[pex_data$age_eti_group == "[18, Inf)", ],
    control = glmerControl(optimizer = "bobyqa"),
    family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
    gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

## BMI

### BMI %ile (age < 20 at ETI start)

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    bmipercentile ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_start < 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("BMI Percentile") +
    theme_classic(base_size = 15)
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(bmipercentile = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### BMI raw value (age >= 20 at ETI start)

```{r}
change_mod <- lmer(
    bmivalue ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_start >= 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("BMI (kg/m2)") +
    theme_classic(base_size = 15)
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(bmivalue = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

## Height

### Height %ile (age < 20 at ETI start)

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    heightpercentile ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_start < 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("heightpercentile") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(heightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Height raw value (age >= 20 at ETI start)

```{r}
change_mod <- lmer(
    height ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_start >= 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("height") +
    theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(height = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

## Weight

### Weight %ile (age < 20 at ETI start)

```{r}
# Fit mixed model with change point
change_mod <- lmer(
    weightpercentile ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_start < 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("weightpercentile") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(weightpercentile = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```

### Weight raw value (age >= 20 at ETI start)

```{r}
change_mod <- lmer(
    weight ~
        Years *
        Race +
        slope_c_years * Race +
        Insurance +
        prior_modulator +
        isOnEnzymes +
        pseudomonasaeruginosa +
        (1 | eDWID),
    data = encounter[encounter$age_eti_start >= 20, ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
    geom_point() +
    geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
    geom_line() +
    xlab("Years from ETI Start") +
    ylab("weight") +
    theme_classic()
# Model summary
change_mod %>%
    gtsummary::tbl_regression(intercept = T) %>%
    as_gt() %>%
    fmt_number(decimals = 3)
# Table of predicted values
pop_dat %>%
    select(-slope_c_years) %>%
    rename(weight = pred, ci.ll = ll, ci.ul = ul) %>%
    gt() %>%
    fmt_number(decimals = 3)
# Contrasts
aa_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
aa_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
hl_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
hl_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
nhw_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
nhw_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
other_pre <- c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
other_c <- c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
contr <- rbind(
    "Black/AA Pre-ETI" = aa_pre,
    "Black/AA Post-ETI" = aa_pre + aa_c,
    "H/L Pre-ETI" = hl_pre,
    "H/L Post-ETI" = hl_pre + hl_c,
    "NHW Pre-ETI" = nhw_pre,
    "NHW Post-ETI" = nhw_pre + nhw_c,
    "Other Pre-ETI" = other_pre,
    "Other Post-ETI" = other_pre + other_c
)
change_mod_slopes <- glht(change_mod, linfct = contr)
change_mod_slopes_ci <- confint(change_mod_slopes, calpha = univariate_calpha())
change_mod_slopes_ci$confint %>%
    as.data.frame() %>%
    gt(caption = "Slopes by Race", rownames_to_stub = T) %>%
    fmt_number(decimals = 3)
```