### 2025-09-16 Christine Chan/CF Registry/cfrd_1hr_ogtt_cgm_predictive.qmd

---
title: "Redefining CFRD Screening and Diagnosis in the Highly Effective Modulator Therapy Era (CF- SuGAR)"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(rspiro)
library(childsds)
library(tidyverse)
library(tableone)
library(pROC)
library(survival)
library(performance)
knitr::opts_knit$set(
  root.dir = "/Volumes/som/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo"
)
```

```{r}
#| include: false
# read in separate data files
ogtt <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/ogtt.csv")
pft <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/pfts.csv")
a1c <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/a1c.csv")
demo <- read.csv(
  "./Christine Chan/1 hour OGTT in CFRD/Data_raw/demographics.csv"
)
dob <- read.csv("./Christine Chan/1 hour OGTT in CFRD/Data_raw/dob.csv")
# merge dob and pfts
pft <- merge(pft, dob, by = "MRN", all.x = T, all.y = F)
pft <- unique(pft)
# checking for dups
# t <- ogtt[,c("MRN","FName","LName")]
# t <- unique(t)
# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
# merge ogtt and demo
demo$FName <- NULL
demo$LName <- NULL
ogtt <- ogtt %>%
  group_by(MRN, Date.of.OGTT) %>%
  filter(row_number() == 1)
ogtt_demo <- merge(ogtt, demo, by = "MRN", all.x = T, all.y = T)
ogtt_demo$Date.of.OGTT <- mdy(ogtt_demo$Date.of.OGTT)
a1c$EncounterDate <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
a1c$FName <- NULL
a1c$LName <- NULL
a1c$Date.of.OGTT <- NULL
a1c$Days.Between.A1C.and.OGTT <- NULL
ogtt_demo$EncounterDate <- ogtt_demo$Date.of.OGTT
ogtt_demo_a1c <- merge(
  ogtt_demo,
  a1c,
  by = c("MRN", "EncounterDate"),
  all.x = T,
  all.y = T
)

# find most recent FVC that is within the 90 days prior to OGTT
pft$EncounterDate <- as.Date(pft$Date.of.PFT, format = "%m/%d/%Y")
pft$FName <- NULL
pft$LName <- NULL
alldata <- merge(
  ogtt_demo_a1c,
  pft,
  by = c("MRN", "EncounterDate"),
  all.x = T,
  all.y = T
)
# get predicted spirometry values
alldata$FVC...Pred <- NULL
alldata$FEV1...Pred <- NULL
alldata$FEF.25.75...Pred <- NULL
# dob <- pft[,c("MRN","DOB")]
# dob <- unique(dob)
# alldata <- merge(alldata,dob,by="MRN",all.x = T,all.y = F)
alldata$DOB <- as.Date(as.character(alldata$DOB), format = "%m/%d/%Y")
alldata$ht_m <- alldata$Height..cm. / 100
alldata$sexnum <- ifelse(alldata$Sex == "Male", 1, 2)
alldata$FEV1.pred <- pred_GLI(
  age = alldata$Age.at.PFT,
  height = alldata$ht_m,
  gender = alldata$sexnum,
  param = "FEV1"
)
alldata$FVC.pred <- pred_GLI(
  age = alldata$Age.at.PFT,
  height = alldata$ht_m,
  gender = alldata$sexnum,
  param = "FVC"
)
alldata$FEF.25.75.pred <- pred_GLI(
  age = alldata$Age.at.PFT,
  height = alldata$ht_m,
  gender = alldata$sexnum,
  param = "FEF2575"
)
alldata$FEV1.per.pred <- (alldata$FEV1.Raw / alldata$FEV1.pred) * 100
alldata$FVC.per.pred <- (alldata$FVC.Raw / alldata$FVC.pred) * 100
alldata$FEF.25.75.per.pred <- (alldata$FEF.25.75.Raw / alldata$FEF.25.75.pred) *
  100
# alldata$Sex <- NULL
# hard code some corrections to data per CC
alldata[alldata$MRN == 686476, ]$X1.hour.Glucose <- ifelse(
  alldata[alldata$MRN == 686476, ]$Age.At.Test..years. == 14.5,
  185,
  alldata[alldata$MRN == 686476, ]$X1.hour.Glucose
)
# severity of genotype
# Red = severe or minimal function mutation
# Orange= unknown mutation status (for now)
# Yellow= partial function or mild
# Green = not CF causing, classify as "other", will most likely be seen in people with CRMS diagnosis
alldata$class2 <- ifelse(
  alldata$Mutation.2 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G"
    ),
  "I",
  ifelse(
    alldata$Mutation.2 %in%
      c(
        "F508",
        "F508del",
        "F508del ",
        "G85E",
        "I507",
        "I 507",
        "N1303K",
        "A559T",
        "R560T",
        "A561E"
      ),
    "II",
    ifelse(
      alldata$Mutation.2 %in% c("G551D", "G551S", "S549N"),
      "III",
      ifelse(
        alldata$Mutation.2 %in%
          c(
            "R334W",
            "R347P",
            "R117H",
            "R117C",
            "P67L",
            "L206W",
            "l206w",
            "I206w",
            "D614G",
            "R347H",
            "D1152H",
            "3849+10kbC>T"
          ),
        "IV",
        ifelse(
          alldata$Mutation.2 %in%
            c(
              "A455E",
              "2789+5G->A",
              "3849+10CT",
              "3849+10kbC->T",
              "711+3A->G",
              "1898+5G->T",
              "P574H",
              "3272-26A->G5",
              "3272-26A->G"
            ),
          "V",
          ""
        )
      )
    )
  )
)
alldata$severity2 <- ifelse(
  alldata$Mutation.2 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G",
      "F508",
      "F508del",
      "F508del ",
      "G85E",
      "I507",
      "I 507",
      "N1303K",
      "A559T",
      "R560T",
      "A561E",
      "G551D",
      "G551S",
      "S549N",
      "1898+1G->A",
      "2143delT",
      "2183AA-G",
      "2183delAA->G",
      "2184insA",
      "S434X",
      "2585delT",
      "3120+1G->A",
      "3905insT",
      "406-1G->A",
      "712-1G->T",
      "L1245X",
      "Q2X",
      "R1158X",
      "R851X",
      "V520F",
      "L1077P",
      "S489X",
      "F311del",
      "W846X",
      "M1101K",
      "2622+1G->A",
      "S945L",
      "Y1092X",
      "R1066C",
      "G178R",
      "2957delT",
      "CFTRdele1",
      "W1204X",
      "q493x",
      "W57G",
      "1585-8G>A",
      "1717-8G->A",
      "G1244E"
    ),
  "Severe",
  ifelse(
    alldata$Mutation.2 %in%
      c(
        "R334W",
        "R347P",
        "R117H",
        "R117C",
        "P67L",
        "L206W",
        "l206w",
        "I206w",
        "D614G",
        "R347H",
        "D1152H",
        "3849+10kbC>T",
        "D110H",
        "R1070W",
        "A455E",
        "2789+5G->A",
        "3849+10CT",
        "3849+10kbC->T",
        "711+3A->G",
        "1898+5G->T",
        "P574H",
        "3272-26A->G5",
        "3272-26A->G",
        "3849G->A",
        "H199Y",
        "H199y",
        "S492F",
        "S531P",
        "L1335P",
        "R352Q",
        "D1270N",
        "2789+2insA",
        "3849+10kbc->T",
        "Q98R",
        "Phe1052Val",
        "F1052V",
        "Tyr1014Cys",
        "Y1014C",
        "T338I",
        "5T",
        "c.2249C>T",
        "P750L"
      ),
    "Mild",
    ifelse(
      alldata$Mutation.2 %in%
        c(
          "1262insA",
          "I506T",
          "3363delGT",
          "Q1209P",
          "3349insT",
          "G178R",
          "L467P",
          "317insC",
          "I336K",
          "ex14a",
          "3500-2A->G",
          "c.3407_3422del16",
          "3349insT",
          "1248+1G->A",
          "M470V",
          "1812-1G->A",
          "del X22-24",
          "Q237H",
          "Q237H",
          "W57G",
          "1078delA",
          "G178R",
          "2957delT",
          "G576A",
          "other",
          "Other",
          "Unknown",
          "unknown",
          "CFTRdele1",
          "1811+1.6kb A->",
          "UNK",
          "I336K",
          "M470V",
          "S1235R",
          "M470V",
          "F508C",
          "R1066H",
          "P205S"
        ),
      "Unk",
      ""
    )
  )
)
alldata$class1 <- ifelse(
  alldata$Mutation.1 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G"
    ),
  "I",
  ifelse(
    alldata$Mutation.1 %in%
      c(
        "F508",
        "F508del",
        "F508del ",
        "G85E",
        "I507",
        "I 507",
        "N1303K",
        "A559T",
        "R560T",
        "A561E"
      ),
    "II",
    ifelse(
      alldata$Mutation.1 %in% c("G551D", "G551S", "S549N"),
      "III",
      ifelse(
        alldata$Mutation.1 %in%
          c(
            "R334W",
            "R347P",
            "R117H",
            "R117C",
            "P67L",
            "L206W",
            "l206w",
            "I206w",
            "D614G",
            "R347H",
            "D1152H",
            "3849+10kbC>T"
          ),
        "IV",
        ifelse(
          alldata$Mutation.1 %in%
            c(
              "A455E",
              "2789+5G->A",
              "3849+10CT",
              "3849+10kbC->T",
              "711+3A->G",
              "1898+5G->T",
              "P574H",
              "3272-26A->G5",
              "3272-26A->G"
            ),
          "V",
          ""
        )
      )
    )
  )
)
alldata$severity1 <- ifelse(
  alldata$Mutation.1 %in%
    c(
      "1717-1G->A",
      "1717-1G-A",
      "G542X",
      "Q493X",
      "R1162X",
      "R553X",
      "R553x",
      "W1089X",
      "W1282X",
      "1078delT",
      "R75X",
      "3659delC",
      "621+1G->T",
      "621+1G>T",
      "394delTT",
      "3120 + 1g->A",
      "1154InsTC",
      "1154insTC",
      "1213delT",
      "1259insA",
      "1288insTA",
      "3791delC",
      "E60X",
      "K710X",
      "2184delA",
      "CFTRdele2,3",
      "663delT",
      "Glu528",
      "1461ins4",
      "306insA",
      "R709X",
      "CFTRdele22-24",
      "2711delT",
      "2183AA- >G",
      "2183delAA>G",
      "F508",
      "F508del",
      "F508del ",
      "G85E",
      "I507",
      "I 507",
      "N1303K",
      "A559T",
      "R560T",
      "A561E",
      "G551D",
      "G551S",
      "S549N",
      "1898+1G->A",
      "2143delT",
      "2183AA-G",
      "2183delAA->G",
      "2184insA",
      "S434X",
      "2585delT",
      "3120+1G->A",
      "3905insT",
      "406-1G->A",
      "712-1G->T",
      "L1245X",
      "Q2X",
      "R1158X",
      "R851X",
      "V520F",
      "L1077P",
      "S489X",
      "F311del",
      "W846X",
      "M1101K",
      "2622+1G->A",
      "S945L",
      "Y1092X",
      "R1066C",
      "G178R",
      "2957delT",
      "CFTRdele1",
      "W1204X",
      "q493x",
      "W57G",
      "1585-8G>A",
      "1717-8G->A",
      "G1244E"
    ),
  "Severe",
  ifelse(
    alldata$Mutation.1 %in%
      c(
        "R334W",
        "R347P",
        "R117H",
        "R117C",
        "P67L",
        "L206W",
        "l206w",
        "I206w",
        "D614G",
        "R347H",
        "D1152H",
        "3849+10kbC>T",
        "D110H",
        "R1070W",
        "A455E",
        "2789+5G->A",
        "3849+10CT",
        "3849+10kbC->T",
        "711+3A->G",
        "1898+5G->T",
        "P574H",
        "3272-26A->G5",
        "3272-26A->G",
        "3849G->A",
        "H199Y",
        "H199y",
        "S492F",
        "S531P",
        "L1335P",
        "R352Q",
        "D1270N",
        "2789+2insA",
        "3849+10kbc->T",
        "Q98R",
        "Phe1052Val",
        "F1052V",
        "Tyr1014Cys",
        "Y1014C",
        "T338I",
        "5T",
        "c.2249C>T",
        "P750L"
      ),
    "Mild",
    ifelse(
      alldata$Mutation.1 %in%
        c(
          "1262insA",
          "I506T",
          "3363delGT",
          "Q1209P",
          "3349insT",
          "G178R",
          "L467P",
          "317insC",
          "I336K",
          "ex14a",
          "3500-2A->G",
          "c.3407_3422del16",
          "3349insT",
          "1248+1G->A",
          "M470V",
          "1812-1G->A",
          "del X22-24",
          "Q237H",
          "Q237H",
          "W57G",
          "1078delA",
          "G178R",
          "2957delT",
          "G576A",
          "other",
          "Other",
          "Unknown",
          "unknown",
          "CFTRdele1",
          "1811+1.6kb A->",
          "UNK",
          "I336K",
          "M470V",
          "S1235R",
          "M470V",
          "F508C",
          "R1066H",
          "P205S"
        ),
      "Unk",
      ""
    )
  )
)
# 3 patients have severity2="Severe" per Christine
alldata$severity2 <- ifelse(
  alldata$MRN %in% c(546678, 664400, 2015506),
  "Severe",
  alldata$severity2
)
alldata$genoRisk <- ifelse(
  alldata$severity1 == "Mild" | alldata$severity2 == "Mild",
  "Low",
  ifelse(
    alldata$severity1 == "Severe" & alldata$severity2 == "Severe",
    "High",
    ""
  )
)
# write file of people with missing genotype risk
# write.csv(alldata[alldata$genoRisk=="",],"./Christine Chan/1 hour OGTT in CFRD/Data_raw/missing_genotype_risk.csv")
# merge back all baseline characteristics
d <- alldata[, c(
  "MRN",
  "FName",
  "LName",
  "Sex",
  "Race",
  "Latino.a.",
  "Mutation.1",
  "Mutation.2",
  "class1",
  "class2",
  "severity1",
  "severity2",
  "genoRisk"
)]
d <- d[!is.na(d$FName), ]
d <- unique(d)
alldata$FName <- NULL
alldata$Latino.a. <- NULL
alldata$LName <- NULL
alldata$Sex <- NULL
alldata$Race <- NULL
alldata$Mutation.1 <- NULL
alldata$Mutation.2 <- NULL
alldata$class1 <- NULL
alldata$class2 <- NULL
alldata$severity1 <- NULL
alldata$severity2 <- NULL
alldata$genoRisk <- NULL
alldata <- merge(alldata, d, by = "MRN", all.x = T, all.y = T)
# checking for dups
# t <- alldata[,c("MRN","FName","LName")]
# t <- unique(t)
# ct <- t %>% group_by(LName,FName) %>% count(unique(MRN))
# find first OGTT
alldata$Date <- alldata$EncounterDate
o <- alldata[!is.na(alldata$X1.hour.Glucose), c("MRN", "Date")]
o <- o[order(o$MRN, o$Date), ]
o <- o %>%
  group_by(MRN) %>%
  filter(row_number() == 1)
colnames(o) <- c("MRN", "Date of first OGTT")
alldata <- merge(alldata, o, by = "MRN", all.x = T, all.y = T)
alldata <- alldata[!is.na(alldata$`Date of first OGTT`), ]
# checking ppts that Collin said had CFRD
col_check <- alldata[
  alldata$MRN %in%
    c(
      1006479,
      1053930,
      1078408,
      1136992,
      1289450,
      1339023,
      1413821,
      1695512,
      664400,
      686476,
      707831,
      712676,
      739231,
      739460,
      787372,
      800007,
      848440,
      860181,
      875128,
      897478,
      905284,
      908966,
      918124,
      932695,
      949042,
      952250,
      952252,
      825118
    ),
]
col_check <- col_check[order(col_check$MRN, col_check$Date), ]
# write.csv(col_check,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd.csv",na="",row.names = F)
# last 2 discrepant cases after fixing date issue
col_check2 <- alldata[alldata$MRN %in% c(686476, 825118), ]
col_check2 <- col_check2[order(col_check2$MRN, col_check2$Date), ]
# write.csv(col_check2,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd2.csv",na="",row.names = F)
# remove all visits before first OGTT
# alldata$`Date of first OGTT` <- as.Date(as.character(alldata$`Date of first OGTT`),format="%m/%d/%Y")
alldata <- alldata[alldata$Date >= alldata$`Date of first OGTT`, ]
alldata <- alldata %>% arrange(MRN, EncounterDate)
# calculate BMI
alldata$BMI <- alldata$Weight..kg. / ((alldata$Height..cm. / 100)^2)
# reorder columns
colorder <- c(
  "MRN",
  "FName",
  "LName",
  "Sex",
  "Race",
  "Latino.a.",
  "Mutation.1",
  "Mutation.2",
  "class1",
  "class2",
  "severity1",
  "severity2",
  "genoRisk",
  "Date of first OGTT",
  "Date",
  "Age.At.Test..years.",
  "Fasting.Glucose",
  "X1.hour.Glucose",
  "X2.hour.Glucose",
  "A1C",
  "DOB",
  "Age.at.PFT",
  "FVC.Raw",
  "Height..cm.",
  "Weight..kg.",
  "FEV1.Raw",
  "FEF.25.75.Raw",
  "ht_m",
  "sexnum",
  "FEV1.pred",
  "FVC.pred",
  "FEF.25.75.pred",
  "FEV1.per.pred",
  "FVC.per.pred",
  "FEF.25.75.per.pred",
  "BMI"
)
alldata <- alldata[, colorder]
alldata <- alldata[order(alldata$MRN, alldata$Date), ]
# create combined race/ethnicity
alldata$race_eth <- ifelse(
  alldata$Latino.a. == "TRUE",
  "Hispanic",
  alldata$Race
)
# get BMI percentile
alldata$age <- ifelse(
  is.na(alldata$Age.at.PFT),
  alldata$Age.At.Test..years.,
  alldata$Age.at.PFT
)
alldata$age_bmi <- ifelse(alldata$age > 20, 20, alldata$age)
alldata$bmip <- sds(
  alldata$BMI,
  age = alldata$age_bmi,
  sex = alldata$Sex,
  male = "Male",
  female = "Female",
  ref = cdc.ref,
  item = "bmi",
  type = "perc"
)
# for demographics data, multiply BMI percentile by 100
alldata$bmipercentile <- alldata$bmip * 100
# classify CFRD at each visit
alldata$cfrd_visit <- NA
alldata$cfrd_visit <- ifelse(
  is.na(alldata$X2.hour.Glucose),
  NA,
  ifelse(alldata$X2.hour.Glucose >= 200, 1, 0)
)
# find people with CFRD at baseline
cfrdbase <- alldata[
  alldata$Date == alldata$`Date of first OGTT` & alldata$cfrd_visit == 1,
]
before_cfrd <- length(unique(alldata$MRN))
alldata <- alldata[!(alldata$MRN %in% cfrdbase$MRN), ]
after_cfrd <- length(unique(alldata$MRN))
# number excluded with CFRD at baseline
nexcl_cfrd <- before_cfrd - after_cfrd
# create variable for CFRD groups (i.e., those who developed CFRD vs those who did not)
# and get date of first OGTT with CFRD
cfrdcheck <- alldata[alldata$cfrd_visit == 1, c("MRN", "Date", "cfrd_visit")]
cfrdcheck <- cfrdcheck[order(cfrdcheck$MRN, desc(cfrdcheck$cfrd_visit)), ]
cfrdcheck <- cfrdcheck[!is.na(cfrdcheck$MRN), ]
cfrdcheck <- cfrdcheck %>%
  group_by(MRN) %>%
  filter(row_number() == 1)
colnames(cfrdcheck) <- c("MRN", "Date of CFRD diagnosis", "CFRD")
alldata <- merge(alldata, cfrdcheck, by = "MRN", all.x = T, all.y = T)
alldata$CFRD <- ifelse(is.na(alldata$CFRD), 0, 1)
# get baseline age
baseage <- alldata[
  alldata$Date == alldata$`Date of first OGTT`,
  c("MRN", "age")
]
colnames(baseage) <- c("MRN", "Base_age")
alldata <- merge(alldata, baseage, by = "MRN", all.x = T, all.y = T)
# get baseline 1 hour and 2 hour glucose
basegluc <- alldata[
  alldata$Date == alldata$`Date of first OGTT`,
  c("MRN", "X1.hour.Glucose", "X2.hour.Glucose")
]
colnames(basegluc) <- c("MRN", "Baseline_1hr_glucose", "Baseline_2hr_glucose")
alldata <- merge(alldata, basegluc, by = "MRN", all.x = T, all.y = T)
# dichotomize baseline 1 hour glucose
alldata$Baseline_1hr_glucose_155 <- ifelse(
  alldata$Baseline_1hr_glucose >= 155,
  1,
  0
)
alldata$Baseline_1hr_glucose_140 <- ifelse(
  alldata$Baseline_1hr_glucose >= 140,
  1,
  0
)
# dichotomize baseline 2 hour glucose
alldata$Baseline_2hr_glucose_140 <- ifelse(
  alldata$Baseline_2hr_glucose >= 140,
  1,
  0
)
# find date of last OGTT
lastogtt <- alldata[
  !is.na(alldata$Fasting.Glucose) |
    !is.na(alldata$X1.hour.Glucose) |
    !is.na(alldata$X2.hour.Glucose),
  c("MRN", "Date")
]
lastogtt <- lastogtt[order(lastogtt$MRN, desc(lastogtt$Date)), ]
lastogtt <- lastogtt %>%
  group_by(MRN) %>%
  filter(row_number() == 1)
colnames(lastogtt) <- c("MRN", "Date of last OGTT")
alldata <- merge(alldata, lastogtt, by = "MRN", all.x = T, all.y = T)
# read in date of modulator use
mod <- read.csv(
  "./Christine Chan/1 hour OGTT in CFRD/Data_raw/modulators.csv",
  na.strings = c("", " ", NA)
)
mod$Kalydeco.Start.Date <- as.Date(mod$Kalydeco.Start.Date, format = "%m/%d/%Y")
mod$Orkambi.Start.Date <- as.Date(mod$Orkambi.Start.Date, format = "%m/%d/%Y")
mod$Symdeko.Start.Date <- as.Date(mod$Symdeko.Start.Date, format = "%m/%d/%Y")
mod$Trikafta.Start.Date <- as.Date(mod$Trikafta.Start.Date, format = "%m/%d/%Y")
mod$modstart <- as.Date(NA)
mod$he_modstart <- as.Date(NA)
mod$me_modstart <- as.Date(NA)
for (i in 1:nrow(mod)) {
  mod[i, ]$modstart <-
    as.Date(pmin(
      mod[i, ]$Kalydeco.Start.Date,
      mod[i, ]$Orkambi.Start.Date,
      mod[i, ]$Symdeko.Start.Date,
      mod[i, ]$Trikafta.Start.Date,
      na.rm = T
    ))
  mod[i, ]$he_modstart <- as.Date(pmin(
    mod[i, ]$Kalydeco.Start.Date,
    mod[i, ]$Trikafta.Start.Date,
    na.rm = T
  ))
  mod[i, ]$me_modstart <- as.Date(pmin(
    mod[i, ]$Orkambi.Start.Date,
    mod[i, ]$Symdeko.Start.Date,
    na.rm = T
  ))
}
modkeep <- mod[, c("MRN", "modstart", "he_modstart", "me_modstart")]
alldata <- merge(alldata, modkeep, by = "MRN", all.x = TRUE, all.y = FALSE)
alldata$onmod <- ifelse(
  !is.na(alldata$modstart) & alldata$modstart <= alldata$Date,
  1,
  ifelse(!is.na(alldata$modstart) & alldata$modstart > alldata$Date, 0, NA)
)
alldata$onmod <- ifelse(
  is.na(alldata$modstart) & is.na(alldata$onmod),
  0,
  alldata$onmod
)
alldata$onhemod <- ifelse(
  !is.na(alldata$he_modstart) & alldata$he_modstart <= alldata$Date,
  1,
  ifelse(
    !is.na(alldata$he_modstart) & alldata$he_modstart > alldata$Date,
    0,
    NA
  )
)
alldata$onhemod <- ifelse(
  is.na(alldata$he_modstart) & is.na(alldata$onhemod),
  0,
  alldata$onhemod
)
alldata$onmemod <- ifelse(
  !is.na(alldata$me_modstart) & alldata$me_modstart <= alldata$Date,
  1,
  ifelse(
    !is.na(alldata$me_modstart) & alldata$me_modstart > alldata$Date,
    0,
    NA
  )
)
alldata$onmemod <- ifelse(
  is.na(alldata$me_modstart) & is.na(alldata$onmemod),
  0,
  alldata$onmemod
)
alldata$FEV.1.per.pred_FVC.per.pred <- alldata$FEV1.per.pred /
  alldata$FVC.per.pred
# for analyses about prediction of CFRD, we need to exclude those who only have one OGTT
# can use alldata for mixed models of PFTs and BMI
oneogtt <- alldata[
  alldata$`Date of first OGTT` == alldata$`Date of last OGTT`,
]
multogtt <- alldata[!(alldata$MRN %in% oneogtt$MRN), ]
# have CFRD indicator
# create variable for follow-up time
# in those who do not develop CFRD, this is the time between first and last OGTT, and they are censored
# for those who develop CFRD, this is time between first OGTT and diagnosis of CFRD
multogtt$Follow_up_time_yrs <- ifelse(
  multogtt$CFRD == 1,
  (multogtt$`Date of CFRD diagnosis` - multogtt$`Date of first OGTT`) / 365.25,
  (multogtt$`Date of last OGTT` - multogtt$`Date of first OGTT`) / 365.25
)
# get age at diagnosis of CFRD
multogtt$age_at_CFRD_dx <- ifelse(
  multogtt$CFRD == 1,
  (multogtt$`Date of CFRD diagnosis` - multogtt$DOB) / 365.25,
  NA
)
# create categorical time
multogtt$cat_time <- as.factor(plyr::round_any(
  multogtt$Follow_up_time_yrs,
  0.5
))
# need a time variable for mixed models
multogtt$time <- as.numeric(
  (multogtt$Date - multogtt$`Date of first OGTT`) / 365.25
)
multogtt$cat_time2 <- as.factor(plyr::round_any(multogtt$time, 0.5))
# get number of participants in each cohort
n_alldata <- length(unique(alldata$MRN))
n_multogtt <- length(unique(multogtt$MRN))
# number of participants in each cohort who developed CFRD
n_alldata_cfrd <- length(unique(alldata[alldata$CFRD == 1, ]$MRN))
n_multogtt_cfrd <- length(unique(multogtt[multogtt$CFRD == 1, ]$MRN))
# create dataset with baseline only
baseline_mult <- multogtt[multogtt$Date == multogtt$`Date of first OGTT`, ]
baseline_mult$onmod_baseline <- as.factor(baseline_mult$onmod)
baseline_all <- alldata[alldata$Date == alldata$`Date of first OGTT`, ]
baseline_all$onmod_baseline <- as.factor(baseline_all$onmod)
# find patients who were put on modulators during follow-up
evermod <- alldata %>%
  group_by(MRN) %>%
  summarise(max(onmod))
baseline_all <- merge(baseline_all, evermod, by = "MRN", all.x = T, all.y = T)
baseline_all$modstarted_fup <- as.factor(ifelse(
  baseline_all$onmod_baseline == 0 & baseline_all$`max(onmod)` == 1,
  1,
  0
))
baseline_mult <- merge(baseline_mult, evermod, by = "MRN", all.x = T, all.y = F)
baseline_mult$modstarted_fup <- as.factor(ifelse(
  baseline_mult$onmod_baseline == 0 & baseline_mult$`max(onmod)` == 1,
  1,
  0
))
################## FILL IN MISSING A1c, BMI, PFT AT BASELINE
# find most recent A1c that is within the 90 days prior to OGTT or up to 14 days after OGTT
a1c$Date.of.A1C <- as.Date(a1c$Date.of.A1C, format = "%m/%d/%Y")
baseline_mult$A1C <- unlist(apply(baseline_mult, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- a1c %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.A1C, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$A1C[which.min(abs(difftime(d$Date.of.A1C, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
baseline_all$A1C <- unlist(apply(baseline_all, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- a1c %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.A1C, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$A1C[which.min(abs(difftime(d$Date.of.A1C, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
# find most recent FEV1 that is within the 90 days prior to OGTT or up to 14 days after OGTT
pfttemp <- alldata[!is.na(alldata$FEV1.per.pred), ]
pfttemp$Date.of.PFT <- as.Date(pfttemp$Date, format = "%m/%d/%Y")
baseline_mult$FEV1.per.pred <- unlist(apply(baseline_mult, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- pfttemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.PFT, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$FEV1.per.pred[which.min(abs(difftime(d$Date.of.PFT, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
baseline_all$FEV1.per.pred <- unlist(apply(baseline_all, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- pfttemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.PFT, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$FEV1.per.pred[which.min(abs(difftime(d$Date.of.PFT, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
# didn't add any PFTs so won't repeat code for FVC
# find most recent BMI that is within the 90 days prior to OGTT or up to 14 days after OGTT
bmitemp <- alldata[!is.na(alldata$BMI), ]
bmitemp$Date.of.BMI <- as.Date(bmitemp$Date, format = "%m/%d/%Y")
baseline_mult$BMI <- unlist(apply(baseline_mult, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- bmitemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.BMI, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$BMI[which.min(abs(difftime(d$Date.of.BMI, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
baseline_all$BMI <- unlist(apply(baseline_all, 1, function(r) {
  Date <- r["Date of first OGTT"]
  ID <- as.numeric(r["MRN"])
  d <- bmitemp %>% filter(MRN == ID)
  d$days <- difftime(d$Date.of.BMI, Date, units = "days")
  d <- d %>% filter(days <= 14 & days >= -90)
  a <- d$BMI[which.min(abs(difftime(d$Date.of.BMI, Date)))]
  if (is.na(Date) | nrow(d) == 0) {
    a <- NA
  }
  return(a)
}))
#############################################################
# create multogtt with time <4 years since there is little data beyond that
multogtt_reduced <- multogtt[multogtt$time <= 4, ]
##############
# DESC STATS #
##############
# table of demographics in those with at least one pre-CFRD OGTT
demovars <- c(
  "Sex",
  "race_eth",
  "age",
  "genoRisk",
  "Fasting.Glucose",
  "X1.hour.Glucose",
  "X2.hour.Glucose",
  "onmod_baseline",
  "modstarted_fup",
  "A1C",
  "FEV1.per.pred",
  "FVC.per.pred",
  "BMI",
  "bmipercentile",
  "FEV.1.per.pred_FVC.per.pred"
)
# Taking out these clinical variables - if they are to be added to table 1, need to figure out how
# because they come from PFT visits, not the same date as OGTT
#              "BMI","FEV1.per.pred","FVC.per.pred","FEF.25.75.per.pred","bmip")
t1_alldata <- CreateTableOne(data = baseline_all, vars = demovars)
t1_alldata <- print(
  t1_alldata,
  printToggle = F,
  varLabels = TRUE,
  missing = T,
  showAllLevels = T,
  test = F
)
# table of demographics in those with at least one pre-CFRD OGTT by sex
t1_alldata_sex <- CreateTableOne(
  data = baseline_all,
  vars = demovars,
  strata = "Sex"
)
t1_alldata_sex <- print(
  t1_alldata_sex,
  printToggle = F,
  varLabels = TRUE,
  missing = T,
  showAllLevels = T
)
# table of demographics in those with more than 1 OGTT
t1_multogtt <- CreateTableOne(
  data = baseline_mult,
  vars = c(demovars, "Follow_up_time_yrs")
)
t1_multogtt <- print(
  t1_multogtt,
  printToggle = F,
  varLabels = TRUE,
  missing = T,
  showAllLevels = T,
  test = F
)
# table of demographics in those with more than 1 OGTT stratified by progression
t1_multogtt_prog <- CreateTableOne(
  data = baseline_mult,
  vars = c(demovars, "Follow_up_time_yrs", "age_at_CFRD_dx"),
  strata = "CFRD"
)
t1_multogtt_prog <- print(
  t1_multogtt_prog,
  printToggle = F,
  varLabels = TRUE,
  missing = T,
  showAllLevels = T,
  test = T
)
# table of demographics in those with more than 1 OGTT by sex
t1_multogtt_sex <- CreateTableOne(
  data = baseline_mult,
  vars = demovars,
  strata = "Sex"
)
t1_multogtt_sex <- print(
  t1_multogtt_sex,
  printToggle = F,
  varLabels = TRUE,
  missing = T,
  showAllLevels = T
)
# compare baseline glucose values in those who developed CFRD and those who did not
# added other vars too
gluctable <- CreateTableOne(baseline_mult, vars = demovars, strata = "CFRD")
gluctable <- print(
  gluctable,
  printToggle = F,
  varLabels = TRUE,
  missing = T,
  showAllLevels = T,
  test = T
)
########################
# LOGISTIC REGRESSIONS #
########################
# logistic regression of 1 hour and 2 hour glucose predicting cfrd
log1 <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_1hr_glucose,
  family = "binomial"
)
or1 <- exp(cbind(OR = coef(log1), confint(log1)))
log1_adj <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_1hr_glucose + Base_age + Sex,
  family = "binomial"
)
or1_adj <- exp(cbind(OR = coef(log1_adj), confint(log1_adj)))
# add interaction with sex
log1_adj_int <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_1hr_glucose + Base_age + Sex + Sex * Baseline_1hr_glucose,
  family = "binomial"
)
or1_adj_int <- exp(cbind(OR = coef(log1_adj_int), confint(log1_adj_int)))
log1_cat <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_1hr_glucose_155,
  family = "binomial"
)
or1_cat <- exp(cbind(OR = coef(log1_cat), confint(log1_cat)))
log1_cat_adj <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_1hr_glucose_155 + Base_age + Sex,
  family = "binomial"
)
or1_cat_adj <- exp(cbind(OR = coef(log1_cat_adj), confint(log1_cat_adj)))
log1_cat_140 <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_1hr_glucose_140,
  family = "binomial"
)
or1_cat_140 <- exp(cbind(OR = coef(log1_cat_140), confint(log1_cat_140)))
log1_cat_adj_140 <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_1hr_glucose_140 + Base_age + Sex,
  family = "binomial"
)
or1_cat_adj_140 <- exp(cbind(
  OR = coef(log1_cat_adj_140),
  confint(log1_cat_adj_140)
))
# add interaction with sex
log1_cat_adj_int <- glm(
  data = baseline_mult,
  CFRD ~
    Baseline_1hr_glucose_155 + Base_age + Sex + Sex * Baseline_1hr_glucose_155,
  family = "binomial"
)
or1_cat_adj_int <- exp(cbind(
  OR = coef(log1_cat_adj_int),
  confint(log1_cat_adj_int)
))
log2 <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_2hr_glucose,
  family = "binomial"
)
or2 <- exp(cbind(OR = coef(log2), confint(log2)))
log2_adj <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_2hr_glucose + Base_age + Sex,
  family = "binomial"
)
or2_adj <- exp(cbind(OR = coef(log2_adj), confint(log2_adj)))
# add interaction with sex
log2_adj_int <- glm(
  data = baseline_mult,
  CFRD ~ Baseline_2hr_glucose + Base_age + Sex + Sex * Baseline_2hr_glucose,
  family = "binomial"
)
or2_adj_int <- exp(cbind(OR = coef(log2_adj_int), confint(log2_adj_int)))
# ROC for 1 hour glucose
# get predicted values from logistic regression
baseline_mult$pred_1hr <- predict(
  log1,
  type = "response",
  data.frame = baseline_mult
)
# perform the ROC analysis
roc1 <- roc(data = baseline_mult, response = CFRD, predictor = X1.hour.Glucose)
# find the cutoff that maximizes the Youden Index
coords1 <- coords(
  roc1,
  "b",
  ret = c("t", "specificity", "sensitivity"),
  best.method = "youden"
) # default
# ROC for 2 hour glucose
baseline_mult$pred_2hr <- predict(
  log2,
  type = "response",
  data.frame = baseline_mult
)
roc2 <- roc(data = baseline_mult, response = CFRD, predictor = X2.hour.Glucose)
coords2 <- coords(
  roc2,
  "b",
  ret = c("t", "specificity", "sensitivity"),
  best.method = "youden"
) # default
##############
# COX MODELS #
##############
# CPH model for baseline 1 hour glucose
cox1hr <- coxph(
  Surv(Follow_up_time_yrs, CFRD) ~ Baseline_1hr_glucose,
  data = baseline_mult
)
# Diagnostics
d1 <- cox.zph(cox1hr)
r2_coxsnell(cox1hr)
```

For the purposes of this power calculation, I took Laura Pyle's analysis code for the "Predictive Value of 1-Hour Glucose" paper @lorenzPredictiveValue1Hour2023 and re-fit the Cox proportional hazard model excluding pancreatic sufficient participants, in order to get an updated value for the apparent Cox-Snell $R^2$ ($R^2_{CSapp}$).