### 2025-12-23 Christine Chan/CF Registry/cf_registry_lung_function_analysis.qmd

---
title: "CF Registry Analysis - Lung Function"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-width: 12
    fig-height: 9
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(ggplot2)
library(gtsummary)
library(gt)
library(pROC)
library(ggrepel)
library(childsds)
library(labelled)
library(Hmisc)
library(lmerTest)
library(emmeans)
library(multcomp)
library(tidyverse)
library(parallel)
library(knitr)
base_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/home/tvigers/Documents/Data",
  "Tims-MacBook-Air.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
  "Tims-Mac-mini.local" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers",
  "Mac" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers"
)
data_dir <- switch(
  Sys.info()["nodename"],
  "togo" = "/CF/Christine Chan/CF Registry",
  "Tims-MacBook-Air.local" = "/CF/Christine Chan/CF Registry",
  "Tims-Mac-mini.local" = "/CF/Christine Chan/CF Registry",
  "Mac" = "/CF/Christine Chan/CF Registry"
)
home_dir <- paste0(base_dir, data_dir)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
#| warning: false
#| message: false
# Import data and codebook
setwd(home_dir)
annual <- read.csv(
  "./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
# Remove those with T1D or T2D
annual <- annual |> arrange(eDWID, ReviewYear)
annual <- annual[!annual$cfrd_status_annual %in% c(4, 5), ]
annual$cfrd_status_annual <- factor(
  annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv")
races <- c(
  "White",
  "Black or African American",
  "American Indian or Alaska Native",
  "Asian",
  "Native Hawaiian or Other Pacific ",
  "Some other race"
)
demogs$Race <- apply(demogs, 1, function(r) {
  w <- which(r[paste0("Race", 1:6)] == 1)
  if (length(w) > 1) {
    return("More than one race")
  } else {
    return(races[w])
  }
})
demogs$Race <- factor(
  demogs$Race,
  levels = c(
    "White",
    "American Indian or Alaska Native",
    "Asian",
    "Black or African American",
    "More than one race",
    "Native Hawaiian or Other Pacific ",
    "Some other race"
  ),
  labels = c(
    "White",
    "Other race",
    "Other race",
    "Black or African American",
    "More than one race",
    "Other race",
    "Other race"
  )
)
demogs$Hispanicrace <- factor(
  demogs$Hispanicrace,
  levels = 1:2,
  labels = c("Yes", "No")
)
demogs <- demogs |> select(eDWID, Gender, MutClass, Race, Hispanicrace)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Get BMI categories
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$bmi <- coalesce(annual$A_bmivalue, annual$bmi)
annual$bmi_perc_cdc <- sds(
  annual$bmi,
  age = annual$Age_YrEnd,
  sex = annual$Gender,
  male = "M",
  female = "F",
  item = "bmi",
  ref = cdc.ref,
  type = "perc"
) *
  100
annual$bmi_perc_who <- sds(
  annual$bmi,
  age = annual$Age_YrEnd,
  sex = annual$Gender,
  male = "M",
  female = "F",
  item = "bmi",
  ref = who.ref,
  type = "perc"
) *
  100
annual$bmi_perc <- coalesce(annual$bmi_perc_cdc, annual$bmi_perc_who)
annual$bmi_cat <- apply(annual, 1, function(r) {
  age <- as.numeric(r["Age_YrEnd"])
  gender <- as.character(r["Gender"])
  bmi <- as.numeric(r["bmi"])
  bmi_perc <- as.numeric(r["bmi_perc"])
  if (age <= 20) {
    return(cut(
      bmi_perc,
      c(-Inf, 10, 50, 85, Inf),
      right = F,
      labels = c("Undernourished", "At Risk", "Target/Goal", "Overweight")
    ))
  } else if (age > 20) {
    if (gender == "M") {
      return(cut(
        bmi,
        c(-Inf, 18.5, 23, 25, Inf),
        right = F,
        labels = c(
          "Undernourished",
          "At Risk",
          "Target/Goal",
          "Overweight"
        )
      ))
    } else if (gender == "F") {
      return(cut(
        bmi,
        c(-Inf, 18.5, 22, 25, Inf),
        right = F,
        labels = c(
          "Undernourished",
          "At Risk",
          "Target/Goal",
          "Overweight"
        )
      ))
    }
  }
})
annual$bmi_cat <- relevel(annual$bmi_cat, ref = "Target/Goal")
# Raw values for adults, percentiles for peds
annual$bmi_adult <- annual$bmi
annual$bmi_adult[annual$Age_YrEnd <= 20] <- NA
annual$bmi_perc <- annual$bmi_perc
annual$bmi_perc[annual$Age_YrEnd > 20] <- NA
# Labels
var_label(annual$bmi) = "BMI (all participants)"
var_label(annual$bmi_perc) = "BMI percentile (age <= 20)"
var_label(annual$bmi_adult) = "BMI value (age > 20)"
var_label(annual$bmi_cat) = "BMI category"
# Format columns
annual$MutClass <- factor(
  annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(
  annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf),
  right = F
)
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa",
  "A_fungalyeast1",
  "A_pulmonarycomplications1",
  "A_mycocultureresults3",
  "A_cultureresults3"
)
# Yes/No variables
annual$cfrd_treatment_options1[annual$cfrd_treatment_options1 == 2] <- NA
annual$cfrd_treatment_options4[annual$cfrd_treatment_options4 == 4] <- NA
yn_vars <- c(
  "A_corticosteroids1",
  "A_feeding2",
  "A_Vx770",
  "A_VX809comb",
  "liver_disease",
  bugs,
  "cfrd_treatment_options2",
  "cfrd_treatment_options3",
  "cfrd_treatment_options4"
)
annual[, yn_vars] <- lapply(
  annual[, yn_vars],
  factor,
  levels = 0:1,
  labels = c("No", "Yes")
)
annual$A_IsOnEnzymes <- factor(
  annual$A_IsOnEnzymes,
  levels = c(1, 0),
  labels = c("Yes", " No")
)
# Males to reference group
annual$Gender <- factor(annual$Gender, levels = c("M", "F"))
# Hispanic no as reference
annual$Hispanicrace <- factor(annual$Hispanicrace, levels = c("No", "Yes"))
# Get first year of CFRD
first_cfrd <- annual |>
  filter(cfrd_status_annual == "CFRD+") |>
  group_by(eDWID) |>
  summarise(first_cfrd = min(ReviewYear), age_dx = min(Age_YrEnd))
annual <- left_join(annual, first_cfrd, by = join_by(eDWID))
# Year prior to first CFRD in registry - assume negative for CFRD. Assume all
# years following are CFRD+
annual$CFRD <- NA
annual$CFRD[annual$ReviewYear < annual$first_cfrd] <- "CFRD-"
annual$CFRD[annual$ReviewYear >= annual$first_cfrd] <- "CFRD+"
annual$CFRD[is.na(annual$first_cfrd)] <- "CFRD-"
annual$CFRD <- factor(annual$CFRD, levels = c("CFRD-", "CFRD+"))
# Check whether they had an OGTT and/or HbA1c the year of first CFRD or the year
# prior. This is super slow but idc idc.
cl <- makeCluster(type = "FORK", 8)
cfrd_only <- parApply(
  cl,
  first_cfrd,
  1,
  function(r) {
    id <- as.numeric(r["eDWID"])
    year <- as.numeric(r["first_cfrd"])
    d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
    a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
    ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
    ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
    d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year + 1), ]
    treatment = any(
      d[, c(
        "cfrd_treatment_options2",
        "cfrd_treatment_options3",
        "cfrd_treatment_options4"
      )] ==
        "Yes",
      na.rm = TRUE
    )
    return(c(id, year, a1c, ogtt0hr, ogtt2hr, treatment))
  }
)
stopCluster(cl)
cfrd_only <- data.frame(t(cfrd_only))
colnames(cfrd_only) <- c(
  "eDWID",
  "dx_year",
  "a1c",
  "ogtt0hr",
  "ogtt2hr",
  "treatment"
)
# cfrd_only_store <- cfrd_only
cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
  not_missing <- paste(
    which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
    collapse = "_"
  )
  return(not_missing)
})
cfrd_only$a1c_ogtt <- factor(
  cfrd_only$a1c_ogtt,
  levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
  labels = c(
    "Neither",
    "HbA1c Only",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT",
    "OGTT"
  )
)
var_label(cfrd_only$a1c_ogtt) <- "Diagnosis Method"
cfrd_only$treatment = factor(
  cfrd_only$treatment,
  levels = 0:1,
  labels = c("No Treatment", "Treatment")
)
var_label(cfrd_only$treatment) = "CFRD Treatment Started Within 1 Year?"
```

# Diagnosed with CFRD between 2004 - 2018

## Participant characteristics at diagnosis

```{r}
# Create a model dataframe
mod_df = inner_join(
  annual,
  cfrd_only |> filter(dx_year >= 2004),
  by = join_by(eDWID)
) |>
  drop_na(a1c_ogtt) |>
  mutate(years_to_cfrd = ReviewYear - first_cfrd) |>
  filter(years_to_cfrd <= 5 & years_to_cfrd >= -5) |>
  mutate(years_to_cfrd = factor(years_to_cfrd))
var_label(mod_df$years_to_cfrd) = "Years From CFRD Dx"
# Write data for Christine to look at
mod_df |>
  select(
    eDWID,
    Race,
    Hispanicrace,
    MutClass,
    ReviewYear,
    Age_YrEnd,
    CFRD,
    first_cfrd,
    age_dx,
    A_weight,
    A_height,
    bmi_adult,
    bmi_perc,
    bmi_cat,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    a1c_ogtt,
    A_hgba1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    treatment,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    NumFEV1,
    NumFVC
  ) %>%
  write.csv(
    .,
    file = "./Data_Clean/fev1_model_data_2004.csv",
    row.names = FALSE,
    na = ""
  )
# Tables
mod_df |>
  filter(ReviewYear == first_cfrd) |>
  select(
    a1c_ogtt,
    age_dx,
    A_weight,
    A_height,
    bmi_adult,
    bmi_perc,
    bmi_cat,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    Race,
    Hispanicrace,
    MutClass,
    a1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    treatment,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    NumFEV1,
    NumFVC
  ) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p() |>
  modify_caption("**All Groups**")
mod_df |>
  filter(ReviewYear == first_cfrd, a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
  select(
    a1c_ogtt,
    age_dx,
    A_weight,
    A_height,
    A_bmivalue,
    A_bmipercentile,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    Race,
    Hispanicrace,
    MutClass,
    a1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    treatment,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    NumFEV1,
    NumFVC
  ) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(test = list(all_continuous() ~ "oneway.test")) |>
  modify_caption("**OGTT vs. HbA1c Only**")
```

## FEV1

### Not adjusted for treatment status

Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, and pancreatic enzyme use.

```{r}
# Fit the model
m <- lmer(
  A_FEV1_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFEV1 +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFEV1") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model tables
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Contrasts - probably a more elegant way to do this but oh well
neither_m_5 = rep(0, times = 33)
neither_m_5[1] = 1
neither_m_1 = rep(0, times = 33)
neither_m_1[5] = 1
neither_0 = rep(0, times = 33)
neither_0[6] = 1
neither_1 = rep(0, times = 33)
neither_1[7] = 1
neither_5 = rep(0, times = 33)
neither_5[11] = 1
hba1c_m_5 = rep(0, times = 33)
hba1c_m_5[12] = 1
hba1c_m_1 = rep(0, times = 33)
hba1c_m_1[16] = 1
hba1c_0 = rep(0, times = 33)
hba1c_0[17] = 1
hba1c_1 = rep(0, times = 33)
hba1c_1[18] = 1
hba1c_5 = rep(0, times = 33)
hba1c_5[22] = 1
ogtt_m_5 = rep(0, times = 33)
ogtt_m_5[23] = 1
ogtt_m_1 = rep(0, times = 33)
ogtt_m_1[27] = 1
ogtt_0 = rep(0, times = 33)
ogtt_0[28] = 1
ogtt_1 = rep(0, times = 33)
ogtt_1[29] = 1
ogtt_5 = rep(0, times = 33)
ogtt_5[33] = 1
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

### Adjusted for treatment status

Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, pancreatic enzyme use, and CFRD treatment.

```{r}
# Fit the model
m <- lmer(
  A_FEV1_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFEV1 +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    treatment +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFEV1") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model table
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

## FVC

### Not adjusted for treatment status

Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, and pancreatic enzyme use.

```{r}
# Fit the model
m <- lmer(
  A_FVC_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFVC +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFVC") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model tables
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

### Adjusted for treatment status

Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, pancreatic enzyme use, and CFRD treatment.

```{r}
# Fit the model
m <- lmer(
  A_FVC_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFVC +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    treatment +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFVC") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model tables
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

# Diagnosed with CFRD between 2010 - 2018

```{r}
# Create a model dataframe
mod_df = inner_join(
  annual,
  cfrd_only |> filter(dx_year >= 2010),
  by = join_by(eDWID)
) |>
  drop_na(a1c_ogtt) |>
  mutate(years_to_cfrd = ReviewYear - first_cfrd) |>
  filter(years_to_cfrd <= 5 & years_to_cfrd >= -5) |>
  mutate(years_to_cfrd = factor(years_to_cfrd))
var_label(mod_df$years_to_cfrd) = "Years From CFRD Dx"
```

## Participant characteristics at diagnosis

```{r}
# Write data for Christine to look at
mod_df |>
  select(
    eDWID,
    Race,
    Hispanicrace,
    MutClass,
    ReviewYear,
    Age_YrEnd,
    CFRD,
    first_cfrd,
    age_dx,
    A_weight,
    A_height,
    bmi_adult,
    bmi_perc,
    bmi_cat,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    a1c_ogtt,
    A_hgba1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    treatment,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    NumFEV1,
    NumFVC
  ) %>%
  write.csv(
    .,
    file = "./Data_Clean/fev1_model_data_2010.csv",
    row.names = FALSE,
    na = ""
  )
# Tables
mod_df |>
  filter(ReviewYear == first_cfrd) |>
  select(
    a1c_ogtt,
    age_dx,
    A_weight,
    A_height,
    bmi_adult,
    bmi_perc,
    bmi_cat,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    Race,
    Hispanicrace,
    MutClass,
    a1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    treatment,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    NumFEV1,
    NumFVC
  ) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p() |>
  modify_caption("**All Groups**")
mod_df |>
  filter(ReviewYear == first_cfrd) |>
  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) |>
  mutate(a1c_ogtt = droplevels(a1c_ogtt)) |>
  select(
    a1c_ogtt,
    age_dx,
    A_weight,
    A_height,
    A_bmivalue,
    A_bmipercentile,
    A_FEV1_pct_predicted,
    A_FVC_pct_predicted,
    Race,
    Hispanicrace,
    MutClass,
    a1c,
    A_ogttfast_bloodglucose,
    A_twohour_bloodglucose,
    A_IsOnEnzymes,
    liver_disease,
    A_pseudomonasaeruginosa,
    A_fungalyeast1,
    A_mycocultureresults3,
    A_cultureresults3,
    treatment,
    cfrd_treatment_options2,
    cfrd_treatment_options3,
    cfrd_treatment_options4,
    NumFEV1,
    NumFVC
  ) |>
  tbl_summary(
    by = a1c_ogtt,
    missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) |>
  add_overall() |>
  add_p(test = list(all_continuous() ~ "oneway.test")) |>
  modify_caption("**OGTT vs. HbA1c Only**")
```

## FEV1

### Not adjusted for treatment status

Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, and pancreatic enzyme use.

```{r}
# Fit the model
m <- lmer(
  A_FEV1_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFEV1 +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFEV1") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model table
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

### Adjusted for treatment status

Adjusted for age at diagnosis, number of FEV1 measures, sex, mutation class, pancreatic enzyme use, and CFRD treatment.

```{r}
# Fit the model
m <- lmer(
  A_FEV1_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFEV1 +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    treatment +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFEV1") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model table
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

## FVC

### Not adjusted for treatment status

Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, and pancreatic enzyme use.

```{r}
# Fit the model
m <- lmer(
  A_FVC_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFVC +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFVC") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model table
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.

### Adjusted for treatment status

Adjusted for age at diagnosis, number of FVC measures, sex, mutation class, pancreatic enzyme use, and CFRD treatment.

```{r}
# Fit the model
m <- lmer(
  A_FVC_pct_predicted ~ years_to_cfrd *
    a1c_ogtt +
    age_dx +
    NumFVC +
    Gender +
    MutClass +
    A_IsOnEnzymes +
    treatment +
    (1 | eDWID),
  data = mod_df
)
t3 = car::Anova(m, type = 'III', contrasts = "contr.sum")
emm = emmeans(m, specs = ~ years_to_cfrd + a1c_ogtt, pbkrtest.limit = 1e5)
ggplot(
  data.frame(emm),
  aes(x = years_to_cfrd, y = emmean, group = a1c_ogtt, color = a1c_ogtt)
) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.25) +
  geom_line() +
  xlab("Years From CFRD Dx") +
  ylab("Estimated Mean ppFVC") +
  labs(color = "Diagnosis Method") +
  theme_bw()
# Model table
m |> tbl_regression() |> add_n(location = "level")
```

The type 3 ANOVA indicates a significant interaction between time and diagnosis method (p =`r format.pval(t3[nrow(t3),"Pr(>Chisq)"], eps = 0.001)`).

```{r}
# Test contrasts
F <- contrast(
  emm,
  list(
    "Neither: -5 to 0 years" = neither_0 - neither_m_5,
    "Neither: 0 to 5 years" = neither_5 - neither_0,
    "Neither: -1 to 0 years" = neither_0 - neither_m_1,
    "Neither: 0 to 1 years" = neither_1 - neither_0,
    "HbA1c Only: -5 to 0 years" = hba1c_0 - hba1c_m_5,
    "HbA1c Only: 0 to 5 years" = hba1c_5 - hba1c_0,
    "HbA1c Only: -1 to 0 years" = hba1c_0 - hba1c_m_1,
    "HbA1c Only: 0 to 1 years" = hba1c_1 - hba1c_0,
    "OGTT: -5 to 0 years" = ogtt_0 - ogtt_m_5,
    "OGTT: 0 to 5 years" = ogtt_5 - ogtt_0,
    "OGTT: -1 to 0 years" = ogtt_0 - ogtt_m_1,
    "OGTT: 0 to 1 years" = ogtt_1 - ogtt_0
  )
)
kable(F, digits = 3)
kable(data.frame(emm), digits = 3)
```

The table above indicates the change in average lung function from a given timepoint to another within each group. For example, a value of -13.9 for "Neither: -5 to 0 years" indicates that among people without HbA1c or OGTT, lung function declined by 13 percentage points on average from 5 years prior to diagnosis to time of diagnosis.