### 2025-12-23 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd

---
title: "fsOGTT Pre- and Post-ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(tidyverse)
library(gtsummary)
library(labelled)
library(plotly)
library(lmerTest)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
setwd(home_dir)
# Import cleaned data
load("./Data_Clean/fsOGTT_pre_post_ETI.RData")
# Subset to people with more than one pre-ETI fsOGTT
multi_pre_df = df |>
  arrange(study_id, Date) |>
  group_by(study_id) |>
  mutate(
    n_pre = sum(pre_post == "Pre"),
    time_diff_first_last = as.numeric(difftime(
      last(Date),
      first(Date),
      units = "days"
    )),
    time_diff_last_first = as.numeric(difftime(
      first(Date[pre_post == "Post"]),
      last(Date[pre_post == "Pre"]),
      units = "days"
    ))
  ) |>
  filter(n_pre > 1)
var_label(multi_pre_df$n_pre) = "Number of fsOGTT pre-ETI"
var_label(multi_pre_df$time_diff_first_last) = "Days From First to Last fsOGTT"
var_label(
  multi_pre_df$time_diff_last_first
) = "Days From Last Pre-ETI fsOGTT to First Post-ETI fsOGTT"
# Calculate days between pre- and post-ETI fsOGTTs
times = multi_pre_df |>
  group_by(study_id, pre_post) |>
  reframe(d = difftime(Date, lag(Date), units = "days"))
# Save for Christine and Monique
write.csv(
  multi_pre_df,
  file = "./Data_Clean/fsOGTT_pre_post_ETI.csv",
  row.names = F,
  na = ""
)
```

# Participants with > 1 pre-ETI fsOGTT

## Table 1: Participant characteristics at first fsOGTT

The median number of days between pre-ETI fsOGTTs was `r as.numeric(summary(times$d[times$pre_post=="Pre"])[3])` (IQR: `r as.numeric(summary(times$d[times$pre_post=="Pre"])[2])`, `r as.numeric(summary(times$d[times$pre_post=="Pre"])[5])`). The median number of days between post-ETI fsOGTTs was `r as.numeric(summary(times$d[times$pre_post=="Post"])[3])` (IQR: `r as.numeric(summary(times$d[times$pre_post=="Post"])[2])`, `r as.numeric(summary(times$d[times$pre_post=="Post"])[5])`).

```{r}
#| label: tbl-1
multi_pre_df |>
  group_by(study_id) |>
  filter(row_number() == 1) |>
  ungroup() |>
  select(
    n_pre,
    time_diff_first_last,
    sex,
    age_visit,
    bmi,
    bmi_adult,
    bmi_perc,
    fev1,
    CFTR,
    pancreatic_status,
    Diagnosis,
    Insulin_0,
    Glucose_0,
    igi,
    iAUC30ins_over_gluc,
    iAUC120ins_over_gluc,
    iAUC180ins_over_gluc,
    odi_igi,
    odi_iauc,
    matsuda
  ) |>
  tbl_summary(type = c("Insulin_0") ~ "continuous")
```

## IGI (delta 30min insulin/delta 30min glucose)

```{r}
#| warning: false
# Plot
p = ggplot(
  multi_pre_df,
  aes(x = years_from_eti_start, y = igi, group = study_id, color = study_id)
) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplotly(p)
# Model
m = lmer(
  igi ~ years_from_eti_start_model + slope_change + (1 | study_id),
  data = multi_pre_df
)
m |>
  tbl_regression(
    exponentiate = F,
    estimate_fun = label_style_sigfig(digits = 4),
    pvalue_fun = label_style_pvalue(digits = 2),
  )
```

## iAUC30ins/iAUC30gluc

```{r}
#| warning: false
# Plot
p = ggplot(
  multi_pre_df,
  aes(
    x = years_from_eti_start,
    y = iAUC30ins_over_gluc,
    group = study_id,
    color = study_id
  )
) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplotly(p)
# Model
m = lmer(
  iAUC30ins_over_gluc ~ years_from_eti_start_model +
    slope_change +
    (1 | study_id),
  data = multi_pre_df
)
m |>
  tbl_regression(
    exponentiate = F,
    estimate_fun = label_style_sigfig(digits = 4),
    pvalue_fun = label_style_pvalue(digits = 2),
  )
```

## iAUC120ins/iAUC120gluc

```{r}
#| warning: false
# Plot
p = ggplot(
  multi_pre_df,
  aes(
    x = years_from_eti_start,
    y = iAUC120ins_over_gluc,
    group = study_id,
    color = study_id
  )
) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplotly(p)
# Model
m = lmer(
  iAUC120ins_over_gluc ~ years_from_eti_start_model +
    slope_change +
    (1 | study_id),
  data = multi_pre_df
)
m |>
  tbl_regression(
    exponentiate = F,
    estimate_fun = label_style_sigfig(digits = 4),
    pvalue_fun = label_style_pvalue(digits = 2),
  )
```

## iAUC180ins/iAUC180gluc

```{r}
#| warning: false
# Plot
p = ggplot(
  multi_pre_df,
  aes(
    x = years_from_eti_start,
    y = iAUC180ins_over_gluc,
    group = study_id,
    color = study_id
  )
) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplotly(p)
# Model
m = lmer(
  iAUC180ins_over_gluc ~ years_from_eti_start_model +
    slope_change +
    (1 | study_id),
  data = multi_pre_df
)
m |>
  tbl_regression(
    exponentiate = F,
    estimate_fun = label_style_sigfig(digits = 4),
    pvalue_fun = label_style_pvalue(digits = 2),
  )
```

## ODI (IGI*Matsuda)

```{r}
#| warning: false
# Plot
p = ggplot(
  multi_pre_df,
  aes(x = years_from_eti_start, y = odi_igi, group = study_id, color = study_id)
) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplotly(p)
# Model
m = lmer(
  odi_igi ~ years_from_eti_start_model + slope_change + (1 | study_id),
  data = multi_pre_df
)
m |>
  tbl_regression(
    exponentiate = F,
    estimate_fun = label_style_sigfig(digits = 4),
    pvalue_fun = label_style_pvalue(digits = 2),
  )
```

## ODI [(iAUC30ins/iAUC30gluc)*Matsuda]

```{r}
#| warning: false
# Plot
p = ggplot(
  multi_pre_df,
  aes(
    x = years_from_eti_start,
    y = odi_iauc,
    group = study_id,
    color = study_id
  )
) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplotly(p)
# Model
m = lmer(
  odi_iauc ~ years_from_eti_start_model + slope_change + (1 | study_id),
  data = multi_pre_df
)
m |>
  tbl_regression(
    exponentiate = F,
    estimate_fun = label_style_sigfig(digits = 4),
    pvalue_fun = label_style_pvalue(digits = 2),
  )
```

## Matsuda

```{r}
#| warning: false
# Plot
p = ggplot(
  multi_pre_df,
  aes(
    x = years_from_eti_start,
    y = matsuda,
    group = study_id,
    color = study_id
  )
) +
  geom_point() +
  geom_line() +
  theme_bw()
ggplotly(p)
# Model
m = lmer(
  matsuda ~ years_from_eti_start_model + slope_change + (1 | study_id),
  data = multi_pre_df
)
m |>
  tbl_regression(
    exponentiate = F,
    estimate_fun = label_style_sigfig(digits = 4),
    pvalue_fun = label_style_pvalue(digits = 2),
  )
```